### 2021-11-02 inst/doc/my-vignette.Rmd
version 0.0.1

---
title: "MultiATSM package - General Guidelines"
author: Rubens Moura
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    number_sections: true
    df_print: paged
    toc: yes
    toc_depth: 2
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Vignette for package MultiATSM}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc} 
bibliography: references.bib
---

This document aims at providing general guidance on the use of the *MultiATSM* package available at the [CRAN repository](https://cran.r-project.org/).

```{r}
library(MultiATSM)
```

The *MultiATSM* package provides several model outputs from various single and multi-country affine term structure of interest models (ATSMs). All the frameworks of this package are based on the unspanned economic risk framework from @JoslinPriebschSingleton2014 (JPS, 2014). In essence, these models assume the absence of arbitrage opportunities, consider a linear state space representations of the yield curve dynamics and offer a tractable approach to simultaneously combine the traditional yield curve factors (spanned factors) along with economic and financial variables (unspanned factors).

Due to the peculiar features of the JPS specification, an efficient estimation of the parameters governing the risk-neutral ($Q$-measure) and the physical ($P$-measure) probability measures can be carried out rather independently. The only exception is the variance-covariance matrix (sigma) term which is a common element to both the $P$ and the $Q$ density functions.

In its current version, the *MultiATSM* package can generate the outputs of 8 different classes of ATSMs. More specifically, the *MultiATSM* includes models in which the estimation is done on a *country-by-country* basis - as in JPS (2014)  - or *jointly* for all the countries of the economic system - as in @JotikasthiraLeLundblad2015 (JLL, 2015) and @CandelonMoura2021 (CM, 2021).  In all cases, the risk factors dynamics under the $P$-measure follows a VAR(1) model of some sort. The table below summarizes the general features of each one of the models available at this package.

```{r ModFea, message=FALSE, echo=FALSE}
ModelLabels <- c("JPS", "JPS jointP", "VAR jointQ", "GVAR sepQ", "GVAR jointQ", 
                 "JLL original", "JLL NoDomUnit", "JLL jointSigma")

# Rows
Tab <- data.frame(matrix(nrow = length(ModelLabels), ncol = 0)) 
rownames(Tab) <- ModelLabels

# Empty columns
EmptyCol <- c("", "", "", "", "", "", "", "") 
Tab$EmptyCol0 <- EmptyCol
# P-dynamics + 2 empty spaces
Tab$PdynIndUnco <- c("x", "x", "", "", "", "", "", "")
Tab$PdynIndCo <- c("", "", "", "", "", "", "", "")
Tab$PdynJointUnco <- c("", "", "x", "", "", "", "", "")
Tab$PdynJointJLL <- c("", "", "", "", "", "x", "x", "x")
Tab$PdynJointGVAR <- c("", "", "", "x", "x", "", "", "")
Tab$EmptyCol1 <- EmptyCol
Tab$EmptyCol2 <- EmptyCol
# Q-dynamics + 2 empty spaces
Tab$QdynInd <- c("x", "", "", "x", "", "", "", "")   
Tab$QdynJoint <- c("", "x", "x", "", "x", "x", "x", "x") 
Tab$EmptyCol3 <- EmptyCol
Tab$EmptyCol4 <- EmptyCol
# Sigma + 2 empty spaces
Tab$Ponly <-  c("", "", "", "", "", "x", "x", "")
Tab$PandQ <- c("x", "x", "x", "x", "x", "", "", "x")
Tab$EmptyCol5 <- EmptyCol
Tab$EmptyCol6 <- EmptyCol
# Dominant Unit
Tab$DomUnit <- c("", "", "", "", "", "x", "", "x")

# Adjust column names
ColNames <- c("","","","","JLL", "GVAR", "", "", "", "", "", "", "","", "", "","")
colnames(Tab) <- ColNames

# Generate the table
suppressWarnings(library(magrittr))

kableExtra::kbl(Tab, align = "c", caption = "Model Features") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 10) %>%
  kableExtra::add_header_above(c(" "=2, "Unrestricted" = 1, "Restricted" = 1, "Unrestricted" = 1, "Restricted" = 2, " " = 11)) %>%
 kableExtra::add_header_above(c(" "=2, "Individual" = 2, "Joint" = 3, " "=2, "Individual" = 1, "Joint" = 1, " "=2, "P only" = 1, "P and Q" = 1, " " = 3))  %>%
  kableExtra::add_header_above(c( " "=2, "P-dynamics"= 5, " "=2, "Q-dynamics"= 2, " "=2, "Sigma matrix estimation" = 2, " "=2, "Dominant Country"=1), bold = T) %>%
kableExtra::pack_rows("Unrestricted VAR", 1, 3 , label_row_css = "background-color: #666; color: #fff;")  %>%
kableExtra::pack_rows("Restricted VAR (GVAR)", 4, 5, label_row_css = "background-color: #666; color: #fff;") %>%
kableExtra::pack_rows("Restricted VAR (JLL)", 6, 8, label_row_css = "background-color: #666; color: #fff;")
```

Some aspects of the multi-country frameworks are worth highlighting. As for the models based on the setup of JLL (2015), the version "JLL original" follows closely the seminal work of JLL (2015), i.e., it is assumed an economic cohort containing a worldwide dominant economy and a set of smaller countries, in addition to the estimation of the sigma matrix be performed exclusively under the $P$-measure. The two other alternative versions assume the absence of a dominant country ("JLL NoDomUnit") and the estimation of sigma under both the $P$ and $Q$ measures ("JLL jointSigma"), as in the standard JPS (2014) model. As for the remaining multi-country ATSMs, it is considered that the dynamics of the risk factors under the $P$-measure evolves according to a GVAR model. The version labeled "GVAR jointQ" is the one presented in CM (2021).

In what follows, the focus will be devoted to describe the various pieces that are necessary to implement ATSMs. Specifically, Section \@ref(S:SectionData) describes the data-set available at this package and the set of functions that are useful to retrieve data from Excel files. Section \@ref(S:SectionInputs) exposes the necessary inputs that have to be specified by the user. In Section \@ref(S:SectionEstimation) the estimation procedure is detailed. Section \@ref(S:SectionReplication) shows how to use the *MultiATSM* package to estimate ATSMs from scratch as well as to closely reproduce some empirical features of academic papers. To ease the exposition, throughout the next sections, the database used by CM (2021) is employed to illustrate the use of the various functions available in this package.

# Data {#S:SectionData}

## Package data-set{#S:SectionDataPackage}
The *MultiATSM* package contains the four data-sets used in CM (2021). The first set of data comprises several time series of zero-coupon bonds yields from four emerging markets: China, Brazil, Mexico, and Uruguay. It is worth noting that this package requires *(i)* for estimation purposes, the maturities of the bond yields to be the same for all countries (although the function *DataForEstimation* handles different maturities across countries as inputs, the outputs generated by this same function are bond yields of the same maturities for all the economies); *(ii)* bond yields to be expressed in percentage terms (and not in basis points) per annum. The *MultiATSM* package does not support a routine to bootstrap zero-coupon yields from coupon bonds. This data manipulation procedure must be handled by the user herself, if necessary.  
```{r}
data('CM_Yields')
```

The second group of data concerns the time series of the risk factors. Specifically, along the terminology defined in JPS (2014), this data-set contains *(i)* country-specific spanned factors (denoted by level, slope, and curvature - see Section \@ref(S:SpaFac) for a more detailed description) and *(ii)* an array of country-specific and global unspanned factors (namely, some measure of economic growth and inflation). Similarly to the case of the bond yields data, the measures of economic and financial variables must be constructed by the user.
```{r}
data('CM_Factors')
```

The two last block of data are necessary for the estimation of the GVAR-based models. The trade flows database
presents the sum of the value of all goods imports and exports between any two countries of the sample on yearly basis since 1948. All the values are free on board and are expressed in U.S. dollars. This data is used to construct the transition matrix of the GVARs models.  
```{r}
data('CM_Trade')
```

The GVAR factors database casts country-specific lists of all the factors that are used in the estimation of each country's VARX(1,1,1). A specific function for computing the star risk factors is detailed the Section \@ref(S:SectionGVAR). 
```{r}
data('CM_Factors_GVAR')
```

## Import data from Excel files {#S:SectionImport}
This package also offers an automatized procedure to extract data from Excel files and to, subsequentially, prepare the risk factors database that are directly used in the model estimation. The use of the package functions requires that the databases *(i)* are constructed in separate Excel files for the unspanned factors, the term structure data, and the measures of interdependence (for the GVAR-based models); *(ii)* contain, in each Excel file, one separate tab per country (in addition to the global variables, in the case of the unspanned factors database); and *(iii)* have identical variable labels across all the tabs within each file. For illustration, see the Excel file available at the package.
One example of list of inputs to be provided is, for instance,
```{r}
Initial_Date <- "2006-09-01" # Format "yyyy-mm-dd"
Final_Date <- "2019-01-01" # Format "yyyy-mm-dd"
DataFrequency <- "Monthly"
GlobalVar <- c("GBC", "VIX") # Global Variables
DomVar <- c("Eco_Act", "Inflation", "Com_Prices", "Exc_Rates") #  Domestic variables
N <-  3 # Number of spanned factors per country
Economies <- c("China", "Mexico", "Uruguay", "Brazil", "Russia")
ModelType <- "JPS"
```

Based on these inputs, one can construct the variable *ZZfull* which contains the complete set of the model risk factors.
```{r}
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
ZZfull <-DataForEstimation(Initial_Date, Final_Date, Economies, N, FactorLabels, ModelType, DataFrequency)
```

# Required user inputs {#S:SectionInputs}

## Basic inputs used in the model estimation

In order to estimate any ATSM of this package, the user needs to specify several general model inputs, namely:

- **Model type**: a string-vector containing the label of the model to be estimated (as per described in Table \@ref(tab:ModFea)); 
- **Frequency of the data**: a string-vector that specifies the frequency of the time series data. The available options are: *"Annually"*, *"Quarterly"*, *"Monthly"*, *"Weekly"*, *"Daily Business Days"*, and *"Daily All Days"*;
- **Economies**: a string-vector containing the names of the economies which are part of the economic system;
- **Number of spanned factors (N)**: number of country-specific spanned factors. In this package, *N* is assumed to be equal across countries;
- **Global variables**: a string-vector containing the labels of the global unspanned factors used in the model estimation;
- **Domestic variables**: a string-vector containing the labels of the domestic unspanned factors used in the model estimation;
- **Bond yield set of maturities**: a numerical-vector containing the maturities of the yields used in the model estimation. The units of this vector must be expressed in *years*. Alternatively, the user can rely on the *Maturities* routine to extract the the bond maturities form the time series of bond yields (see the documentation of the *Maturities* function). In this case, the user must specify the **unit in which the label of the bond yields** are expressed, e.g. if the label of the 10-years bond is selected as *Y120M* (*Y10y*), then the option *Month* (*Year*) must be selected;
- **Factor labels**: a string-list based which contains the labels of all the variables present in the model. This input is particularly useful for constructing the plots. The routine *LabFac* provides the desired list of factor labels;
- **Stationarity constraint under the $Q$-dynamics**: a binary variable which takes value equal to "1" if the user wishes the largest eigenvalue under the $Q$-measure to be strictly smaller than 1. Otherwise, this same variable must be set to "0";
- **Output label**: a single element string-vector which contains the of the name that appears in the file name that stores the model outputs.

One possible example of the basic model inputs is
```{r}
ModelType <- "JPS"
DataFrequency <- "Monthly"
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
N <- 3 # number of spanned factors per country
GlobalVar <- c("GBC", "CPI_OECD") # Global variables
DomVar <- c("Eco_Act", "Inflation") # Domestic variables 
mat <- c(0.25, 0.5, 1, 3, 5, 10) # vector of maturities
FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) 
StationarityUnderQ <- 0 # 1 = set stationarity condition under  the Q; 0 = no stationarity condition
OutputLabel <- "Model_demo" # output label
```

### GVARinputs and JLLinputs {#S:SectionGVARJLLinputs}
Some additional inputs are required if the user intends to estimate the GVAR-ATSM or JLL-ATSM related setups. For a convenient use, the extra outputs should be stored in a separate list for each model. The GVAR model list of inputs must contain:

- **Economies**: a string-vector containing the names of the economies as described above;  
- **GVAR list of risk factors**:  a list of risk factors sorted by country in addition to the global variables. See function *DatabasePrep*;
- **VARX type**: a string-vector containing the desired estimated form of the VARX(1,1,1). Two possibilities are available. The *"unconstrained"* form estimates the model without any constrains using standard OLS regressions for each one of the model equations. The *"constrained"* form prevents foreign-spanned-factors to impact any domestic factor. In such a case, the VARX(1,1,1) is estimated by restricted least squares.

```{r}
GVARinputs <- list()
GVARinputs$Economies <- Economies
GVARinputs$GVARFactors <- FactorsGVAR
GVARinputs$VARXtype <- "unconstrained"
```

Furthermore, the estimation of the GVAR requires the specification of the necessary inputs to build the **transition matrix** (see Section \@ref(S:SectionGVAR) for more details). One example of this list of inputs is
```{r}
t_First <- "2000" # First year of the sample
t_Last <-  "2015" # Last year of the sample
W_type <- 'Sample Mean' 
```

Concerning the JLL-ATSM frameworks, the list of inputs includes:

- **JLL type**: a string-vector containing the label of the JLL model to be estimated (as described in Table \@ref(tab:ModFea));
- **Economies**: a string-vector containing the names of the economies as described above;
- **Dominant Country**: a string-vector containing the name of the economy which is assigned as the dominant country (applicable for the *JLL original* and *JLL jointSigma* models) or "None" (applicable for the *JLL NoDomUnit*).
 - **Wish the estimation of the sigmas matrices**: this is a binary variable which assume value equal to 1 if the user wishes the estimation of all JLL sigma matrices (i.e. variance-covariance and the Cholesky factorization matrices) and, 0 otherwise. The estimation of the sigma matrices can take several minutes;
- **Sigma of the non-orthogonalized variance-covariance matrix**: to save time, the user may provide the variance-covariance matrix from the non-orthogonalized dynamics. Otherwise, this input should be set as *NULL*.

One possible list of JLL inputs is as follows:
```{r}
  JLLinputs <- list()
  ModelType <- "JLL original"  
  JLLinputs$Economies <- Economies
  JLLinputs$DomUnit <- "China"
  JLLinputs$WishSigmas <- 1 
  JLLinputs$SigmaNonOrtho <- NULL
  JLLinputs$JLLModelType <- ModelType
```

## Additional inputs for numerical and graphical outputs
Once the model parameters from the ATSM have been estimated, the *MultiATSM* package enables the numerical and graphical compilation of the following additional outputs:

- Model fit of the bond yields;
- Impulse response functions (IRFs);
- Forecast error variance decompositions (FEVDs);
- Generalized impulse response functions (GIRFs);
- Generalized forecast error variance decompositions (GFEVDs).

For the IRFs, GIRFs, FEVDs or GFEVDs, a horizon of analysis has to be specified, e.g.:
```{r}
Horiz <- 100
```

For the graphical outputs, the user must indicate the desired types of graphs in a string-based vector:
```{r}
DesiredGraphs <- c("Fit", "GIRF", "GFEVD") # Available options are: "Fit", "IRF", "FEVD", "GIRF", "GFEVD"
```

Moreover, the user must select the types of variables of interest (yields, risk factors or both) and, for the JLL type of models, whether the orthogonalized version should be additionally included   
```{r}
WishGraphRiskFac <- 0 #   YES: 1; No = 0.
WishGraphYields <- 1 #    YES: 1; No = 0.
WishOrthoJLLgraphs <- 0 # YES: 1; No = 0.
```

The function *InputsForOutputs* can provide some guidance for customizing the features of the wished outputs. Conditional to these settings, individual folders are created at your *temporary directory* to store the different types of the desired graphical outputs.

### Bootstrap settings
If the user intends to generate confidence bands through some bootstrap procedure, an additional list of inputs is required. Specifically:

- **methodBS**: the desired bootstrap procedure. Available options are *(i)* standard residual bootstrap ('bs'); *(ii)* wild bootstrap ('wild') and block bootstrap ('block'). If the latest is selected, then the block length must be indicated;
- **ndraws**:  the number of bootstrap draws;
- **pctg**: the confidence level expressed in percentage points.

```{r}
Bootlist <- list()
Bootlist$methodBS <- 'block' 
Bootlist$BlockLength <- 4 
Bootlist$ndraws <-  3  
Bootlist$pctg   <-  95 
```

### Out-of-sample forecast settings
To perform out-of-sample forecasts, the following list-based features have to be detailed:

- **ForHoriz**: number of periods-ahead that the forecasts are to be generated;
- **t0Sample**: time-dimension index of the first observations belonging to the information set;
- **t0Forecast**: time-dimension index of the last observation of the information set used to perform the first forecast set.

```{r}
ForecastList <- list()
ForecastList$ForHoriz <- 12 # forecast horizon
ForecastList$t0Sample <- 1 # initial sample date
ForecastList$t0Forecast <- 70 # last sample date for the first forecast
```

# Model Estimation {#S:SectionEstimation}
Having gathered bond yields and other economic time series data, the ATSM estimation is carried out in three steps. Firstly, one needs to provide the number of country-specific spanned factors to be included in the global ATSM. Secondly, it is decided on the form of the risk factor dynamics under the $P$-measure. Finally, the user needs to select general features for the model optimization.

## Spanned Factors {#S:SpaFac}
The spanned factors are yield-related variables that are used to fit the cross-section dimensions of the term structures. Typically, the spanned factors of country $i$, $P_{i,t}$, are computed as the *N* first principal components (PCs) of the set of observed bond yields. Formally, $P_{i,t}$ is constructed as $P_{i,t} = w_i Y_{i,t}$ where $w_i$ is the PC weight matrix and $Y_{i,t}$ is a country-specific column-vector of yields with increasing maturities.

For *N=3*, the spanned factors are traditionally interpreted as level, slope, and curvature. The nature of such interpretability results from the features of the PC weight matrix as illustrated below:

```{r}
w <- pca_weights_one_country(Yields, Economy = "Uruguay") 
```

In matrix *w*, each row stores the weights used for constructing each spanned factor. The entries of the first row are linked to the composition of the *level* factor in that they load rather equally on all yields. Accordingly, high (low) values of the level factor indicate an overall high (low) value of yields across all maturities. In the second row, the weights monotonically increase with the maturities and, therefore, they capture the degree of steepness (*slope*) of the term structure. High slope factor values imply a steep yield curve, whereas low ones entail flat (or, possibly, downward) curves. In the third row, the weights of the *curvature* factor are presented. The name of this factor follows from the fact that the weights have a more pronounced effect on the middle range maturities of the curve. These concepts are graphically illustrated below.

```{r, fig.cap = "Yield loading on the spanned factors", echo=FALSE}
LabSpaFac <- c("Level", "Slope", "Curvature")
N <- length(LabSpaFac)

w_pca <- data.frame(t(w[1:N,]))
colnames(w_pca) <- LabSpaFac
w_pca$mat <- mat

# Prepare plots
colors <- c("Level" = "blue", "Slope" = "green", "Curvature" = "red")

g <-  ggplot2::ggplot(data = w_pca, ggplot2::aes(x=  mat)) +  
    ggplot2::geom_line(ggplot2::aes(y = Level, color = "Level"), size = 0.7) + 
    ggplot2::geom_line(ggplot2::aes(y = Slope, color = "Slope"), size = 0.7) +
    ggplot2::geom_line(ggplot2::aes(y = Curvature, color = "Curvature"),  size = 0.7) +
      ggplot2::labs(color = "Legend") + ggplot2::scale_color_manual(values = colors) + ggplot2::theme_classic() +
    ggplot2::theme(axis.title.y= ggplot2::element_blank(), legend.position="top", legend.title=ggplot2::element_blank(), legend.text= ggplot2::element_text(size=8) ) + 
   ggplot2::xlab("Maturity (Years)") + ggplot2::geom_hline(yintercept=0)

print(g)
```

To directly obtain the time series of the country-specific spanned factors, the user can simply use the *Spanned_Factors* function as follows:

```{r}
data('CM_Yields')
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
N <- 2
SpaFact <- Spanned_Factors(Yields, Economies, N)
```

## The P-dynamics estimation
As presented in Table \@ref(tab:ModFea), the dynamics of the risk factors under the $P$-measure evolves according to a VAR(1) model, which may be fully unrestricted (as in the case of the JPS-related models) or somewhat restricted (as in the GVAR and JLL frameworks). Below, the usage of each one of these dimensions are illustrated.

### VAR {#S:SectionVAR}
Using the *VAR* function of this package requires simply selecting the appropriated set of risk factors for the desired estimated model. For instance, for the models *JPS jointP* and *VAR jointQ*, the estimation of the $P$-dynamics parameters is obtained as  
```{r}
data("CM_Factors")
PdynPara <- VAR(RiskFactors, VARtype= "unconstrained")
```
whereas the estimation of a *JPS* model for China is 
```{r}
FactorsChina <- RiskFactors[1:7,]
PdynPara <- VAR(FactorsChina, VARtype= "unconstrained")
```

In both cases, the outputs generated are the vector of intercepts in addition to the feedback and the variance-covariance matrices.

### GVAR {#S:SectionGVAR}
The estimation of a GVAR model requires defining a measure of interdependence among the countries of the economic system. This information is reported in the transition matrix, the entries of which reflect the degree of interconnections of two entities of this same economic system. In this package, the illustration of the transition matrix is based on the average of the cross-borders trade flow weights for the period spanning the years from 2006 to 2019. Note that each row sums up to 1.

```{r}
data("CM_Trade")
t_First <- "2006"
t_Last <-  "2019"
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
type <- "Sample Mean"
W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, DataPath = NULL, TradeFlows)
```

Having defined the form of the transition matrix, one can complete the *GVARinputs* variable along the lines discussed in Section \@ref(S:SectionGVARJLLinputs). 
```{r}
data("CM_Factors_GVAR")

GVARinputs <- list()
GVARinputs$Economies <- Economies
GVARinputs$GVARFactors <- FactorsGVAR
GVARinputs$VARXtype <- "unconstrained"
GVARinputs$Wgvar <- W_gvar

N <- 3

GVARpara <- GVAR(GVARinputs, N)
```

A separate routine is provided for computing the foreign-specific factors (also commonly referred as the star variables) used in the estimation of the VARX models.
```{r}
data('CM_Factors')
StaFac <- StarFactors(RiskFactors, Economies, W_gvar)
```

### JLL {#S:SectionJLL}
Calculating the $P$-dynamics parameters in the form proposed by JLL (2015) requires the following inputs: *(i)* the time series of the risk factors in non-orthogonalized form; *(ii)* the number of country-specific spanned factors and *(iii)* the specification of the *JLLinputs* as presented in Section \@ref(S:SectionGVARJLLinputs). See, for instance:

```{r, eval=FALSE}
data("CM_Factors")
N <- 3
JLLinputs <- list()
ModelType <- "JLL original"  
JLLinputs$Economies <- Economies
JLLinputs$DomUnit <- "China"
JLLinputs$WishSigmas <- 1 
JLLinputs$SigmaNonOrtho <- NULL
JLLinputs$JLLModelType <- ModelType
JLLpara <- JLL(RiskFactors, N, JLLinputs)
```

## ATSM estimation
The estimation of ATSM involves defining the model inputs from the log-likelihood function and, subsequently, the parameters used in its optimization process. The structure proposed in this part of the code is in a great extent based on the term structure package by @LeSingleton2018.

### The log-likelihood function 
For the sake of simplicity, one illustrates the construction of a log-likelihood function based on a *VAR jointQ* model. For both the JLL and GVAR-based models, one must further specify the JLLinputs or GVARinputs along the lines described in the Section \@ref(S:SectionGVARJLLinputs).
```{r}
# Inputs to be specified by the user
data("CM_Yields")
data("CM_Factors")
ModelType <- "VAR jointQ"
Yields <- Yields
ZZ <- RiskFactors
Economies <- c("China", "Brazil", "Mexico", "Uruguay") 
mat <-  Maturities(Yields, Economies, UnitYields = "Month")
DataFrequency <- "Monthly"
GlobalVar <- c("GBC", "CPI_OECD") # Global variables
DomVar <- c("Eco_Act", "Inflation") # Domestic variables 
N <- 3 # Number of country-specific spanned factors

# Generate the "Factor Labels" list (necessary preliminarily step) 
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)

# Prepare the inputs for log-Likelihood function
ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency)

# Log-Likelihood function
f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)
```

### Optimization parameters
JPS (2014) requires the estimation of a set of parameters containing 6 elements, namely: the risk-neutral
long-run mean of the short rate (*r0*), the risk-neutral feedback matrix (*K1XQ*), the variance-covariance matrix (*SSZ*) from the VAR processes, the standard deviation of the errors from the portfolios of yields observed with error (*se*), in addition to the intercept (*K0Z*) and the feedback (*K1Z*) matrices of the physical dynamics. Each one of these parameters must be cast in a individual list that should contain the *(i)* the starting value of the parameter (if any); *(ii)* the variable label followed by a ':' and a type of constraint; *(iii)* a lower bound (if any), and *(iv)* an upper bound (if any). The variable labels of *r0*, *se*, *K0Z* and *K1Z* should be preceded by the symbol @ as manner to account for the fact that the solution for these parameters are known in closed-form. For the remaining ones (*K1XQ* and *SSZ*), @LeSingleton2018 provide standardized routines (already incorporated in this package) to set good initial values so that the optimization process runs fast.

The type of constraint to be set for the parameters *r0*, *se*, *K0Z* and *K1Z* is typically *bounded* and are used for bounded matrices. For *K1XQ* and *SSZ*, the *MultiATSM* package currently provides the following options:

- **K1XQ:** *'Jordan'* (or *Jordan MultiCountry'*) for a matrix of Jordan type for single (multi) country specifications. These labels can be extended by *'; stationary'*, if one wishes to impose the largest eigenvalue of the risk-neutral feedback matrix to be strictly smaller than 1;   
- **SSZ:** *'psd'* for a positive semi-definite matrix; *'diag'* for a diagonal matrix; *'BlockDiag'* for a block diagonal matrix (typical from the GVAR-based models) and *'JLLstructure'* for the models containing the restrictions along the lines of JLL (2015).

One example for the list of parameters specification is

```{r}
K1XQinputs <- list(NULL, "K1XQ: Jordan" , NULL , NULL)
SSZinputs <- list(NULL, "SSZ: psd", NULL, NULL)
r0inputs <- list(NULL, "@r0: bounded", NULL, NULL)
seinputs <- list(NULL, "@se: bounded", 1e-6, NULL)
K0Zinputs<- list(NULL, "@K0Z: bounded", NULL, NULL)
K1Zinputs<- list(NULL, "@K1Z: bounded", NULL, NULL)
```

To complete the optimization setting, the user needs to specify the level of convergence tolerance (usually, 1e-4 is a reasonable value) and the vector *OptRun*. The first element of this vector must be filled by the word *'iter off'*, if the user wishes to switch off the printouts of the numerical optimization routines, or simply by *'iter'* otherwise. The second element of the vector concerns the algorithm that is used in the optimization: the option *'fminunc only'* only uses fminunc, whereas *'fminsearch only'* only applies fminsearch. If *OptRun* is formed exclusively by *'iter off'*, then both fminunc and fminsearch are employed in the optimization.

# Examples of full implementation of ATSMs {#S:SectionReplication}
This section presents some examples on how to use the *MultiATSM* package to fully implement ATSMs. Further on, the functions of this package are used to replicate some of the results presented in the original papers of JPS (2014) and CM (2021).

## General template
```{r, eval=FALSE}
###############################################################################################################
#################################### USER INPUTS ##############################################################
###############################################################################################################
# A) Load database data
data("CM_Factors")
data('CM_Factors_GVAR')
data('CM_Trade')
data('CM_Yields')

# B) Decide on general model inputs
ModelType <- "JLL original" # available options are "JPS", "JPS jointP", "GVAR sepQ", "VAR jointQ", "GVAR jointQ", "JLL original", "JLL NoDomUnit", "JLL jointSigma".

StationarityUnderQ <- 1 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1, NO:0

Economies <- c("China", "Brazil", "Mexico", "Uruguay") # Names of the economies from the economic system
GlobalVar <- c("GBC", "CPI_OECD") # Global Variables
DomVar <- c("Eco_Act", "Inflation") # Country-specific variables

N <- 3 # Number of spanned factors per country

OutputLabel <- "Test" # label of the model for saving the file
DataFrequency <- "Monthly" # Frequency of the data
UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"

# B.1) Decide on specific model inputs
#################################### GVAR-based models #########################################################
if (ModelType == 'GVAR sepQ' || ModelType == 'GVAR jointQ'){
  # Transition matrix inputs:
  t_First <- "2006" # First year of the sample
  t_Last <-  "2019" # Last year of the sample
  W_type <- 'Sample Mean' # Method to compute the transition matrix
  # VARX input:
  VARXtype <- "unconstrained"
}

#################################### JLL-based models #########################################################
if (ModelType =="JLL original" || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma"){
  DominantCountry <- "China" # (i) one of the countries of the system or (ii) "None"
}
################################################################################################################

# C) Decide on Settings for numerical outputs
Horiz <- 50
DesiredGraphs <- c("IRF", "FEVD") #c("Fit", "IRF", "FEVD", "GIRF", "GFEVD")
WishGraphRiskFac <- 0
WishGraphYields <- 1
WishOrthoJLLgraphs <- 0

# D) Bootstrao settings
WishBootstrap <- 1 #  YES: 1; No = 0.
Bootlist <- list()
Bootlist$methodBS <- 'bs' # (i) 'bs' ; (ii) 'wild'; (iii) 'block'
Bootlist$BlockLength <- 4 # necessary input if one chooses the block bootstrap
Bootlist$ndraws <-  30
Bootlist$pctg   <-  95 # confidence level

# E) Out-of-sample forecast
WishForecast <- 1 #  YES: 1; No = 0.
ForecastList <- list()
ForecastList$ForHoriz <- 12 # forecast horizon
ForecastList$t0Sample <- 1 # initial sample date
ForecastList$t0Forecast <- 145 # last sample date for the first forecast

###############################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #############################################
###############################################################################################################

# 2) Minor preliminary work
C <- length(Economies)
FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) # Generate the set of labels
mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields) # Vector of maturities expressed in years
# (alternatively, it can be completed manually)
ZZ <- RiskFactors
# 2.1) Build the GVARinputs
if (ModelType == 'GVAR sepQ' || ModelType == 'GVAR jointQ'){
  GVARinputs <- list()
  GVARinputs$Economies <- Economies
  GVARinputs$GVARFactors <- FactorsGVAR
  GVARinputs$VARXtype <- VARXtype
  GVARinputs$Wgvar <- Transition_Matrix(t_First, t_Last, Economies, W_type, DataPath = NULL, Data = TradeFlows)
} else { GVARinputs <- NULL }

# 2.2) Build the JLLinputs
if (ModelType =="JLL original" || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma"){
  JLLinputs <- list()
  JLLinputs$Economies <- Economies
  JLLinputs$DomUnit <- DominantCountry
  JLLinputs$WishSigmas <- 1 # Sigma matrix is estimated within the "InputsForMLEdensity" function
  JLLinputs$SigmaNonOrtho <- NULL
  JLLinputs$JLLModelType <- ModelType
}else{
  JLLinputs <- NULL
}

# 3) Prepare the inputs of the likelihood function
ModelParaList <- list()
for (i in 1:C){
  if (( ModelType == "GVAR jointQ" || ModelType == "VAR jointQ" || ModelType == "JLL original"
        || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma" )   & i >1 ){break} # the break avoids the models with joint estimation under the Q to be estimated C times

  # 3.1) Compute the inputs that go directly into the log-likelihood function
  ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, JLLinputs,   GVARinputs)

  # 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
  K1XQ <- ATSMInputs$K1XQ
  if (ModelType == "JLL original" || ModelType == "JLL NoDomUnit" ){ SSZ <- NULL} else{SSZ <- ATSMInputs$SSZ}

  # 4) Build the objective function
  f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)

  # 5) Set the optimization settings
  VarLab <- ParaLabels(ModelType, StationarityUnderQ)

  varargin <- list()
  varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
  varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
  varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
  varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
  varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
  varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
  varargin$OptRun <- c("iter off")

  LabelVar<- c('Value', 'Label', 'LB', 'UB') # Elements of each parameter
  for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}

  tol <- 1e-4

  # 6) Optimization of the model
  if (ModelType == 'JPS' || ModelType == 'JPS jointP' || ModelType == "GVAR sepQ"){
    ModelParaList[[ModelType]][[Economies[i]]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType)$Summary
  }else{ ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs)$Summary}

}

# 7) Numerical and graphical outputs
InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,                  WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishBootstrap, Bootlist, WishForecast, ForecastList)
# A) Fit, IRF, FEVD, GIRF and GFEVD
NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)

# B) Bootstrap
Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs, FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs)

# C) Out-of-sample forecasting
Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies, DataFrequency, JLLinputs, GVARinputs)

```

## Paper replications
### Joslin, Priebisch and Singleton (2014)
To the best of my knowledge, the original data-set used in @JoslinPriebschSingleton2014 is not publicly available. Nevertheless, there exists an alternative source of data that closely matches JPS (2014) in terms of the construction method and the sample period. This database is constructed by @BauerRudebusch2017 (BR, 2017) and is available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/). In that paper, BR (2017) investigate whether macro-finance term structure models better suit the unspanned macro risk framework by JPS (2014) or the earlier traditional spanned settings as the one by @AngPiazzesi2003JME. Accordingly, BR (2017) intend to replicate some of the empirical results reported in JPS (2014). The R-code used by BR (2017) is also available at available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).

Supported by BR (2017)'s data-set, the code below uses the *MultiATSM* package to estimate the key model parameters from the ATSM along the lines of JPS (2014).

```{r, eval=FALSE} 
###############################################################################################################
#################################### USER INPUTS ##############################################################
###############################################################################################################
# A) Load database data
data("BR_jps_gro_R3")

# B) Decide on general model inputs
ModelType <- "JPS"

StationarityUnderQ <- 0 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1, NO:0

Economies <- c("U.S.") # Names of the economies from the economic system
GlobalVar <- c()
DomVar <- c("GRO", "INF") # Country-specific variables

N <- 3 # Number of spanned factors per country

DataFrequency <- "Monthly" # Frequency of the data
UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"

mat <- c(0.25, 0.5, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10) # maturities expressed in years

###############################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #############################################
###############################################################################################################
# 2) Minor preliminary work
C <- length(Economies)
M <- length(DomVar)
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType) # Generate the set of labels
Yields <- t(BR_jps_out$Y)
W <- BR_jps_out$W[1:N,] # Use the Wpca matrix from BR (2017)
SpaFac <- W%*%Yields
rownames(SpaFac) <- FactorLabels$Tables$U.S.[-(1:M)]
ZZ <- rbind(t(BR_jps_out$M.o), SpaFac) # Complete set of risk factors

# 3) Prepare the inputs of the likelihood function
for (i in 1:C){
# 3.1) Compute the inputs that go directly into the log-likelihood function
ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency)
# 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
K1XQ <- ATSMInputs$K1XQ
SSZ <- ATSMInputs$SSZ
# 3.3) Adjust the inputs which are funtion of the W matrix
ATSMInputs$Wpca <- W
ATSMInputs$We <- t(pracma::null(W))
ATSMInputs$WpcaFull <- rbind(ATSMInputs$Wpca, ATSMInputs$We)
ATSMInputs$PP <- SpaFac

# 4) Build the objective function
f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)

# 5) Set the optimization settings
VarLab <- ParaLabels(ModelType, StationarityUnderQ)

varargin <- list()
varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
varargin$OptRun <-  c("iter off")

LabelVar<- c('Value', 'Label', 'LB', 'UB') # Elements of each parameter
for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}

tol <- 1e-4

# 6) Optimization of the model
ModelPara <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType)$Summary
}
```

The tables below compare the ATSM parameter estimates generated from BR (2017) and the *MultiATSM*. Overall, one can note that the differences in the estimates are economically modest.

```{r, echo= FALSE}
options(scipen = 100) # eliminate the scientific notation
data("BR_jps_gro_R3")
data("JPSrep")

RowsQ <- c("$r0$", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_3$" )
TableQ <- data.frame(matrix(NA, ncol = 0, nrow =length(RowsQ)))
row.names(TableQ) <- RowsQ

PackageQ<- c(ModelPara$ests$r0, diag(ModelPara$ests$K1XQ))
BRq <- c(BR_jps_out$est.llk$rho0.cP, diag(BR_jps_out$est.llk$KQ.XX))
TableQ$MultiATSM <- PackageQ
TableQ$'BR (2017)' <- BRq

TableQ <- round(TableQ, digits = 5)

kableExtra::kbl(TableQ, align = "c", caption = "$Q$-dynamics parameters") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::footnote(general = " $\\lambda$'s are the eigenvalues from the risk-neutral feedback matrix and $r0$ is the long-run mean of the short rate under Q.")
```

```{r, echo= FALSE}
data("BR_jps_gro_R3")
data("JPSrep")

RowsP <- c("PC1", "PC2", "PC3", "GRO", "INF")
ColP <- c(" ", RowsP)

# 1) K0Z and K1Z
# Bauer and Rudebusch coefficients
TablePbr <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
row.names(TablePbr) <- RowsP
colnames(TablePbr) <- ColP

TablePbr[[ColP[1]]] <- BR_jps_out$est.llk$KP.0Z
for(j in 1:length(RowsP) ){TablePbr[[RowsP[j]]] <- BR_jps_out$est.llk$KP.ZZ[,j]}

TablePbr <- round(TablePbr, digits = 5)

# MultiATSM coefficients
TablePMultiATSM <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
row.names(TablePMultiATSM) <- RowsP
colnames(TablePMultiATSM) <- ColP

IdxVar <- c(3:5, 1:2) # indexes to flip the order of the spanned and unspanned factors
TablePMultiATSM[[ColP[1]]] <- ModelPara$ests$K0Z[IdxVar]
ModelPara$ests$K1Z <- ModelPara$ests$K1Z[IdxVar,IdxVar]
for(j in 1:length(RowsP) ){ TablePMultiATSM[[RowsP[j]]] <- ModelPara$ests$K1Z[,j]}

TablePMultiATSM <- round(TablePMultiATSM, digits = 5)
TableP <- rbind(TablePbr,TablePMultiATSM)
row.names(TableP) <- c(RowsP,paste(RowsP," ",sep="")) # trick to avoid rows to have the same name

kableExtra::kbl(TableP, align = "c", caption = "$P$-dynamics parameters") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::add_header_above(c(" "= 1, "K0Z"=1, "K1Z" = 5), bold = T) %>%
  kableExtra::pack_rows("BR (2017)", 1, 5) %>%
  kableExtra::pack_rows("MultiATSM", 6, 10) %>%
  kableExtra::footnote(general = " $K0Z$ is the intercept and $K1Z$ is feedback matrix from the $P$-dynamics.")
```

```{r, echo= FALSE}
data("BR_jps_gro_R3")
data("JPSrep")

se <- data.frame(BR_jps_out$est.llk$sigma.e, ModelPara$ests$se )
rownames(se) <- "se"
colnames(se) <- c("MultiATSM","BR (2017)")
se <- round(se, digits = 7)

kableExtra::kbl(se, align = "c", caption ="Portfolio of yields with errors") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::footnote(general = " $se$ is the standard deviation of the portfolio of yields observed with errors.")
```

It is relevant to highlight that the script above makes use of the principal component weight provided by BR (2017). Such a matrix is simply a scaled-up version of the one provided by the function *pca_weights_one_country* of this package. Accordingly, despite the numerical differences on the weight matrices, both methods generate time series of spanned factors which are perfectly correlated. Another difference between the two approaches relates to the construction form of the log-likelihood function (llk): while in the BR (2017) code the llk is expressed in terms of portfolio of yields, the *MultiATSM* package generates this same input directly as a function of observed yields (i.e. both procedures lead to equivalent llk up to the Jacobian term).

### Candelon and Moura (2021)
```{r, eval=FALSE}
###############################################################################################################
#################################### USER INPUTS ##############################################################
###############################################################################################################
# A) Load database data
data("CM_Factors")
data('CM_Factors_GVAR')
data('CM_Trade')
data('CM_Yields')

# B) Decide on general model inputs
ModelType <- "GVAR jointQ" # Set "GVAR jointQ" for the GVAR-ATSM and "JLL original" for the JLL-ATSM

StationarityUnderQ <- 0

Economies <- c("China", "Brazil", "Mexico", "Uruguay") 
GlobalVar <- c("GBC", "CPI_OECD") # Global Variables
DomVar <- c("Eco_Act", "Inflation") # Country-specific variables

N <- 3 # Number of spanned factors per country

OutputLabel <- "CM_2021" # label of the model for saving the file
DataFrequency <- "Monthly" # Frequency of the data
UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"

# B.1) Decide on specific model inputs
#################################### GVAR-based models #########################################################
if (ModelType == 'GVAR jointQ'){
# Transition matrix inputs:
t_First <- "2006" # First year of the sample
t_Last <-  "2019" # Last year of the sample
W_type <- 'Sample Mean' # Method to compute the transition matrix
# VARX input:
VARXtype <- "unconstrained"
}

#################################### JLL-based models #########################################################
if (ModelType =="JLL original"){
    DominantCountry <- "China"
}
################################################################################################################

# C) Decide on Settings for numerical outputs
Horiz <- 25
DesiredGraphs <- c("Fit", "GIRF", "GFEVD")
WishGraphRiskFac <- 0
WishGraphYields <- 1
WishOrthoJLLgraphs <- 1

# D) Bootstrao settings
WishBootstrap <- 0 #  If bootstrap outputs are desired set this variable equal to 1 (it may take several hours/days).
Bootlist <- list()
Bootlist$methodBS <- 'bs'
Bootlist$BlockLength <- c()
Bootlist$ndraws <-  1000
Bootlist$pctg   <-  95

# E) Out-of-sample forecast
WishForecast <- 0 #  If out-of-sample forecasts outputs are desired set this variable equal to 1
ForecastList <- list()
ForecastList$ForHoriz <- 12
ForecastList$t0Sample <- 1
ForecastList$t0Forecast <- 90

###############################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #############################################
###############################################################################################################

# 2) Minor preliminary work
C <- length(Economies)
FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) # Generate the set of labels
mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields) # Vector of maturities expressed in years

ZZ <- RiskFactors
# 2.1) Build the GVARinputs
if (ModelType == "GVAR jointQ"){
  GVARinputs <- list()
  GVARinputs$Economies <- Economies
  GVARinputs$GVARFactors <- FactorsGVAR
  GVARinputs$VARXtype <- VARXtype
  GVARinputs$Wgvar <- Transition_Matrix(t_First, t_Last, Economies, W_type, DataPath = NULL, Data = TradeFlows)
  } else { GVARinputs <- NULL }

# 2.2) Build the JLLinputs
if (ModelType =="JLL original"){
  JLLinputs <- list()
  JLLinputs$Economies <- Economies
  JLLinputs$DomUnit <- DominantCountry
  JLLinputs$WishSigmas <- 1
  JLLinputs$SigmaNonOrtho <- NULL
  JLLinputs$JLLModelType <- ModelType
}else{  JLLinputs <- NULL }

# 3) Prepare the inputs of the likelihood function
ModelParaList <- list()

# 3.1) Compute the inputs that go directly into the log-likelihood function
ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, JLLinputs,
GVARinputs)

  # 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
K1XQ <- ATSMInputs$K1XQ
if (ModelType == "JLL original"){ SSZ <- NULL} else{SSZ <- ATSMInputs$SSZ}

# 4) Build the objective function
f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)

# 5) Set the optimization settings
VarLab <- ParaLabels(ModelType, StationarityUnderQ)

varargin <- list()
varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
varargin$OptRun <-  c("iter off")

LabelVar<- c('Value', 'Label', 'LB', 'UB')
for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}

tol <- 1e-4

# 6) Optimization of the model
ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs)$Summary

# 7) Numerical and graphical outputs
InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ, WishGraphYields,  WishGraphRiskFac, WishOrthoJLLgraphs, WishBootstrap, Bootlist, WishForecast, ForecastList)

# A) Fit, IRF, FEVD, GIRF and GFEVD
NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)

# B) Bootstrap
Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs, FactorLabels,
DataFrequency, varargin, JLLinputs, GVARinputs)

# C) Out-of-sample forecasting
Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies, DataFrequency, JLLinputs, GVARinputs)
```

# References

### 2021-11-02 R/A0N__computeBnAn.R

@@ -1,0 +1,172 @@
+#' Compute the cross-section loadings of yields of a canonical A0_N model ("sep Q" models)
+#'
+#'@param mat       vector of maturities (J x 1). Maturities are in multiples of the discrete interval
+#'                 used in the model
+#'@param K1XQ      risk neutral feedback matrix (N x N)
+#'@param dx        state loadings for the one-period rate (1xN). Default is a vector of ones
+#'@param r0        the long run risk neutral mean of the short rate (scalar)
+#'@param SSX       the covariance matrix of the errors (N x N)
+#'
+#'@return
+#' List containing:
+#' \itemize{
+#' \item Intercept (Jx1)
+#' \item slope (JxN)
+#' \item the betan (JX1, part of the intercepts unrelated to the long run risk neutral mean r0)
+#' coefficients of a canonical A_0(N).
+#' }
+#'@references
+#'  \itemize{
+#'  \item This function is based on the "A0N__computeBnAn" function by Le and Singleton (2018). \cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+#'
+#'  \item Dai and Singleton (2000). "Specification Analysis of Affine Term Structure Models" (The Journal of Finance)
+#'  }
+
+
+
+A0N__computeBnAn_sepQ <- function(mat, K1XQ, dx, r0, SSX){
+
+
+  mat <- round(mat[])
+  K <- max(mat) # longest maturity of interest
+  N <- dim(K1XQ)[1] # Number of latent factors
+
+
+  if (!exists("dX", inherits = FALSE) || is.null(dX) ){
+    dX <- rep(1, times=N)
+  }
+  dX <- as.vector(dX)
+
+  BnX <- matrix(0, nrow = K+1, ncol= N)
+  AnX <- matrix(0, nrow = K+1, ncol=1)
+
+  # Generate the loadings from a canonical A_0(N) model:
+  for ( k in 1:K){
+    BnX[k+1,] <- dX + BnX[k,]%*%K1XQ # Note: BnX depends only on K1XQ and dx NOT r0 or SSX.
+    if (!exists("r0", inherits = FALSE) || is.null(r0) ){
+      AnX <- as.numeric(vector())
+    } else {
[…]

### 2021-11-02 R/Bootstrap.R

@@ -1,0 +1,346 @@
+#' Generates the bootstrap-related outputs
+#'
+#'@param ModelType   string-vector containing the label of the model to be estimated
+#'@param ModelParaPE point estimate from the model parameters (see the outputs of the "Optimization" function)
+#'@param NumOutPE    point estimate from the numerical outputs (see the outputs of the "NumOutputs" function)
+#'@param mat         vector of maturities (in years) used in the estimation
+#'@param Economies   string-vector containing the names of the economies which are part of the economic system
+#'@param InputsForOutputs list containing the desired inputs for the construction of IRFs, GIRFs, FEVDs, and GFEVDs.
+#'@param FactorLabels   string-list based which contains the labels of all the variables present in the model
+#'@param DataFrequency    character-based vector: "Daily All Days", "Daily Business Days", "Weekly", "Monthly", "Quarterly", "Annually"
+#'@param vararginPE    list containg starting values and constraints (see arguments of the "Optimization" function)
+#'@param JLLinputs    list of necessary inputs for the estimation of JLL-based models (see "JLL" function)
+#'@param GVARinputs list of necessary inputs for the estimation of GVAR-based models (see "GVAR" function)
+#'
+#'
+#'@importFrom pracma ceil rand strcmp num2str tril
+#'
+#'
+#'
+#'@examples
+#' # See examples in the vignette file of this package (Section 4).
+#'
+#'@returns
+#'list containing the following elements:
+#'\itemize{
+#' \item list of model parameters for one each one the draws;
+#' \item list of numerical outputs (IRFs, GIRFs, FEVDs, GFEVDs) for each one of the draws;
+#' \item Confidence bands for the chosen level of significance.
+#' }
+
+#'@references
+#' This function is a modified and extended version of the "VARirbound" function from "A toolbox for VAR analysis"
+#' by Ambrogio Cesa-Bianchi (https://github.com/ambropo/VAR-Toolbox)
+#' @export
+
+
+Bootstrap <- function(ModelType, ModelParaPE, NumOutPE, mat, Economies, InputsForOutputs,
+                      FactorLabels, DataFrequency, vararginPE, JLLinputs = NULL, GVARinputs = NULL){
+
+
[…]

### 2021-11-02 R/CM_Factors.R

@@ -1,0 +1,10 @@
+#'@title Data: Risk Factors - Candelon and Moura (2021)
+#'
+#' @description Risk factors data used in Candelon and Moura (2021)
+#' @name RiskFactors
+#' @docType data
+#' @usage data("CM_Factors")
+#' @format matrix containing the  risk factors of the models
+#' @references Candelon, B. and Moura, R. "A Multi-Country Model of the Term Structures of Interest Rates with a GVAR".
+#' @keywords Risk Factors
+NULL

### 2021-11-02 R/CM_Trade.R

@@ -1,0 +1,10 @@
+#'@title Data: Trade Flows - Candelon and Moura (2021)
+#'
+#' @description Trade Flows data used in Candelon and Moura (2021)
+#' @name TradeFlows
+#' @docType data
+#' @usage data("CM_Trade")
+#' @format list containing the bilateral trade flows
+#' @references Candelon, B. and Moura, R. "A Multi-Country Model of the Term Structures of Interest Rates with a GVAR".
+#' @keywords Trade Flows
+NULL

### 2021-11-02 R/CM_Yields.R

@@ -1,0 +1,10 @@
+#'@title Data: Yields - Candelon and Moura (2021)
+#'
+#' @description Yields data used in Candelon and Moura (2021)
+#' @name Yields
+#' @docType data
+#' @usage data("CM_Yields")
+#' @format matrix containing the Yields of the models
+#' @references Candelon, B. and Moura, R. "A Multi-Country Model of the Term Structures of Interest Rates with a GVAR".
+#' @keywords Yields
+NULL

### 2021-11-02 R/ConfidenceBoundsBootstrap.R

@@ -1,0 +1,1464 @@
+#' Builds the confidence bounds and graphs (Bootstrap set)
+#'
+#'@param ModelType string-vector containing the label of the model to be estimated
+#'@param ModelBootstrap list containing the complete set of model parameters after the bootstrap estimation procedure
+#'@param NumOutPE      point estimate from the numerical outputs (see the outputs of the "NumOutputs" function)
+#'@param InputsForOutputs list conataining the desired inputs for the construction of IRFs, GIRFs, FEVDs, and GFEVDs
+#'@param Economies string-vector containing the names of the economies which are part of the economic system
+#'
+
+
+BootstrapBoundsSet <- function(ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies){
+
+  # Generate the graph paths and the graph folders
+  dir.create(paste(tempdir(), "/Outputs/", ModelType, "/Bootstrap", sep=""))
+  PathsGraphs <- FolderCreationBoot(ModelType, Economies)
+
+
+  # If one chooseS models in which the estimation is done country-by-country
+  if ( "JPS" %in% ModelType || "JPS jointP" %in% ModelType ||  "GVAR sepQ" %in% ModelType){
+
+    for (i in 1:length(Economies)){dir.create( paste(tempdir(), "/Outputs/", ModelType, "/Bootstrap/Model ", Economies[i], sep=""))}
+
+
+    IRFandGIRFsep <- IRFandGIRFbs_sepQ(ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies, PathsGraphs)
+    FEVDandGFEVDsep <- FEVDandGFEVDbs_sepQ(ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies, PathsGraphs)
+
+
+    NumOutSep <- append(IRFandGIRFsep, FEVDandGFEVDsep)
+  }
+
+  # If one chooseS models in which the estimation is done jointly for all countries
+  if ( "GVAR jointQ" %in% ModelType || "VAR jointQ" %in% ModelType || "JLL original" %in% ModelType
+       || "JLL NoDomUnit" %in% ModelType || "JLL jointSigma" %in% ModelType){
+
+
+    IRFandGIRFjoint <- IRFandGIRFbs_jointQ(ModelType, ModelBootstrap, NumOutPE, InputsForOutputs,
+                                           Economies, PathsGraphs)
[…]

### 2021-11-02 R/DataForEstimation.R

@@ -1,0 +1,72 @@
+#' Retrieve data from Excel and build the database used in the model estimation
+#'
+#' @param t0 Sample starting date (yyyy-mm-dd)
+#' @param tF Sample last date (yyyy-mm-dd)
+#' @param Economies string-vector containing the names of the economies which are part of the economic system
+#' @param N   Number of country-specific spanned factor (scalar)
+#' @param FactorLabels String-list based which contains the labels of all the variables present in the model
+#' @param ModelType   String-vector containing the label of the model to be estimated
+#' @param DataFrequency  Character-based-vector. Avaialable options are: "Daily All Days", "Daily Business Days", "Weekly", "Monthly", "Quarterly", "Annually"
+#' @param DataPathMacro  Path of the Excel file conating the macroeconomic data (if any). The default is linked to the excel file present in the package.
+#' @param DataPathYields  Path of the Excel file conating the yields data (if any). The default is linked to the excel file present in the package.
+#' @param Wgvar   GVAR transition matrix, if GVAR type model is chosen; default is set to NULL.
+#'
+#'
+#'@return   A list containing the
+#'\enumerate{
+#'\item time series of the complete set of bond yields (matrix, JxT or CJxT);
+#'\item time series of the complete set risk factors (matrix, KxT);
+#'\item 'GVARFactors': list of all variables that are used in the estimation of the VARX \cr
+#'                (see e.g. 'CM_Factors_GVAR' file). If the estimated model type is not GVAR-based, then returns NULL.
+#'}
+#'
+#'@examples
+#'DomVar <- c("Eco_Act", "Inflation")
+#'GlobalVar <- c("GBC", "CPI_OECD")
+#' t0 <- "2006-09-01"
+#' tF <-  "2019-01-01"
+#' Economies <- c("China", "Brazil", "Mexico", "Uruguay", "Russia")
+#' N <- 2
+#' ModelType <- "JPS"
+#' FactorLabels <-  LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+#' DataFrequency <- "Monthly"
+#'
+#'
+#' DataModel <- DataForEstimation(t0, tF, Economies, N, FactorLabels, ModelType, DataFrequency)
+#'
+#' @export
+
[…]

### 2021-11-02 R/DatabasePrep.R

@@ -1,0 +1,253 @@
+#' Prepare the GVARFactors database
+
+
+#'@param t_First             sample starting date (yyyy-mm-dd)
+#'@param t_Last              sample last date (yyyy-mm-dd)
+#'@param Economies           string-vector containing the names of the economies which are part of the economic system
+#'@param N                   number of country-specific spanned factor (scalar)
+#'@param FactorLabels        list containing the factor labels
+#'@param ModelType           string-vector containing the label of the model to be estimated
+#'@param Wgvar               GVAR transition matrix (CxC), if GVAR type model is chosen; default is set to NULL.
+#'@param DataPathMacro       path of the Excel file containing the macroeconomic data (if any). The default is linked to the Excel file available in the package.
+#'@param DataPathYields      path of the Excel file containing the yields data (if any). The default is linked to the Excel file available in the package.
+#'
+#'@return List of the risk factor set used in the estimation of the GVAR model
+#'
+#'
+#'@importFrom pracma isempty
+#'
+#'
+#'@examples
+#' DomVar <- c("Eco_Act", "Inflation")
+#' GlobalVar <- c("GBC", "CPI_OECD")
+#' t0 <- "2006-09-01"
+#' tF <-  "2019-01-01"
+#' Economies <- c("China", "Brazil", "Mexico", "Uruguay", "Russia")
+#' N <- 3
+#' ModelType <- "JPS jointQ"
+#' FactorLabels <-  LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+#'
+#' GVARFactors <- DatabasePrep(t0, tF, Economies, N, FactorLabels, ModelType)
+#'
+#'
+#'@returns
+#' List containing the risk factor set used in the estimation of the GVAR-based models
+#'
+#'@export
+
+
+
+DatabasePrep <- function(t_First, t_Last, Economies, N, FactorLabels, ModelType, Wgvar = NULL,
+                         DataPathMacro = NULL, DataPathYields = NULL){
+
+
+  if (length(DataPathMacro) ==0 ) { DataPathMacro <- system.file("extdata", "MacroData.xlsx", package = "MultiATSM") }
[…]

### 2021-11-02 R/EstimationSigma_Ye.R

@@ -1,0 +1,46 @@
+#' Estimate numerically the Cholesky-factorization from the JLL-based models
+#'
+
+#'@param SigmaUnres   unrestricted variance-covariance matrix (K x K)
+#'@param res          residuals from the VAR of the JLL model (K x T)
+#'@param M          number of domestic unspanned factors per country (scalar)
+#'@param G          number of global unspanned factors (scalar)
+#'@param Economies  string-vector containing the names of the economies which are part of the economic system
+#'@param DomUnit Name of the economy which is assigned as the dominant unit. \cr
+#'               If no dominant unit is assigned, then this variable is defined as "none"
+#'
+#'
+#'@keywords internal
+#'@return    Cholesky-factorization after the maximization (K x K)
+#'
+
+
+EstimationSigma_Ye <- function(SigmaUnres, res, M, G, Economies, DomUnit){
+
+  # SIGMA_Ye
+  K <- nrow(SigmaUnres)
+  C <- length(Economies)
+  N <- (K - G - M*C)/C
+
+  # Set the constraints in the Sigma matrix
+  Se <- CholRestrictionsJLL(SigmaUnres, M, G, N, Economies, DomUnit)
+  IdxNONzeroSigmaJLL <- which(Se!=0)
+  x <- Se[IdxNONzeroSigmaJLL] # vector containing the initial guesses
+
+  MLfunction <- functional::Curry(llk_JLL_Sigma, res=res, IdxNONzero= IdxNONzeroSigmaJLL, K=K)
+
+  iter <- 'off' # hides the outputs of each iteration. If one wants to display these features then set 'iter'
+  options200 <- neldermead::optimset(MaxFunEvals = 200000*numel(x), Display = iter,
+                         MaxIter = 200000, GradObj='off', TolFun= 10^-2, TolX= 10^-2)
+
+
+  Xmax <- neldermead::fminsearch(MLfunction, x, options200)$optbase$xopt
+  SIGMA_Ye <- matrix(0, K,K)
+  SIGMA_Ye[IdxNONzeroSigmaJLL]<- Xmax # Cholesky term (orthogonalized factors)
+
+  #Labels
+  rownames(SIGMA_Ye) <- rownames(SigmaUnres)
+  colnames(SIGMA_Ye) <- rownames(SigmaUnres)
+
+  return(SIGMA_Ye)
+}

### 2021-11-02 R/Folder_Creation.R

@@ -1,0 +1,144 @@
+#' Creates the folders and the path in which the graphical outputs are stored (ponit estimate version)
+#'
+#'@param ModelType a string-vector containing the label of the model to be estimated
+#'@param Economies a string-vector containing the names of the economies which are part of the economic system
+#'
+#'@keywords internal
+#'
+
+
+FolderCreationPoint <- function(ModelType, Economies){
+
+C <- length(Economies)
+
+PathsGraphs <- list()
+
+OutputTypeSet <- c("Fit", "IRF", "FEVD", "GIRF", "GFEVD")
+VarTypeSet <- c("Factors", "Yields")
+
+
+  for (h in 1:length(OutputTypeSet)){
+    PathsGraphs[[ModelType]][[OutputTypeSet[h]]] <- list()
+  }
+
+
+if (ModelType == "JPS" || ModelType == "JPS jointP" || ModelType == "GVAR sepQ"){
+for (h in 1:length(OutputTypeSet)){
+  for (v in 1:length(VarTypeSet)){
+  for (i in 1:C){
+    if (h==1){
+    PathsGraphs[[ModelType]][[OutputTypeSet[h]]][[Economies[i]]]  <- paste(tempdir(), "/Outputs/", ModelType,
+                                                                           "/Point Estimate/Model ", Economies[i], "/",
+                                                                           OutputTypeSet[h], sep="")
+    } else{
+      PathsGraphs[[ModelType]][[OutputTypeSet[h]]][[Economies[i]]][[VarTypeSet[v]]]  <- paste(tempdir(), "/Outputs/",
+                                                                                              ModelType, "/Point Estimate/Model ",
+                                                                                              Economies[i], "/", OutputTypeSet[h], "/",
+                                                                                              VarTypeSet[v], sep="")
+       }
+    }
+}
+}
+} else{
+
+  for (h in 1:length(OutputTypeSet)){
+    for (v in 1:length(VarTypeSet)){
+        if (h==1){
+          PathsGraphs[[ModelType]][[OutputTypeSet[h]]]  <- paste(tempdir(),"/Outputs/", ModelType, "/Point Estimate/",
+                                                                  OutputTypeSet[h], sep="")
+        } else{
+          PathsGraphs[[ModelType]][[OutputTypeSet[h]]][[VarTypeSet[v]]]  <- paste(tempdir(),"/Outputs/", ModelType,
[…]

### 2021-11-02 R/ForecastYields.R

@@ -1,0 +1,580 @@
+#' Gather bond yields forecasts for all the model types
+#'
+#'@param ModelType a string-vector containing the label of the model to be estimated
+#'@param ModelPara  List of model parameter estimates (See the "Optimization" function)
+#'@param InputsForOutputs list conataining the desired horizon of analysis for the IRFs, GIRFs, FEVDs, and GFEVDs
+#'@param FactorLabels  a string-list based which contains all the labels of all the variables present in the model
+#'@param Economies string-vector containing the names of the economies which are part of the economic system
+#'@param DataFrequency text: "Daily All Days", "Daily Business Days", "Weekly", "Monthly", "Quarterly", "Annually"
+#'@param JLLinputs list of necessary inputs for the estimation of JLL-based models (see "JLL" function)
+#'@param GVARinputs list of necessary inputs for the estimation of GVAR-based models (see "GVAR" function)
+#'
+
+#'@examples
+#' # See examples in the vignette file of this package (Section 4).
+#'
+#'
+#'@returns
+#' List containg the following elements
+#' \enumerate{
+#' \item Out-of-sample forecasts of bond yields per forecast horizon
+#' \item Out-of-sample forecast errors of bond yields per forecast horizon
+#' \item Root mean square errors per forecast horizon
+#'}
+#'
+#'@export
+
+ForecastYields <- function(ModelType, ModelPara, InputsForOutputs, FactorLabels, Economies, DataFrequency,
+                               JLLinputs, GVARinputs){
+
+
+  WishForecast <- InputsForOutputs[[ModelType]]$Forecasting$WishForecast
+
+  if (WishForecast ==0){ print( "No bond yields forecasts were generated")
+    }else{
+      Jmisc::tic()
+  # If one chooseS models in which the estimation is done country-by-country
+  if ( "JPS" %in% ModelType || "JPS jointP" %in% ModelType ||  "GVAR sepQ" %in% ModelType){
+    ForeYie <- ForecastYieldsSepQ(ModelType, ModelPara, InputsForOutputs, FactorLabels, Economies, DataFrequency,
+                                  JLLinputs, GVARinputs)
+
+  }
+  # If one chooseS models in which the estimation is done jointly for all countries
[…]

### 2021-11-02 R/GVAR.R

@@ -1,0 +1,408 @@
+#' Estimate a GVAR(1) and a VARX(1,1,1)
+#'
+#
+#'@param GVARinputs List containing the following necessary inputs for the estimation of the GVAR:
+#' \enumerate{
+#'        \item  Economies:  string-vector containing the names of the economies which are part of the economic system
+#'        \item  'GVARFactors': list of all variables that are used in the estimation of the VARX \cr
+#'                (see e.g. 'CM_Factors_GVAR' file);
+#'        \item 'VARXtype': character-vector containing two possibilities:
+#'  \itemize{
+#'        \item 'unconstrained': model is estimated without any constrained (each equation is estimated individually by OLS);
+#'        \item "constrained": model is estimated taking into account the fact that foreign-pricing-factors
+#'                              do NOT impirge on
+#'                               (i) domestic economic variables and (ii) domestic pricing factors.
+#'                               (equations are estimated by restricted least squares)
+#'          }
+#'          \item 'Wgvar':  GVAR transition matrix (C x C) - see the output from 'Transition_Matrix' function
+#' }
+#'@param N    number of country-specific spanned factors (scalar)
+#'
+#'@importFrom pracma numel
+#'
+#'@return   A list containing
+#'\enumerate{
+#'\item parameters of the country-specific VARX(1,1,1)
+#'\itemize{
+#'\item intercept (M+Nx1);
+#'\item phi_1   (M+N x M+N);
+#'\item phi_1^star (M+N x M+N);
+#'\item phi_g (M+N x M+N);
+#'\item Sigma (M+N x G)
+#'}
+#'\item parameters of the GVAR.
+#'\itemize{
+#' \item F0 (F X 1);
+#' \item F1 (F x F);
+#' \item Sigma_y (F x F)
+#'}
+#'}
+#'
+#'
+#'@examples
[…]

### 2021-11-02 R/GaussianDensity.R

@@ -1,0 +1,59 @@
+#' computes the density function of a gaussian process
+
+#'@param res    matrix of residuals (N x T)
+#'@param SS     covariance matrice or array of covariance matrices \cr
+#'               (If dim(SS) > 3, then the model has a stochastic volatility) (N x N) or (N x N x T)
+#'
+#'@param invSS  Inverse of SS (N x N) or (N x N x T) - optional input
+#'@param logabsdetSS   log(abs(|SS|)) (1 x T) - optional input
+#'
+#'@return y   - vector of density (1 x T)
+#'
+#'@references
+#' This function is based on the "Gaussian" function by Le and Singleton (2018).\cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+#'
+
+
+
+GaussianDensity <- function(res,SS, invSS, logabsdetSS){
+
+
+  nargin <- nargs()
+
+
+  N <- dim(res)[1]
+  T <- dim(res)[2]
+
+
+  if (is.na(dim(SS)[3])){ # If SS is a matrix and NOT a 3-dimension array.
+    if (!missing("invSS")){
+      SSres <-  invSS%*%res
+    }else{
+      SSres <- solve(SS,res)} # SSres <- (SS)^(-1)*res
+
+
+    if (missing("logabsdetSS")){
+      logabsdetSS <- 0.5*logdet(SS%*%t(SS),opt="chol")} # why do we multiply by 0.5? Because in "logdet", we do
+                                                        # 2 *sum(log(abs)).
+
+  }else{
+    if(nargin<3){
+      invSS <- mult__inv(SS, whichoutput= NULL , 2)$inva
+      logabsdetSS <- mult__inv(SS, whichoutput= NULL , 2)$logabsdet
+    }else if (nargin==3){
+      logabsdetSS <- mult__inv(SS, whichoutput= NULL,'logabsdet')$logabsdet
+    }
+
+    a <- invSS[[1]]
+    b <- array(res, c(N, 1, T))
+    SSres <- array(mult__prod(a, b, c=NULL), c(N, T))
+    logabsdetSS <- logabsdetSS[[1]]
+  }
+  y <- -0.5*N*log(2*pi)-0.5*logabsdetSS-0.5*abs(colSums(res*SSres)) # Last term is equal to "t(res)*(SS)^(-1)*res"
+
+  return(y)
+
+}

### 2021-11-02 R/GraphicalOutputs.R

@@ -1,0 +1,1662 @@
+#' Generate the graphical outputs for the selected models (Point estimate)
+
+
+#'@param ModelType a string-vector containing the label of the model to be estimated
+#'@param ModelPara List of model parameter estimates (See the "Optimization" function)
+#'@param NumOut list of computed outputs containing the model fit, IRFs, FEVDs, GIRFs, and GFEVDs
+#'@param InputsForOutputs list containing the desired inputs for the construction of the desired output
+#'@param Economies string-vector containing the names of the economies which are part of the economic system
+#'@param FactorLabels string-list based which contains the labels of all the variables present in the model
+#'
+
+
+
+GraphicalOutputs <- function(ModelType, ModelPara, NumOut, InputsForOutputs, Economies, FactorLabels){
+
+# Generate the graph paths and the graph folders
+  # Create the folder output folder in which the outputs will be stored
+  dir.create(paste(tempdir(), "/Outputs", sep=""))
+  dir.create(paste(tempdir(), "/Outputs/",  ModelType, sep=""))
+  dir.create(paste(tempdir(), "/Outputs/",  ModelType, "/Point Estimate", sep=""))
+
+
+  PathsGraphs <- FolderCreationPoint(ModelType, Economies)
+
+# 0) Plot the set of risk factors
+  RiskFactorsGraphs(ModelType, ModelPara, Economies, FactorLabels)
+
+
+# 1) Models for which the estimation is done on a country-by-country basis
+if ( "JPS" %in% ModelType || "JPS jointP" %in% ModelType ||  "GVAR sepQ" %in% ModelType){
+
+    for (i in 1:length(Economies)){dir.create(paste(tempdir(),"/Outputs/", ModelType, "/Point Estimate/Model ", Economies[i], sep=""))}
+
+  # Model fit
+  FitgraphsSep(ModelType, InputsForOutputs[[ModelType]]$Fit$WishGraphs, ModelPara, NumOut, Economies, PathsGraphs)
+
+  # IRF and FEVD
+  IRFgraphsSep(ModelType, NumOut, InputsForOutputs[[ModelType]]$IRF$WishGraphs$RiskFactors,
+               InputsForOutputs[[ModelType]]$IRF$WishGraphs$Yields, InputsForOutputs[[ModelType]]$IRF$horiz,
+               PathsGraphs, Economies)
+
+  FEVDgraphsSep(ModelType, NumOut, InputsForOutputs[[ModelType]]$FEVD$WishGraphs$RiskFactors,
[…]

### 2021-11-02 R/InputsForMLEdensity.R

@@ -1,0 +1,385 @@
+#' Generates several  inputs that are necessary to build the likelihood function
+#'
+#'@param ModelType string-vector containing the label of the model to be estimated
+#'@param Yields time series of yields (JxT or CJ x T)
+#'@param PdynamicsFactors time series of the risk factors (K x T)
+#'@param FactorLabels string-list based which contains the labels of all the variables present in the model
+#'@param mat  vector of maturities (in years) used in the estimation
+#'@param Economies  string-vector containing the names of the economies of the system. \cr
+#'                 If the ModelType selected is "JPS", "JPS jointP", "GVAR sepQ", then only one economy can be selected. \cr
+#'                  For the other models, more than one economy must be selected.
+#'@param DataFrequency  character-based-vector. Avaialable options are: "Daily All Days", "Daily Business Days", "Weekly", "Monthly", "Quarterly", "Annually"
+#'@param JLLinputs    list of necessary inputs for the estimation of JLL-based models (see "JLL" function)
+#'@param GVARinputs list of necessary inputs for the estimation of GVAR-based models (see "GVAR" function)
+#'
+#'
+#'@importFrom pracma null
+#'
+#'@return List of necessary inputs  for constructing the model's log-likelihood function
+#'@examples
+#'\donttest{
+#' # Example 1:
+#' data(CM_Factors)
+#' data(CM_Yields)
+#'
+#' ModelType <- "JPS"
+#' Economies <- "Mexico"
+#' Factors <- RiskFactors
+#' N <- 3
+#' GlobalVar <- c("GBC", "CPI_OECD") # Global Variables
+#' DomVar <- c("Eco_Act", "Inflation") # Domestic Variables
+#' FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType)
+#'
+#' mat <- c(0.25, 0.5, 1, 3, 5, 10)
+#' DataFrequency <- "Monthly"
+#'
+#' i <- length(Economies)
+#' ATSMInputs <- InputsForMLEdensity(ModelType, Yields, Factors, FactorLabels, mat,
+#'                                  Economies, DataFrequency)
+
+
[…]

### 2021-11-02 R/InputsForOutputs.R

@@ -1,0 +1,148 @@
+#' Collect the inputs that are used to construct the numerical and the graphical outputs
+#'
+#'@param ModelType String-vector containing the label of the model to be estimated
+#'@param Horiz Single numerical vector conataining the desired horizon of analysis for the outputs
+#'@param ListOutputWished List of desired graphical outputs. Available options are: "Fit","IRF", "FEVD", "GIRF", "GFEVD".
+#'@param OutputLabel Name of the output label to be stored
+#'@param WishStationarityQ User must set 1 is she whises to impose the largest eigenvalue under the Q to be strictly
+#'                       smaller than 1, otherwise set 0.
+#'@param WishGraphYields Binary variable: set 1, if the user wishes graphs to be generated; or set 0, otherwise. Default is set as "0".
+#'@param WishGraphRiskFactors Binary variable: set 1, if the user wishes graphs to be generated; or set 0, otherwise. Default is set as "0".
+#'@param WishOrthoJLLgraphs Binary variable: set 1, if the user wishes orthogonalized JLL-based graphs to be generated; or set 0, otherwise.
+#'                         Default is set as "0"
+#'@param WishBootstrap  Binary variable: set 1, if the user wishes graphs to be generated; or set 0, otherwise. Default is set as "0".
+#'@param ListBoot       List containing the four following elements:
+#'\enumerate{
+#'  \item "methodBS": Desired bootstrap method among (a) 'bs' (standard residual bootstrap), (b) 'wild' (wild bootstrap),
+#'                   (c) 'block' (block  bootstrap);
+#'  \item "BlockLength": if block bootstrap is chosen, then the user has to specify the lenght of the block (single numerical vector);
+#'  \item "ndraws": number of draws;
+#'  \item "pctg": level of confidence (single numerical vector expressed in basis points)
+#'}
[…]

### 2021-11-02 R/JLL.R

@@ -1,0 +1,471 @@
+#' Set of inputs present at JLL's P-dynamics
+#'
+#'@param NonOrthoFactors Risk factors before the orthogonalization (FxT)
+#'@param N Number of country-specific spanned factors
+#'@param JLLinputs List of necessary inputs to estimate JLL outputs:
+#'  \enumerate{
+#'      \item Economies:  set of economies that are part of the economic system (string-vector)
+#'      \item "DomUnit": name of the economy which is assigned as the dominant unit. \cr
+#'                  If no dominant unit is assigned, then this variable is defined as "None"
+#'      \item WishSigmas: equal to "1" if one wishes the variance-covariance matrices and the Cholesky factorizations
+#'                  (can take long if they need to be estimated). Set "0", otherwise.
+#'      \item SigmaNonOrtho: NULL or some K x K matrix from the non-orthogonalized dynamics
+#'      \item JLLModelType: available options are "JLL original", "JLL jointSigma"  or "JLL NoDomUnit"
+#'  }
+
+#'@examples
+#'\donttest{
+#'if (requireNamespace('neldermead', quietly = TRUE)) {
+#'
+#'data(CM_Factors)
+#' ZZ <- RiskFactors
+#' N <- 3
+#'
+#' JLLinputs <- list()
+#' JLLinputs$Economies <- c( "China", "Brazil", "Mexico", "Uruguay")
+#' JLLinputs$DomUnit <- "China"
+#' JLLinputs$WishSigmas <- 1
+#' JLLinputs$SigmaNonOrtho <- NULL
+#' JLLinputs$JLLModelType <- "JLL original"
+
+#' JLL(ZZ, N, JLLinputs)
+#'
+#' #'} else {
+#'  message("skipping functionality due to missing Suggested dependency")
+#'}
+#'}
+
+#'@details
+#' For the models 'JLL original' or "JLL jointSigma" the name of one dominant economy must assigned.\cr
[…]

### 2021-11-02 R/K1XQStationary.R

@@ -1,0 +1,29 @@
+#' Impose stationarity under the Q-measure
+#'
+#'@param StationaryEigenvalues Binary variable: set "1" if the user whises the largest eigenvalue
+#'                            to be strictly smaller than 1. Set "0", otherwise
+#'
+#'@examples
+#' stat <- 1 # Takes values 1 and 0
+#'K1XQStationary(stat)
+#'
+#'
+#'@returns
+#'list
+#'
+#'@export
+
+K1XQStationary<- function(StationaryEigenvalues){
+
+  K1Type <- list()
+
+  if (StationaryEigenvalues == 1){
+    K1Type$SepQ <- paste("K1XQ: ", "Jordan", "; stationary", sep="")
+    K1Type$JointQ <- paste("K1XQ: ", "Jordan", " MultiCountry", "; stationary", sep="")
+  }else{
+    K1Type$SepQ <- paste("K1XQ: ", "Jordan", sep="")
+    K1Type$JointQ <- paste("K1XQ: ", "Jordan", " MultiCountry", sep="")
+  }
+
+  return(K1Type)
+}

### 2021-11-02 R/LabFac.R

@@ -1,0 +1,109 @@
+#' Generates the labels factors
+#'
+#'@param N number of spanned factors per country (scalar)
+#'@param DomVar character-vector containing the names of the domestic variables
+#'@param GlobalVar character-vector containing the names of the global variables
+#'@param Economies string-vector containing the names of the economies which are part of the economic system
+#'@param ModelType string-vector containing the label of the model to be estimated
+#'
+#'
+#'@examples
+#'N <- 2
+#'DomVar <- c("inflation", "Economic growth")
+#'GlobalVar <- "Commodity Prices"
+#'Economies <- c("U.S.", "Canada", "Germany", "Japan")
+#'ModelType <- "JPS"
+#'
+#'VarLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+#'
+#'
+#'@returns
+#' List containing the country-specific risk factor labels
+#'
+#'@export
+
+LabFac <- function(N, DomVar, GlobalVar, Economies, ModelType){
+
+  M <- length(DomVar) # Number of domestic unspanned factors
+  C <- length(Economies) # Number of countries of the economic system
+
+  FactorLabels <- list()
+  FactorLabels$Spanned <- LabelsSpanned(N)
+  FactorLabels$Domestic <- c(DomVar, FactorLabels$Spanned)
+  FactorLabels$Star <-  LabelsStar(FactorLabels$Domestic)
+  FactorLabels$Global <- GlobalVar
+
+
+  labelsDomVar <- c()
+  labelsDomVarJLL <- c()
+  j <-0
+  for (i in 1:C){
+    count0 <- (i-1)*j
+    for (j in 1:(N+M)){
+      count1 <- count0 + j
+      labelsDomVar[count1] <- paste(FactorLabels$Domestic[j], Economies[i])
+      if ("JLL original" %in% ModelType || "JLL NoDomUnit" %in% ModelType || "JLL jointSigma" %in% ModelType){
+        labelsDomVarJLL[count1] <- paste(FactorLabels$Domestic[j], Economies[i], "JLL")
+      }
+    }
+    FactorLabels$Tables[[Economies[i]]] <- labelsDomVar[(count0+1):count1]
+    FactorLabels$TablesJLL[[Economies[i]]] <- labelsDomVarJLL[(count0+1):count1]
+  }
+
+
+  idx0 <- 0
+  for (a in 1:C){
+    idx1 <- idx0 + N + M
[…]

### 2021-11-02 R/MinorOptimizationFunctions.R

@@ -1,0 +1,201 @@
+####################################################################################
+#' Check whether one element is a subset of another element
+#'
+#' @param s1 smaller subset
+#' @param s2 complete set
+#'@references
+#' This function is based on the "contain" function by Le and Singleton (2018). \cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+#'
+
+contain <- function(s1, s2){
+  y <- grepl(s1, s2)
+  return(y)
+}
+
+
+
+###################################################################################
+#' Transform a number bounded between a lower bound and upper bound to x by:
+#'
+#' @param y  Number to be transformed (scalar)
+#' @param lb lower bound (scalar)
+#' @param ub upper bound (scalar)
+#'
+#'@references
+#' This function is based on the "bound2x" function by Le and Singleton (2018). \cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+
+bound2x <- function(y, lb, ub){
+
+  x <- pos2x((y - lb)/(ub - y))
+  return(x)
+}
+
+#######################################################################################################
+#' Transform a positive number y to back to x by:
+#'
+#' @param y scalar
+#'@references
+#' This function is based on the "pos2x" function by Le and Singleton (2018).\cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+
+pos2x <- function(y){
+
+  x <- y + log1p(-expm1(-y)-1)
+  return(x)
+}
+
+########################################################################################################
+#' Transform x to a positive number by: y = log(e^x + 1)
+#'
[…]

### 2021-11-02 R/NumOutputs.R

@@ -1,0 +1,2158 @@
+#' Construct the model numerical outputs (model fit, IRFs, GIRFs, FEVDs, and GFEVDs)
+#'
+#'@param ModelType a string-vector containing the label of the model to be estimated
+#'@param ModelPara List of model parameter estimates (See the "Optimization" function)
+#'@param InputsForOutputs list conataining the desired horizon of analysis for the model fit, IRFs, GIRFs, FEVDs, and GFEVDs
+#'@param FactorLabels  a string-list based which contains all the labels of all the variables present in the model
+#'@param Economies a string-vector containing the names of the economies which are part of the economic system
+#'
+#'
+#'@examples
+#' # See examples in the vignette file of this package (Section 4).
+#'
+#'
+#'@returns
+#'List of the model numerical outputs, namely
+#'\enumerate{
+#'\item Model fit of bond yields
+#'\item IRFs
+#'\item FEVDs
+#'\item GIRFs
+#'\item GFEVDs
+#'}
+#'
+#'
+#'@export
+
+
+
+NumOutputs <- function(ModelType, ModelPara, InputsForOutputs, FactorLabels, Economies){
+
+
+  AllNumOutputs <- list()
+
+  # If one chooseS models in which the estimation is done country-by-country
+  if ( "JPS" %in% ModelType || "JPS jointP" %in% ModelType ||  "GVAR sepQ" %in% ModelType){
+    NumOutSep <- OutputConstructionSep(ModelType, ModelPara, InputsForOutputs, FactorLabels, Economies)
+
+  }
+  # If one chooseS models in which the estimation is done jointly for all countries
+  if ( "GVAR jointQ" %in% ModelType || "VAR jointQ" %in% ModelType || "JLL original" %in% ModelType
+       || "JLL NoDomUnit" %in% ModelType || "JLL jointSigma" %in% ModelType){
+
+    NumOutJoint <- OutputConstructionJoint(ModelType, ModelPara,InputsForOutputs, FactorLabels, Economies)
+
+
+  }
+
+
+
+  # Prepare final list of outputs
+  if (!exists("NumOutJoint")){ AllNumOutputs <- NumOutSep}
+  if (!exists("NumOutSep")){ AllNumOutputs <- NumOutJoint}
+  if ( exists("NumOutSep") & exists("NumOutJoint")){
+
[…]

### 2021-11-02 R/NumOutputsBootstrap.R

@@ -1,0 +1,1900 @@
+#' Numerical outputs (IRFs, GIRFs, FEVD, and GFEVD) for bootstrap
+#'
+#'@param ModelType string-vector containing the label of the model to be estimated
+#'@param ModelParaBoot list of model parameter estimates (see the "Optimization" function) after a bootstrap draw
+#'@param InputsForOutputs list conataining the desired inputs for the construction of the model fit, IRFs, GIRFs, FEVDs, and GFEVDs
+#'@param FactorLabels string-list based which contains all the labels of all the variables present in the model
+#'@param Economies  string-vector containing the names of the economies which are part of the economic system
+#'
+
+
+NumOutputs_Bootstrap <- function(ModelType, ModelParaBoot, InputsForOutputs, FactorLabels, Economies){
+
+
+  AllNumOutputs <- list()
+
+
+  # If one chooseS models in which the estimation is done country-by-country
+  if ( "JPS" %in% ModelType || "JPS jointP" %in% ModelType ||  "GVAR sepQ" %in% ModelType){
+    NumOutSep <- OutputConstructionSep_BS(ModelType, ModelParaBoot, InputsForOutputs, FactorLabels, Economies)
+  }
+
+  # If one chooseS models in which the estimation is done jointly for all countries
+  if ( "GVAR jointQ" %in% ModelType || "VAR jointQ" %in% ModelType || "JLL original" %in% ModelType
+       || "JLL NoDomUnit" %in% ModelType || "JLL jointSigma" %in% ModelType){
+
+    NumOutJoint <- OutputConstructionJoint_BS(ModelType, ModelParaBoot, InputsForOutputs, FactorLabels, Economies)
+
+  }
+
+
+  # Prepare final list of outputs
+  if (!exists("NumOutJoint")){ AllNumOutputs <- NumOutSep}
+  if (!exists("NumOutSep")){ AllNumOutputs <- NumOutJoint}
+  if ( exists("NumOutSep") & exists("NumOutJoint")){
+
+    for (i in 1:length(NumOutSep)){AllNumOutputs[[i]] <- append(NumOutSep[[i]], NumOutJoint[[i]]) }
+    names(AllNumOutputs) <- names(NumOutSep)
+  }
+
+  return(AllNumOutputs)
+}
+
+
+
+
+
+######################################################################################################
[…]

### 2021-11-02 R/Optimization.R

@@ -1,0 +1,391 @@
+#' Peform the minimization of mean(f)
+#'
+#' @param f vector-valued objective function (function)
+#' @param tol convergence tolerance (scalar). For ML estimation, a reasonable value is tol <- 1e-4
+#' @param varargin list containg starting values and constraints:
+#'                        for each input argument K (of f), we need four inputs that look like:
+#'    \enumerate{
+#'        \item a starting value: K0
+#'        \item a variable label ('K0') followed by a ':' followed by a type of constraint. The constraint can be:
+#'                  \itemize{
+#'                      \item 'bounded': bounded matrix;
+#'                      \item 'Jordan' or 'Jordan MultiCountry': a matrix of Jordan type;
+#'                      \item 'psd': psd matrix;
+#'                      \item 'stationary': largest eigenvalue of the risk-neutral feedback matrix is strictly smaller than 1;
+#'                      \item 'diag' or 'BlockDiag': a diagonal or block diagonal matrix.
+#'                      \item 'JLLstructure': to impose the zero-restrictions on the variance-voriance matrix along
+#'                              the lines of the JLL models
+#'                          }
+#'        \item a lower bound lb (lb <- NULL -> no lower bound)
+#'        \item an upper bound ub (ub <- NULL -> no upper bound)
+#'        \item Specification of the optimization settings:
+#'        \itemize{
+#'                  \item 'iter off': hide the printouts of the numerical optimization routines;
+#'                  \item 'fminunc only': only uses fminunc for the optimization;
+#'                  \item ''fminsearch only': only uses fminsearch for the optimization.
+#'
+#'   }
+#'   }
+#' @param FactorLabels string-list based which contains the labels of all the variables present in the model
+#' @param Economies string-vector containing the names of the economies which are part of the economic system
+#' @param ModelType string-vector containing the label of the model to be estimated
+#' @param JLLinputs inputs used in the estimation of the JLL-based models; Default is set to NULL
+#'
+#'
+#'@examples
+#'#' # See examples in the vignette file of this package (Section 4).
+#'
+#'@return
[…]

### 2021-11-02 R/Reg_K1Q.R

@@ -1,0 +1,166 @@
+#'Estimate the risk-neutral feedbak matrix K1Q using linear regressions
+#'
+#
+#'@param Y        matrix of yields used in estimation  (J x T)
+#'@param mat     vector of maturities (in years) of yields used in estimation (J x 1)
+#'@param Z        pricing factors (can be yields-based or non-yields/macro variables) (N x T)
+#'@param dt       time unit of the model (scalar). For instance, if data is (i) monthly, dt <- 12; (ii) quarterly, dt <- 4; (iii) yearly, dt <- 1.
+#'@param type     'Jordan' -> K1Q will be of the Jordan type \cr
+#                 'NULL' --> no adjustment will be made
+#'
+#'@importFrom pracma mldivide
+#'
+#'@return Risk neutral feedback matrix K1Q.
+#'@examples
+#'data(CM_Yields)
+#'
+#' Y_China <- Yields[1:6,]
+#' Z_China <- Spanned_Factors(Y_China, Economies ="China", N=3)
+#' mat <-c(0.25 , 0.5 , 1, 3, 5, 10)
+#' dt <- 1/12
+#'type <- 'Jordan'
+#'Reg_K1Q(Y_China, mat, Z_China, dt, type)
+#'@references
+#' This function is based on the "Reg_K1Q" function by Le and Singleton (2018). \cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+#'
+#'@export
+
+
+
+
+Reg_K1Q <- function(Y, mat, Z, dt,type){
+
+
+  # Step 1: Define the boundaries of the maturities that will be interpolated
+  M <- round(max(mat)/dt) # longest maturity adjusted for the time units of the model
+  m <- round(min(mat)/dt) # shortest maturity adjusted for the time units of the model
+
+  # Step 2: Interpolate yields by spline. Interpolation is made for each time unit of the model.
+  b <- matrix(data=NA, ncol= ncol(Y) , nrow = M )
+
+  for(i in 1:ncol(Y)){
[…]

### 2021-11-02 R/Reg__OLSconstrained.R

@@ -1,0 +1,51 @@
+#' Restricted OLS regression
+#"
+#'@param Y    left hand side variables (M x T)
+#'@param X    regressors (i.e. N-1 variables + the intercept) (N x T)
+#'@param Bcon   constraints matrix (M x N). If Bcon(i,j) = nan --> B(i,j) is a free parameter
+#'@param G      weighting matrix (psd) - (M x M). Default is set to be identity
+#
+#
+#'@return  matrix of coefficient (M x N)
+#'@details
+#'# Estimate of B is obtained by minimizing the objective:\cr
+#'   sum_t (Y_t-B X_t)' G^{-1} (Y_t-B*X_t) \cr
+#' subject to the constraint that B = Bcon for all non-nan entries of Bcon
+#'
+#'@references
+#' This function is based on the "Reg__OLSconstrained" function by Le and Singleton (2018). \cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#' Available at: https://cepr.org/40029
+
+
+
+Reg__OLSconstrained <-function(Y, X, Bcon, G){
+
+
+  M <- dim(Y)[1]
+  T <- dim(Y)[2]
+
+  if (!exists('G')||is.null(G)){
+    G <- diag(M)
+  }
+
+  N <- dim(X)[1]
+
+  id <- is.nan(Bcon) # id= TRUE -> free parameter; id= FALSE -> constrained
+
+  B <- matrix(0, nrow=M, ncol=N)
+  B[!id] <- Bcon[!id]
+
+  if (any(B!=0)){
+    Y <- Y - B%*%X
+  }
+
+  bigX <- kronecker((X%*%t(X))/T, solve(G))
+  bigY <- solve(G,Y%*%t(X)) /T
+
+
+  B[id] <- solve(bigX[id, id], bigY[id])
+
+  return(B)
+}

### 2021-11-02 R/Transition_Matrix.R

@@ -1,0 +1,118 @@
+#' Compute the transition matrix required in the estimation of the GVAR model
+#'
+#'@param t_First      Sample starting date (year)
+#'@param t_Last       Sample last date (year)
+#'@param Economies    Vector containing the names of all the economies of the system.
+#'@param type         Three possibilities:
+#'\itemize{
+#'      \item "Full Sample": if one wishes ALL weight matrices of each year from which data is available (it may extrapolate the sample period);
+#'      \item "Sample Mean": if one wishes a SINGLE weight matrix containing the average of weights over of the entire sample period;
+#'      \item Some year in particular (e.g. "1998", "2005" ...).
+#'}
+#'
+#'@param DataPath     path of the Excel file containing the data (if any). The default is linked to the Excel file available in the package.
+#'@param Data         Data for computing the transition matrix. Default is set to NULL.
+#'
+#
+#'@return matrix or list of matrices
+#'
+#'@details
+#' NOTE: if there is missing data for any country of the system for that particularly year,
+#' then the transition matrix will include only NAs.
+#'
+#'@examples
+#'data(CM_Trade)
+#'
+#' t_First <- "2006"
+#' t_Last <-  "2019"
+#' Economies <- c("China", "Brazil", "Mexico", "Uruguay")
+#' type <- "Sample Mean"
+#' Transition_Matrix(t_First, t_Last, Economies, type, DataPath = NULL, Data = TradeFlows)
+#'
+#'
+#'@export
+
+
+Transition_Matrix <- function(t_First, t_Last, Economies, type, DataPath = NULL, Data = NULL){
+
+
+  if (sjmisc::is_empty(Data)){
+
+  if (sjmisc::is_empty(DataPath)){ DataPath <- system.file("extdata", "TradeData.xlsx", package = "MultiATSM") }
+
+  tab_names_Trade <- readxl::excel_sheets(DataPath)
+  list_all_Trade <- lapply(tab_names_Trade, function(x) readxl::read_excel(path = DataPath, sheet = x))
+  names(list_all_Trade) <- tab_names_Trade
+
+  L <- length(list_all_Trade)
+
[…]

### 2021-11-02 R/VAR.R

@@ -1,0 +1,60 @@
+#' Estimates a VAR(1)
+
+#' @param RiskFactors matrix containing all the risk factors (K x T)
+#' @param VARtype string-vector which accommodates two possibilities: 'unconstrained' or 'constrained'
+#' @param Bcon constraints matrix (K+1 x N) - should contain an intercept. IfBcon(i,j) = nan --> B(i,j) is a free parameter. \cr
+#'              Default is set to NULL.
+
+#'@return  intercept, feedback matrix and the variance-covariance matrix of a VAR(1)
+#'@examples
+#' data("CM_Factors")
+#' #Example 1
+#' VAR(RiskFactors, VARtype= 'unconstrained')
+
+#' #Example 2
+#' K <- nrow(RiskFactors)
+#' Bcon <-matrix(0, nrow = K, ncol = K+1)
+#' Bcon[,1:3] <- NaN
+#' VAR(RiskFactors, VARtype= 'constrained', Bcon)
+#'
+#'
+#'@export
+
+VAR <- function(RiskFactors, VARtype, Bcon = NULL){
+
+  K <- nrow(RiskFactors)
+  T <- ncol(RiskFactors)
+  LHS <- RiskFactors[,2:T]
+  RHS <- RiskFactors[,1:(T-1)]
+
+  # Unconstrained
+  if (VARtype == 'unconstrained'){
+    RegVAR <-stats::lm( t(LHS)~ t(RHS)) # VAR(1) under the P.
+    K0Z <- t(t(RegVAR$coefficients[1,]))
+    K1Z <- t(RegVAR$coefficients[2:(K+1),])
+    eZ <- RegVAR$residuals
+    eZ <- t(eZ)
+    SSZ <- eZ%*%t(eZ)/dim(eZ)[2]
+  }
+
+  # Constrained
+  if (VARtype == 'constrained'){
+
+    intercept <- rep(1, times= T-1)
+    RHS <- rbind(intercept, RHS)
+
+    Coeff <- Reg__OLSconstrained(LHS, RHS, Bcon, G=NULL)
+    colnames(Coeff) <- rownames(RHS)
+    rownames(Coeff) <- rownames(RHS)[-1]
+
+    K0Z <- as.matrix(Coeff[,1])
+    K1Z <- Coeff[,2:(K+1)]
+    eZ <- LHS - Coeff%*%RHS
[…]

### 2021-11-02 R/VariableLabelsOptimization.R

@@ -1,0 +1,49 @@
+#' Create the variable labels used in the estimation
+#'
+#'@param ModelType a string-vector containing the label of the model to be estimated
+#'@param WishStationarityQ User must set "1" is she whises to impose the largest eigenvalue under the Q to be strictly
+#'                       smaller than 1. Otherwise set "0"
+#'
+#'
+#'@examples
+#'ModelType <- "GVAR jointQ"
+#'WishStationarityQ <- 1
+#'ParaLabels(ModelType, WishStationarityQ)
+#'
+#'@returns
+#' list containing the features of the parameters that will be used in the estimation
+#'@export
+
+
+
+ParaLabels <- function(ModelType, WishStationarityQ){
+
+ParaLabelsList <- list()
+
+
+
+ParaLabelsList[[ModelType]][["r0"]] <- "@r0: bounded"
+ParaLabelsList[[ModelType]][["se"]] <- "@se: bounded"
+ParaLabelsList[[ModelType]][["K0Z"]] <- "@K0Z: bounded"
+ParaLabelsList[[ModelType]][["K1Z"]] <- "@K1Z: bounded"
+
+
+# K1XQ
+K1XQType <- K1XQStationary(WishStationarityQ)$SepQ
+
+if (ModelType == "JPS" || ModelType == "JPS jointP" || ModelType == 'GVAR sepQ' ){
+  ParaLabelsList[[ModelType]][["K1XQ"]] <- K1XQStationary(WishStationarityQ)$SepQ
+} else{
+  ParaLabelsList[[ModelType]][["K1XQ"]] <- K1XQStationary(WishStationarityQ)$JointQ
+}
+
+
+# SSZ
+if (ModelType == "JPS" || ModelType == "JPS jointP" || ModelType == 'VAR jointQ' ){ ParaLabelsList[[ModelType]][["SSZ"]] <- "SSZ: psd" }
+if (ModelType == "GVAR sepQ" || ModelType == 'GVAR jointQ'){  ParaLabelsList[[ModelType]][["SSZ"]] <- "SSZ: BlockDiag" }
+if (ModelType == "JLL original" || ModelType == "JLL NoDomUnit"){ ParaLabelsList[[ModelType]][["SSZ"]] <- "@SSZ: bounded" } # Variance-covariance matrix is not estimated under Q
+if (ModelType == "JLL jointSigma"){ ParaLabelsList[[ModelType]][["SSZ"]] <- "SSZ: JLLstructure" }
[…]

### 2021-11-02 R/df__dx.R

@@ -1,0 +1,104 @@
+#' Computes numerical first order derivative of f(x)
+#'
+#' @param f function which contains vector (J x T) valued function handle
+#' @param x parameter values
+#'
+#' @return
+#' transformed matrix (MN x JT)
+#' @references
+#' This function is based on the "df__dx" function by Le and Singleton (2018). \cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+#'
+
+
+
+df__dx <-function(f, x){
+
+  h0 <- array(NaN,c(dim(x)))
+  h0[] <- pmax(pmin(abs(x)*1e-3, 1e-3), 1e-6) # delta
+
+  fx0_o <- f(x=x) # f(x)
+  fis <- 0
+
+  if (is.list(fx0_o)){
+    fx0 <- f(x=x)
+    fis <- 1
+  }else{
+    fx0 <- fx0_o
+  }
+
+  # checking if f(x+h) and f(x-h) is admissible:
+  hxp <- h0; hxm <- h0
+
+  for (i in 1:numel(x) ){
+    fixed <- 0
+    count <- 1
+    while (!fixed && count<10){
+      xp <- x
+      xp[i] <- x[i]+hxp[i] # f(x+ delta)
+      fxp <- f(x=xp)
+      if (abs(mean((t(fxp)-t(fx0))/hxp[i]))>1e8){ # [f(x+ delta) - f(x)]/delta
+        hxp[i] <- hxp[i]/2
+        count <- count+1
+      }else{
+        fixed <- 1
+      }
+    }
+    if (!fixed){ hxp[i] <- 0}
+
+    fixed <- 0
+    count <- 1
+    while (!fixed && count<10){
+      xm <- x
+      xm[i] <- x[i]-hxm[i] # f(x-delta)
+      fxm <- f(x=xm)
+      if ( abs(mean((t(fx0) -t(fxm))/hxm[i]))>1e8){ # [f(x+ delta) - f(x)]/delta
+        hxm[i] <- hxm[i]/2
+        count <- count+1
+      }else{
+        fixed <- 1
+      }
+    }
+    if (!fixed){ hxm[i] <-0 }
+  }
+
[…]

### 2021-11-02 R/f_with_vectorized_parameters.R

@@ -1,0 +1,589 @@
+#'Use function f to generate the outputs from a ATSM
+#'
+#'@param x vector containing all the vectorized auxiliary parameters
+#'@param sizex matrix (6x2) containing the size information of all parameters
+#'@param f vector-valued objective function (function)
+#'@param con if con = 'concentration', then set the value of the parameter whose name
+#'          contains @@ to empty
+#'@param varargin variable inputs used in the optimization (see inputs from "optimization" function)
+#'@param ModelType string-vector containing the label of the model to be estimated
+#'@param FactorLabels string-list based which contains the labels of all the variables present in the model
+#'@param Economies string-vector containing the names of the economies which are part of the economic system
+#'@param JLLinputs Set of necessary inputs used in the estimation of the JLL-based models (see "JLL" function)
+#'@param nargout if nargout <- 1, returns only the values of the likelihood function.\cr
+#'               If nargout <- 2, generates the entire set of outputs
+#'
+#'@references
+#' This function is modified version of the "f_with_vectorized_parameters" function by Le and Singleton (2018).\cr
+#' "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
+#' (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+#'
+
+
+f_with_vectorized_parameters <- function(x, sizex, f, con, varargin, ModelType, FactorLabels, Economies,
+                                         JLLinputs, nargout){
+
+
+  FF <- 1e9; out <- NULL; to_continue <- NULL
+
+  if (!(any(is.na(x)) || any(is.infinite(x)) ||any(Im(x))) ){
+    to_continue <- update_para(x, sizex, ii = NULL, con, FactorLabels, Economies, JLLinputs, varargin)
+    para <- getpara(to_continue) # extract only the numerical parameters of each variable that will not be concetrated out of the likelihood function.
+  }
+
+
[…]

### 2021-11-02 R/mult_Set_Functions.R

@@ -1,0 +1,386 @@
+###############################################################################################################
+#' Inverts an array of matrices so that:  inva[,,i] = inv(a[,,i])
+
+#'@param a        matrix array (N x N x T)
+#'@param whichoutput      if = 'lobabsdet' computes the log(abs(det(a))) only (text).
+#'@param nargout    "nargout == 1" or "nargout == 2"(scalar)
+
+
+#'@importFrom pracma numel size
+#'
+#'@keywords internal
+#'@return "nargout == 1" returns inva -> matrix array: a^{-1} (N x N x T) \cr
+#'        "nargout == 2" returns inva -> matrix array: a^{-1} (N x N x T) and  logabsdet ->  vector of log(abs(det(a)))  (1 x T)
+
+#'@references
+#' This function is modified version of the "mult__inv" function by Le and Singleton (2018).\cr
+#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models."\cr
+#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
+#'  Available at: https://cepr.org/40029
+
+
+
+mult__inv <- function(a, whichoutput, nargout){
+
+
+  if (!exists("whichoutput")){
+    whichoutput = ""
+  }
+
+  ss <- size(a)
+
+  if (numel(ss)<=2){
+    if (nargout>1){
+      inva <- solve(a)
+      logabsdet <- log(abs(det(a)))
+    } else {
+      if (whichoutput =="logabsdet"){
+        inva <- log(abs(det(a)))
+      } else{
+        inva <- solve(a)
+      }
+    }
+  }
+
+  if (numel(ss)>2){
+
+    if (dim(a)[1]<5){
+      inva <- mult_inv_small(a)$y
+      dd <-  mult_inv_small(a)$dd
+
+      if (nargout>1) {
+        logabsdet <- log(abs(dd))
+      } else{
+        if (whichoutput == 'logabsdet'){
+          inva <- log(abs(dd))
+        }
+      }
+    } else{
+      if (nargout>1){
+        inva <- mult_inv_large(a)
+        logabsdet <- mult_logabsdet(a)
+      } else{
+        if (whichoutput == 'logabsdet') {
+          inva <- mult_logabsdet(a)
+        } else{
+          inva <- mult_inv_large(a)
+        }
[…]

### 2021-11-02 R/pca_weights_one_country.R

@@ -1,0 +1,44 @@
+#' Weigth matrix from principal components (matrix of eigenvectors)
+#
+#'@param Y        matrix dimension (J x T), where  J - the number of maturities and  T - time series length
+#'@param Economy  string-vector containg the name of one economy
+#'
+#'
+#'@return matrix (J x J)
+#'@examples
+#'data("CM_Yields")
+#'pca_weights_one_country(Yields, Economy= "Brazil")
+#'@export
+
+
+
+
+pca_weights_one_country <- function(Y, Economy) {
+
+
+  # Extract the yields of a single country
+  Idx <- grepl(Economy, rownames(Y))
+  Y_CS <- Y[Idx,] # Country-specific yields
+
+  # Store the eigenvectors of the matrix of yields
+  H <- eigen(stats::cov(t(Y_CS)))
+  W <- t(H$vectors)
+
+  J <- nrow(Y_CS) # Total number of maturities
+  M <- round(stats::median(1:J)) # Median of the maturity range
+
+
+  # Adjust the weights to ease the interpretability of the spanned factors
+  # a) LEVEL: if the weights of the level are negative, then invert the sign
+  # (ideally, high values of level should be interpreted as a high values for all yields)
+  if (all(W[1,] < 0)){   W[1,] <- W[1,]*(-1)   }
+  # b) SLOPE: if the slope is downward sloping, then invert the sign
+  # (ideally, high values of slope should be interpreted as a steep high curve)
+  if (W[2,1] > W[2,J]){   W[2,] <- W[2,]*(-1)   }
+  # c) Curvature: if the curvature is U-shapped, then invert the sign
+  # (ideally, the curvature factor should capture an effect that is linked to mid-term maturities)
+  if (W[3,1] > W[3,M] & W[3,J] > W[3,M]){ W[3,] <- W[3,]*(-1) }
+
+
+  return(W)
+}

### 2021-11-02 inst/doc/my-vignette.R

@@ -1,0 +1,681 @@
+## ------------------------------------------------------------------------
+library(MultiATSM)
+
+## ----ModFea, message=FALSE, echo=FALSE-----------------------------------
+ModelLabels <- c("JPS", "JPS jointP", "VAR jointQ", "GVAR sepQ", "GVAR jointQ", 
+                 "JLL original", "JLL NoDomUnit", "JLL jointSigma")
+
+# Rows
+Tab <- data.frame(matrix(nrow = length(ModelLabels), ncol = 0)) 
+rownames(Tab) <- ModelLabels
+
+# Empty columns
+EmptyCol <- c("", "", "", "", "", "", "", "") 
+Tab$EmptyCol0 <- EmptyCol
+# P-dynamics + 2 empty spaces
+Tab$PdynIndUnco <- c("x", "x", "", "", "", "", "", "")
+Tab$PdynIndCo <- c("", "", "", "", "", "", "", "")
+Tab$PdynJointUnco <- c("", "", "x", "", "", "", "", "")
+Tab$PdynJointJLL <- c("", "", "", "", "", "x", "x", "x")
+Tab$PdynJointGVAR <- c("", "", "", "x", "x", "", "", "")
+Tab$EmptyCol1 <- EmptyCol
+Tab$EmptyCol2 <- EmptyCol
+# Q-dynamics + 2 empty spaces
+Tab$QdynInd <- c("x", "", "", "x", "", "", "", "")   
+Tab$QdynJoint <- c("", "x", "x", "", "x", "x", "x", "x") 
+Tab$EmptyCol3 <- EmptyCol
+Tab$EmptyCol4 <- EmptyCol
+# Sigma + 2 empty spaces
+Tab$Ponly <-  c("", "", "", "", "", "x", "x", "")
+Tab$PandQ <- c("x", "x", "x", "x", "x", "", "", "x")
+Tab$EmptyCol5 <- EmptyCol
+Tab$EmptyCol6 <- EmptyCol
+# Dominant Unit
+Tab$DomUnit <- c("", "", "", "", "", "x", "", "x")
+
+# Adjust column names
+ColNames <- c("","","","","JLL", "GVAR", "", "", "", "", "", "", "","", "", "","")
+colnames(Tab) <- ColNames
+
+
+# Generate the table
+suppressWarnings(library(magrittr))
+
+
+kableExtra::kbl(Tab, align = "c", caption = "Model Features") %>%
[…]

### 2022-02-28 inst/doc/my-vignette.R
version 0.1.1

@@ -1,7 +1,7 @@
-## ------------------------------------------------------------------------
+## -----------------------------------------------------------------------------
 library(MultiATSM)
 
-## ----ModFea, message=FALSE, echo=FALSE-----------------------------------
+## ----ModFea, message=FALSE, echo=FALSE----------------------------------------
 ModelLabels <- c("JPS", "JPS jointP", "VAR jointQ", "GVAR sepQ", "GVAR jointQ", 
                  "JLL original", "JLL NoDomUnit", "JLL jointSigma")
 
@@ -13,16 +13,16 @@ rownames(Tab) <- ModelLabels
 EmptyCol <- c("", "", "", "", "", "", "", "") 
 Tab$EmptyCol0 <- EmptyCol
 # P-dynamics + 2 empty spaces
-Tab$PdynIndUnco <- c("x", "x", "", "", "", "", "", "")
+Tab$PdynIndUnco <- c("x", "", "", "", "", "", "", "")
 Tab$PdynIndCo <- c("", "", "", "", "", "", "", "")
-Tab$PdynJointUnco <- c("", "", "x", "", "", "", "", "")
+Tab$PdynJointUnco <- c("", "x", "x", "", "", "", "", "")
 Tab$PdynJointJLL <- c("", "", "", "", "", "x", "x", "x")
 Tab$PdynJointGVAR <- c("", "", "", "x", "x", "", "", "")
 Tab$EmptyCol1 <- EmptyCol
 Tab$EmptyCol2 <- EmptyCol
 # Q-dynamics + 2 empty spaces
-Tab$QdynInd <- c("x", "", "", "x", "", "", "", "")   
-Tab$QdynJoint <- c("", "x", "x", "", "x", "x", "x", "x") 
+Tab$QdynInd <- c("x", "x", "", "x", "", "", "", "")   
+Tab$QdynJoint <- c("", "", "x", "", "x", "x", "x", "x") 
 Tab$EmptyCol3 <- EmptyCol
 Tab$EmptyCol4 <- EmptyCol
 # Sigma + 2 empty spaces
@@ -52,19 +52,19 @@ kableExtra::pack_rows("Unrestricted VAR", 1, 3 , label_row_css = "background-col
 kableExtra::pack_rows("Restricted VAR (GVAR)", 4, 5, label_row_css = "background-color: #666; color: #fff;") %>%
 kableExtra::pack_rows("Restricted VAR (JLL)", 6, 8, label_row_css = "background-color: #666; color: #fff;")
 
[…]

### 2022-04-14 inst/doc/my-vignette.Rmd
version 0.2.1

@@
 All the frameworks of this package are based on the unspanned economic risk framework from @JoslinPriebschSingleton2014 (JPS, 2014).
-In essence, these models assume the absence of arbitrage opportunities, consider a linear state space representations of the yield curve dynamics and offer a tractable approach to simultaneously combine the traditional yield curve factors (spanned factors) along with economic and financial variables (unspanned factors).
+In essence, these models assume the absence of arbitrage opportunities, consider a linear state space representation of the yield curve dynamics and offer a tractable approach to simultaneously combine the traditional yield curve factors (spanned factors) along with economic and financial variables (unspanned factors).
 
@@
 More specifically, the *MultiATSM* includes models in which the estimation is done on a *country-by-country* basis - as in JPS (2014)  - or *jointly* for all the countries of the economic system - as in @JotikasthiraLeLundblad2015 (JLL, 2015) and @CandelonMoura2021 (CM, 2021).
-In all cases, the risk factors dynamics under the $P$-measure follows a VAR(1) model of some sort.
+In all cases, the risk factor dynamics under the $P$-measure follow a VAR(1) model of some sort.
 The table below summarizes the general features of each one of the models available at this package.
@@
 # P-dynamics + 2 empty spaces
-Tab$PdynIndUnco <- c("x", "x", "", "", "", "", "", "")
+Tab$PdynIndUnco <- c("x", "", "", "", "", "", "", "")
 Tab$PdynIndCo <- c("", "", "", "", "", "", "", "")
-Tab$PdynJointUnco <- c("", "", "x", "", "", "", "", "")
+Tab$PdynJointUnco <- c("", "x", "x", "", "", "", "", "")
 Tab$PdynJointJLL <- c("", "", "", "", "", "x", "x", "x")
@@
 # Q-dynamics + 2 empty spaces
-Tab$QdynInd <- c("x", "", "", "x", "", "", "", "")   
-Tab$QdynJoint <- c("", "x", "x", "", "x", "x", "x", "x") 
+Tab$QdynInd <- c("x", "x", "", "x", "", "", "", "")   
+Tab$QdynJoint <- c("", "", "x", "", "x", "x", "x", "x") 
 Tab$EmptyCol3 <- EmptyCol
@@
 The *MultiATSM* package contains the four data-sets used in CM (2021).
-The first set of data comprises several time series of zero-coupon bonds yields from four emerging markets: China, Brazil, Mexico, and Uruguay.
+The first set of data comprises several time series of zero-coupon bond yields from four emerging markets: China, Brazil, Mexico, and Uruguay.
 It is worth noting that this package requires *(i)* for estimation purposes, the maturities of the bond yields to be the same for all countries (although the function *DataForEstimation* handles different maturities across countries as inputs, the outputs generated by this same function are bond yields of the same maturities for all the economies); *(ii)* bond yields to be expressed in percentage terms (and not in basis points) per annum.
@@
 Specifically, along the terminology defined in JPS (2014), this data-set contains *(i)* country-specific spanned factors (denoted by level, slope, and curvature - see Section \@ref(S:SpaFac) for a more detailed description) and *(ii)* an array of country-specific and global unspanned factors (namely, some measure of economic growth and inflation).
-Similarly to the case of the bond yields data, the measures of economic and financial variables must be constructed by the user.
+Similarly to the case of the bond yield data, the measures of economic and financial variables must be constructed by the user.
 ```{r}
@@
 
-The two last block of data are necessary for the estimation of the GVAR-based models. The trade flows database
-presents the sum of the value of all goods imports and exports between any two countries of the sample on yearly basis since 1948.
+The two last blocks of data are necessary for the estimation of the GVAR-based models. The trade flows database
+presents the sum of the value of all goods imports and exports between any two countries of the sample on a yearly basis since 1948.
 All the values are free on board and are expressed in U.S.
 dollars.
-This data is used to construct the transition matrix of the GVARs models.
+These data are used to construct the transition matrix of the GVARs models.
 ```{r}
@@
 The GVAR factors database casts country-specific lists of all the factors that are used in the estimation of each country's VARX(1,1,1).
-A specific function for computing the star risk factors is detailed the Section \@ref(S:SectionGVAR).
+A specific function for computing the star risk factors is detailed in the Section \@ref(S:SectionGVAR).
 ```{r}
@@
 ## Import data from Excel files {#S:SectionImport}
-This package also offers an automatized procedure to extract data from Excel files and to, subsequentially, prepare the risk factors database that are directly used in the model estimation.
-The use of the package functions requires that the databases *(i)* are constructed in separate Excel files for the unspanned factors, the term structure data, and the measures of interdependence (for the GVAR-based models); *(ii)* contain, in each Excel file, one separate tab per country (in addition to the global variables, in the case of the unspanned factors database); and *(iii)* have identical variable labels across all the tabs within each file.
+This package also offers an automatized procedure to extract data from Excel files and to, subsequentially, prepare the risk factor databases that are directly used in the model estimation.
+The use of the package functions requires that the databases *(i)* are constructed in separate Excel files for the unspanned factors, the term structure data, and the measures of interdependence (for the GVAR-based models); *(ii)* contain, in each Excel file, one separate tab per country (in addition to the global variables, in the case of the unspanned factors' database); and *(iii)* have identical variable labels across all the tabs within each file.
 For illustration, see the Excel file available at the package.
@@
 The units of this vector must be expressed in *years*.
-Alternatively, the user can rely on the *Maturities* routine to extract the the bond maturities form the time series of bond yields (see the documentation of the *Maturities* function).
+Alternatively, the user can rely on the *Maturities* routine to extract the bond maturities from the time series of bond yields (see the documentation of the *Maturities* function).
 In this case, the user must specify the **unit in which the label of the bond yields** are expressed, e.g.
@@
 Two possibilities are available.
-The *"unconstrained"* form estimates the model without any constrains using standard OLS regressions for each one of the model equations.
-The *"constrained"* form prevents foreign-spanned-factors to impact any domestic factor.
-In such a case, the VARX(1,1,1) is estimated by restricted least squares.
+The *"unconstrained"* form estimates the model without any constrains, using standard OLS regressions for each model equation.
+The *"constrained: Spanned Factors"* form prevents foreign-spanned-factors to impact any domestic factor in the feedback matrix, whereas *"constrained: '* followed by the name of the risk factor restricts this same factor to be influenced only by its own lagged values and the lagged values of its own star variables.
+In the last two cases, the VARX(1,1,1) is estimated by restricted least squares.
 
@@
 GVARinputs$GVARFactors <- FactorsGVAR
-GVARinputs$VARXtype <- "unconstrained"
+GVARinputs$VARXtype <- "constrained: Inflation"
 ```
@@
 - **Dominant Country**: a string-vector containing the name of the economy which is assigned as the dominant country (applicable for the *JLL original* and *JLL jointSigma* models) or "None" (applicable for the *JLL NoDomUnit*).
- - **Wish the estimation of the sigmas matrices**: this is a binary variable which assume value equal to 1 if the user wishes the estimation of all JLL sigma matrices (i.e.
+ - **Wish the estimation of the sigmas matrices**: this is a binary variable which assumes value equal to 1 if the user wishes the estimation of all JLL sigma matrices (i.e.
 variance-covariance and the Cholesky factorization matrices) and, 0 otherwise.
@@
 ### Out-of-sample forecast settings
-To perform out-of-sample forecasts, the following list-based features have to be detailed:
+To perform out-of-sample forecasts of bond yields, the following list-based features have to be detailed:
 
@@
 
-```{r, fig.cap = "Yield loading on the spanned factors", echo=FALSE}
+```{r, fig.cap = "Yield loadings on the spanned factors", echo=FALSE}
 LabSpaFac <- c("Level", "Slope", "Curvature")
@@
 ## The P-dynamics estimation
-As presented in Table \@ref(tab:ModFea), the dynamics of the risk factors under the $P$-measure evolves according to a VAR(1) model, which may be fully unrestricted (as in the case of the JPS-related models) or somewhat restricted (as in the GVAR and JLL frameworks).
-Below, the usage of each one of these dimensions are illustrated.
+As presented in Table \@ref(tab:ModFea), the dynamics of the risk factors under the $P$-measure evolve according to a VAR(1) model, which may be fully unrestricted (as in the case of the JPS-related models) or somewhat restricted (as in the GVAR and JLL frameworks).
+Below, the usage of each of these dimensions are illustrated.
 
@@
 
-In both cases, the outputs generated are the vector of intercepts in addition to the feedback and the variance-covariance matrices.
+In both cases, the outputs generated are the vector of intercepts and both the feedback and variance-covariance matrices.
 
@@
 The estimation of a GVAR model requires defining a measure of interdependence among the countries of the economic system.
-This information is reported in the transition matrix, the entries of which reflect the degree of interconnections of two entities of this same economic system.
-In this package, the illustration of the transition matrix is based on the average of the cross-borders trade flow weights for the period spanning the years from 2006 to 2019.
+This information is reported in the transition matrix, the entries of which reflect the degree of interconnection of two entities of this same economic system.
+In this package, the illustration of the transition matrix is based on the average of the cross-border trade flow weights for the period spanning the years from 2006 to 2019.
 Note that each row sums up to 1.
@@
 long-run mean of the short rate (*r0*), the risk-neutral feedback matrix (*K1XQ*), the variance-covariance matrix (*SSZ*) from the VAR processes, the standard deviation of the errors from the portfolios of yields observed with error (*se*), in addition to the intercept (*K0Z*) and the feedback (*K1Z*) matrices of the physical dynamics.
-Each one of these parameters must be cast in a individual list that should contain the *(i)* the starting value of the parameter (if any); *(ii)* the variable label followed by a ':' and a type of constraint; *(iii)* a lower bound (if any), and *(iv)* an upper bound (if any).
-The variable labels of *r0*, *se*, *K0Z* and *K1Z* should be preceded by the symbol @ as manner to account for the fact that the solution for these parameters are known in closed-form.
+Each of these parameters must be cast in an individual list that should contain the *(i)* the starting value of the parameter (if any); *(ii)* the variable label followed by a ':' and a type of constraint; *(iii)* a lower bound (if any), and *(iv)* an upper bound (if any).
+The variable labels of *r0*, *se*, *K0Z* and *K1Z* should be preceded by the symbol @ as a manner to account for the fact that the solution of these parameters are known in closed-form.
 For the remaining ones (*K1XQ* and *SSZ*), @LeSingleton2018 provide standardized routines (already incorporated in this package) to set good initial values so that the optimization process runs fast.
@@
   if (ModelType == 'JPS' || ModelType == 'JPS jointP' || ModelType == "GVAR sepQ"){
-    ModelParaList[[ModelType]][[Economies[i]]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType)$Summary
-  }else{ ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs)$Summary}
+    ModelParaList[[ModelType]][[Economies[i]]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs, GVARinputs)$Summary
+  }else{ ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs, GVARinputs)$Summary}
 
@@
 Accordingly, BR (2017) intend to replicate some of the empirical results reported in JPS (2014).
-The R-code used by BR (2017) is also available at available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).
+The R-code used by BR (2017) is also available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).
 
@@
 
-It is relevant to highlight that the script above makes use of the principal component weight provided by BR (2017).
+It is relevant to highlight that the script above makes use of the principal component weights provided by BR (2017).
 Such a matrix is simply a scaled-up version of the one provided by the function *pca_weights_one_country* of this package.
 Accordingly, despite the numerical differences on the weight matrices, both methods generate time series of spanned factors which are perfectly correlated.
-Another difference between the two approaches relates to the construction form of the log-likelihood function (llk): while in the BR (2017) code the llk is expressed in terms of portfolio of yields, the *MultiATSM* package generates this same input directly as a function of observed yields (i.e.
+Another difference between the two approaches relates to the construction form of the log-likelihood function (llk): while in the BR (2017) code the llk is expressed in terms of a portfolio of yields, the *MultiATSM* package generates this same input directly as a function of observed yields (i.e.
 both procedures lead to equivalent llk up to the Jacobian term).
@@
 varargin <- list()
-varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
+varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]], NULL , NULL)
 varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
@@
 # 6) Optimization of the model
-ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs)$Summary
+ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs, GVARinputs)$Summary
 

### 2022-04-14 inst/doc/my-vignette.R

@@ -137,7 +137,7 @@ ForecastList$t0Forecast <- 70 # last sample date for the first forecast
 ## -----------------------------------------------------------------------------
 w <- pca_weights_one_country(Yields, Economy = "Uruguay") 
 
-## ---- fig.cap = "Yield loading on the spanned factors", echo=FALSE------------
+## ---- fig.cap = "Yield loadings on the spanned factors", echo=FALSE-----------
 LabSpaFac <- c("Level", "Slope", "Curvature")
 N <- length(LabSpaFac)
  

### 2023-08-16 inst/doc/Paper-Replications.Rmd
version 0.3.3

---
title: "Paper Replication"
author: Rubens Moura
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    number_sections: true
    df_print: paged
    toc: yes
    toc_depth: 2
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Vignette for package MultiATSM_v2}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc} 
bibliography: references.bib
---

This vignette leverages the *MultiATSM* package to reproduce some empirical findings from three academic papers centered around the unspanned macroeconomic risk framework. The first segment focuses on replicating the outcomes of @JoslinPriebschSingleton2014, which laid the foundation for the unspanned risk ATSM literature through a single-country analysis of the U.S. economy. Subsequently, I present the multi-country model extension proposed by @CandelonMoura2021, employing a GVAR approach to capture risk factor dynamics. Lastly, I introduce the program designed to replicate results detailed in @CandelonMoura2023, an extension of the previous framework to investigate the term structure behavior of major emerging markets during the unprecedented COVID$-19$ pandemic.

A thorough explanation on the usage of the additional functionalities of the *MultiATSM* package is available at the [package vignette](my-vignette.html).

```{r}
library(MultiATSM)
```

# Paper replications
## Joslin, Priebisch and Singleton (2014)
To the best of my knowledge, the original data-set used in @JoslinPriebschSingleton2014 is not publicly available. Nevertheless, there exists an alternative source of data that closely matches JPS (2014) in terms of the construction method and the sample period. This database is constructed by @BauerRudebusch2017 (BR, 2017) and is available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/). In that paper, BR (2017) investigate whether macro-finance term structure models better suit the unspanned macro risk framework by JPS (2014) or the earlier traditional spanned settings as the one by @AngPiazzesi2003JME. Accordingly, BR (2017) intend to replicate some of the empirical results reported in JPS (2014). The R-code used by BR (2017) is also available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).

Supported by BR (2017)'s data-set, the code below uses the *MultiATSM* package to estimate the key model parameters from the ATSM along the lines of JPS (2014). The model estimation is performed using the *JPS* modeling setup.

```{r, eval=FALSE} 
#########################################################################################################
#################################### USER INPUTS ########################################################
#########################################################################################################
# A) Load database data
data("BR_jps_gro_R3")

# B) Decide on general model inputs
ModelType <- "JPS"

StationarityUnderQ <- 0
BiasCorrection <- 0
WishForwardPremia <- 1

Economies <- c("U.S.") # Names of the economies from the economic system
GlobalVar <- c()
DomVar <- c("GRO", "INF") # Country-specific variables

N <- 3 # Number of spanned factors per country

DataFrequency <- "Monthly" # Frequency of the data
UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"

mat <- c(0.25, 0.5, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10) # maturities expressed in years

#########################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
#########################################################################################################
# 2) Minor preliminary work
C <- length(Economies)
M <- length(DomVar)
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType) # Generate the set of labels
Yields <- t(BR_jps_out$Y)
W <- BR_jps_out$W[1:N,] # Use the Wpca matrix from BR (2017)
SpaFac <- W%*%Yields
rownames(SpaFac) <- FactorLabels$Tables$U.S.[-(1:M)]
ZZ <- rbind(t(BR_jps_out$M.o), SpaFac) # Complete set of risk factors

# 3) Prepare the inputs of the likelihood function
i <- length(Economies)
  # 3.1) Compute the inputs that go directly into the log-likelihood function
  ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency)
  # 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
  K1XQ <- ATSMInputs$K1XQ
  SSZ <- ATSMInputs$SSZ
  # 3.3) Adjust the inputs which are funtion of the W matrix
  ATSMInputs$Wpca <- W
  ATSMInputs$We <- t(pracma::null(W))
  ATSMInputs$WpcaFull <- rbind(ATSMInputs$Wpca, ATSMInputs$We)
  ATSMInputs$PP <- SpaFac

  # 4) Build the objective function
  f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)

  # 5) Set the optimization settings
  VarLab <- ParaLabels(ModelType, StationarityUnderQ)

  varargin <- list()
  varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
  varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
  varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
  varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
  varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
  varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
  varargin$OptRun <-  c("iter off")

  LabelVar<- c('Value', 'Label', 'LB', 'UB') # Elements of each parameter
  for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}

  tol <- 1e-4

  # 6) Optimization of the model
  ModelPara <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType)$Summary

```

The tables below compare the ATSM parameter estimates generated from BR (2017) and the *MultiATSM*. Overall, one can note that the differences in the estimates are economically modest.

```{r, echo= FALSE}
options(scipen = 100) # eliminate the scientific notation
data("BR_jps_gro_R3")
data("JPSrep")

RowsQ <- c("$r0$", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_3$" )
TableQ <- data.frame(matrix(NA, ncol = 0, nrow =length(RowsQ)))
row.names(TableQ) <- RowsQ

PackageQ<- c(ModelPara$ests$r0, diag(ModelPara$ests$K1XQ))
BRq <- c(BR_jps_out$est.llk$rho0.cP, diag(BR_jps_out$est.llk$KQ.XX))
TableQ$MultiATSM <- PackageQ
TableQ$'BR (2017)' <- BRq

TableQ <- round(TableQ, digits = 5)

suppressWarnings(library(magrittr))

kableExtra::kbl(TableQ, align = "c", caption = "$Q$-dynamics parameters") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::footnote(general = " $\\lambda$'s are the eigenvalues from the risk-neutral feedback matrix and $r0$ is the long-run mean of the short rate under Q.")
```

```{r, echo= FALSE}
data("BR_jps_gro_R3")
data("JPSrep")

RowsP <- c("PC1", "PC2", "PC3", "GRO", "INF")
ColP <- c(" ", RowsP)

# 1) K0Z and K1Z
# Bauer and Rudebusch coefficients
TablePbr <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
row.names(TablePbr) <- RowsP
colnames(TablePbr) <- ColP

TablePbr[[ColP[1]]] <- BR_jps_out$est.llk$KP.0Z
for(j in 1:length(RowsP) ){TablePbr[[RowsP[j]]] <- BR_jps_out$est.llk$KP.ZZ[,j]}

TablePbr <- round(TablePbr, digits = 5)

# MultiATSM coefficients
TablePMultiATSM <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
row.names(TablePMultiATSM) <- RowsP
colnames(TablePMultiATSM) <- ColP

IdxVar <- c(3:5, 1:2) # indexes to flip the order of the spanned and unspanned factors
TablePMultiATSM[[ColP[1]]] <- ModelPara$ests$K0Z[IdxVar]
ModelPara$ests$K1Z <- ModelPara$ests$K1Z[IdxVar,IdxVar]
for(j in 1:length(RowsP) ){ TablePMultiATSM[[RowsP[j]]] <- ModelPara$ests$K1Z[,j]}

TablePMultiATSM <- round(TablePMultiATSM, digits = 5)
TableP <- rbind(TablePbr,TablePMultiATSM)
row.names(TableP) <- c(RowsP,paste(RowsP," ",sep="")) # trick to avoid rows to have the same name

kableExtra::kbl(TableP, align = "c", caption = "$P$-dynamics parameters") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::add_header_above(c(" "= 1, "K0Z"=1, "K1Z" = 5), bold = T) %>%
  kableExtra::pack_rows("BR (2017)", 1, 5) %>%
  kableExtra::pack_rows("MultiATSM", 6, 10) %>%
  kableExtra::footnote(general = " $K0Z$ is the intercept and $K1Z$ is feedback matrix from the $P$-dynamics.")
```

```{r, echo= FALSE}
data("BR_jps_gro_R3")
data("JPSrep")

se <- data.frame(BR_jps_out$est.llk$sigma.e, ModelPara$ests$se )
rownames(se) <- "se"
colnames(se) <- c("MultiATSM","BR (2017)")
se <- round(se, digits = 7)

kableExtra::kbl(se, align = "c", caption ="Portfolio of yields with errors") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 14) %>%
  kableExtra::footnote(general = " $se$ is the standard deviation of the portfolio of yields observed with errors.")
```

It is relevant to highlight that the script above makes use of the principal component weights provided by BR (2017). Such a matrix is simply a scaled-up version of the one provided by the function *pca_weights_one_country* of this package. Accordingly, despite the numerical differences on the weight matrices, both methods generate time series of spanned factors which are perfectly correlated. Another difference between the two approaches relates to the construction form of the log-likelihood function (llk): while in the BR (2017) code the llk is expressed in terms of a portfolio of yields, the *MultiATSM* package generates this same input directly as a function of observed yields (i.e. both procedures lead to equivalent llk up to the Jacobian term).

## Candelon and Moura (2021)
The multi-country framework introduced in @CandelonMoura2021 was meticulously designed to enhance the tractability of large-scale ATSMs, providing a deeper understanding of the intricate global economic mechanisms that underlie yield curve fluctuations. This framework also reinforces the statistical validity of inference and enhances the predictive capabilities of these models. This novel setup, embodied by the *GVAR jointQ* model class, is benchmarked against the findings of @JotikasthiraLeLundblad2015, which are captured by the *JLL original* model class. The paper showcases an empirical illustration involving China, Brazil, Mexico, and Uruguay.

```{r, eval=FALSE}
#########################################################################################################
#################################### USER INPUTS ########################################################
#########################################################################################################
# A) Load database data
data("CM_Factors")
data('CM_Factors_GVAR')
data('CM_Trade')
data('CM_Yields')

# B) Decide on general model inputs
ModelType <- "GVAR jointQ" # Options: "GVAR jointQ", "JLL original"

StationarityUnderQ <- 0  
BiasCorrection <- 0 
WishForwardPremia <- 0 
FPmatLim <- c(47,48)

Economies <- c("China","Brazil","Mexico", "Uruguay") 
GlobalVar <- c("GBC", "CPI_OECD") 
DomVar <- c("OECD_Det", "Inflation")

N <- 3 # Number of spanned factors per country

OutputLabel <- "CM_2021"
DataFrequency <- "Monthly"
UnitMatYields <- "Month"

# B.1) Decide on specific model inputs
#################################### GVAR-based models ##################################################
t_First_Wgvar <- "2004"
t_Last_Wgvar <-  "2019"
W_type <- 'Sample Mean'
VARXtype <- "unconstrained"
#################################### JLL-based models ###################################################
DomUnit <- "China"
WishSigmas <- 0
SigmaNonOrtho <- NULL
JLLModelType <- ModelType
###################################### BRW inputs  ######################################################
flag_mean <- TRUE
gamma <- 0.1
N_iter <- 500
N_burn <- N_iter*0.15
B <- 50
checkBRW <- 1
B_check <- 1000
#########################################################################################################

# C) Decide on Settings for numerical outputs
Horiz <- 25
DesiredGraphs <- c("Fit", "GIRF", "GFEVD")
WishGraphRiskFac <- 0
WishGraphYields <- 1
WishOrthoJLLgraphs <- 1

# D) Bootstrap settings
WishBootstrap <- 0 
Bootlist <- list()
Bootlist$methodBS <- 'bs' 
Bootlist$BlockLength <- 4 
Bootlist$ndraws <-  1000
Bootlist$pctg   <-  95

# E) Out-of-sample forecast
WishForecast <- 0 
ForecastList <- list()
ForecastList$ForHoriz <- 12 
ForecastList$t0Sample <- 1 
ForecastList$t0Forecast <- 90

#########################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
#########################################################################################################

# 2) Minor preliminary work
C <- length(Economies)
FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType)
ZZ <- RiskFactors

Data <- list()
Data$GVARFactors <- FactorsGVAR
mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields)

# 2.1) Generate GVARinputs, JLLinputs and BRWinputs
ModInputs <- ListModelInputs(ModelType, Data, Economies, VARXtype, t_First_Wgvar, t_Last_Wgvar, W_type,
                             DomUnit, WishSigmas, SigmaNonOrtho, BiasCorrection, flag_mean, gamma, 
                             N_iter, N_burn, B, checkBRW, B_check)

GVARinputs <- ModInputs$GVARinputs
JLLinputs <- ModInputs$JLLinputs
BRWinputs <- ModInputs$BRWinputs

# 3) Prepare the inputs of the likelihood function
ModelParaList <- list()
for (i in 1:C){
  if (( ModelType == "GVAR jointQ" || ModelType == "VAR jointQ" || ModelType == "JLL original"
        || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma" )   & i >1 ){break}

# 3.1) Compute the inputs that go directly into the log-likelihood function
ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, 
                                  JLLinputs, GVARinputs, BRWinputs)

# 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
K1XQ <- ATSMInputs$K1XQ
if (ModelType == "JLL original" || ModelType == "JLL NoDomUnit"){ SSZ <- NULL}else{SSZ <- ATSMInputs$SSZ}

# 4) Build the objective function
f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)

# 5) Set the optimization settings
VarLab <- ParaLabels(ModelType, StationarityUnderQ)

varargin <- list()
varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
varargin$OptRun <-  c("iter off")

LabelVar<- c('Value', 'Label', 'LB', 'UB') 
for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}

tol <- 1e-4
# 6) Optimization of the model
ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType,
                                           JLLinputs, GVARinputs)$Summary
}

# 7) Numerical Outputs
InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,
                                     UnitMatYields, WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, 
                                     WishForwardPremia, FPmatLim, WishBootstrap, Bootlist, WishForecast,
                                     ForecastList)
# A) Fit, IRF, FEVD, GIRF, GFEVD and Risk premia
NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)

# B) Bootstrap
Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs,
                       FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs, BRWinputs)

# C) Out-of-sample forecasting
Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
                            DataFrequency, JLLinputs, GVARinputs, BRWinputs)

```

## Candelon and Moura (2023)
In this paper, @CandelonMoura2023 delve into an exploration of the underlying factors that shape the sovereign yield curves of Brazil, India, Mexico, and Russia in the midst of the COVID$-19$ pandemic crisis. To comprehensively address the intricate web of global macro-financial and particularly health-related interdependencies, the study employs a modeling approach based on a *GVAR jointQ* type framework. By delving into these aspects, the authors aim to provide a nuanced understanding of the dynamics influencing sovereign yield curves in these countries during the challenging circumstances of the pandemic.

```{r, eval=FALSE}
# A) Load database data
data("CM_Factors_2023")
data('CM_Factors_GVAR_2023')
data('CM_Trade_2023')
data('CM_Yields_2023')

# B) Decide on general model inputs
ModelType <- "GVAR jointQ"

StationarityUnderQ <- 0 
BiasCorrection <- 0

WishForwardPremia <- 1 
FPmatLim <- c(47,48)

Economies <- c("Brazil", "India", "Russia", "Mexico") # Names of the economies from the economic system
GlobalVar <- c("US_Output_growth", "China_Output_growth", "SP500") # Global Variables
DomVar <- c("Inflation","Output_growth", "CDS", "COVID") # Country-specific variables

N <- 2 # Number of spanned factors per country

OutputLabel <- "CM_2023" 
DataFrequency <- "Weekly" 
UnitMatYields <- "Month"

# C.1) Decide on specific model inputs
#################################### GVAR-based models ##################################################
t_First_Wgvar <- "2015" 
t_Last_Wgvar <-  "2020" 
W_type <- 'Sample Mean' 
VARXtype <- "constrained: COVID" 
###################################### BRW inputs  ######################################################

# D) Decide on Settings for numerical outputs
Horiz <- 12
DesiredGraphs <- c("GIRF", "GFEVD", "TermPremia") 
WishGraphRiskFac <- 0
WishGraphYields <- 1
WishOrthoJLLgraphs <- 0

# E) Bootstrap settings
WishBootstrap <- 0 #  YES: 1; No = 0.
Bootlist <- list()
Bootlist$methodBS <- 'bs' 
Bootlist$BlockLength <- 4 # necessary input if one chooses the block bootstrap
Bootlist$ndraws <-  100
Bootlist$pctg   <-  95 # confidence level

# F) Out-of-sample forecast
WishForecast <- 0 
ForecastList <- list()
ForecastList$ForHoriz <- 12 # forecast horizon
ForecastList$t0Sample <- 1 # initial sample date
ForecastList$t0Forecast <- 50 # last sample date for the first forecast

##########################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE ########################################
##########################################################################################################
# 2) Minor preliminary work
C <- length(Economies)
FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType)
ZZ <- RiskFactors

Data <- list()
Data$GVARFactors <- GVARFactors
Data$Wgvar <-  Trade_Flows
mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields)

# 2.1) Generate GVARinputs, JLLinputs and BRWinputs
ModInputs <- ListModelInputs(ModelType, Data, Economies, VARXtype, t_First_Wgvar, t_Last_Wgvar, W_type)

GVARinputs <- ModInputs$GVARinputs
JLLinputs <- NULL

# 3) Prepare the inputs of the likelihood function
# 3.1) Compute the inputs that go directly into the log-likelihood function
ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, 
                                  JLLinputs, GVARinputs)

# 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
K1XQ <- ATSMInputs$K1XQ
SSZ <- ATSMInputs$SSZ

# 4) Build the objective function
f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)

# 5) Set the optimization settings
VarLab <- ParaLabels(ModelType, StationarityUnderQ)

varargin <- list()
varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
varargin$OptRun <-  c("iter off")

LabelVar<- c('Value', 'Label', 'LB', 'UB') # Elements of each parameter
for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}

tol <- 1e-4
# 6) Optimization of the model
ModelParaList <- list()
ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, 
                                           JLLinputs, GVARinputs)$Summary

# 7) Numerical Outputs
InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,
                                     UnitMatYields, WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs,
                                     WishForwardPremia, FPmatLim, WishBootstrap, Bootlist, WishForecast,
                                     ForecastList)
# A) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)

# B) Bootstrap
Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs,
                       FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs)
```

# References

### 2023-08-16 inst/doc/my-vignette.Rmd
version 0.3.3

@@
 In Section \@ref(S:SectionEstimation) the estimation procedure is detailed.
-Section \@ref(S:SectionReplication) shows how to use the *MultiATSM* package to estimate ATSMs from scratch as well as to closely reproduce some empirical features of academic papers.
+Section \@ref(S:SectionReplication) shows how to use the *MultiATSM* package to estimate ATSMs from scratch.
 To ease the exposition, throughout the next sections, the database used by CM (2021) is employed to illustrate the use of the various functions available in this package.
@@
 The GVAR factors database casts country-specific lists of all the factors that are used in the estimation of each country's VARX(1,1,1).
-A specific function for computing the star risk factors is detailed in the Section \@ref(S:SectionGVAR).
+The function titled *DatabasePrep* provides assistance to structure the data in a similar fashion to this form.
+Moreover, a specific function for computing the star risk factors is detailed in the Section \@ref(S:SectionGVAR).
 ```{r}
@@
 - **Model type**: a string-vector containing the label of the model to be estimated (as per described in Table \@ref(tab:ModFea)); 
+- **Model parameters corrected for bias**: following the bias correction method described in @BauerRudebuschWu2012 (BRW, 2012), this variable assumes value equal to "1" if the user wishes to implement the framework of BRW (2012) and "0", otherwise;
 - **Frequency of the data**: a string-vector that specifies the frequency of the time series data.
@@
 Alternatively, the user can rely on the *Maturities* routine to extract the bond maturities from the time series of bond yields (see the documentation of the *Maturities* function).
-In this case, the user must specify the **unit in which the label of the bond yields** are expressed, e.g.
+In this case, the user must specify the **unit of time in which the label of the bond yields** are expressed, e.g.
 if the label of the 10-years bond is selected as *Y120M* (*Y10y*), then the option *Month* (*Year*) must be selected;
-- **Factor labels**: a string-list based which contains the labels of all the variables present in the model.
-This input is particularly useful for constructing the plots.
-The routine *LabFac* provides the desired list of factor labels;
 - **Stationarity constraint under the $Q$-dynamics**: a binary variable which takes value equal to "1" if the user wishes the largest eigenvalue under the $Q$-measure to be strictly smaller than 1.
 Otherwise, this same variable must be set to "0";
+- **Computation of forward premia**: this is another binary variable which takes the value equal to "1" if the user wishes the computation of the forward premia or "0", otherwise.
+If the this variable is set to "1", then the user must specify a two-element numerical vector containing the maturties associated with the starting and the ending date of the loan;
 - **Output label**: a single element string-vector which contains the of the name that appears in the file name that stores the model outputs.
@@
 ModelType <- "JPS"
+BiasCorrection <- 1 # 1 = estimate model with bias correction; 0 = estimate model without bias correction
 DataFrequency <- "Monthly"
+UnitMatYields <- "Month" # Available options are: "Month" or "Year".
 Economies <- c("China", "Brazil", "Mexico", "Uruguay")
@@
 mat <- c(0.25, 0.5, 1, 3, 5, 10) # vector of maturities
-FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) 
 StationarityUnderQ <- 0 # 1 = set stationarity condition under  the Q; 0 = no stationarity condition
+WishForwardPremia <- 1 # Wish to estimate the forward premia: 1= Yes; 0= No
+FPmatLim <- c(60,120) # maturty of the loan starts in 60 months and ends in 120 months in the future
 OutputLabel <- "Model_demo" # output label
@@
 Some additional inputs are required if the user intends to estimate the GVAR-ATSM or JLL-ATSM related setups.
-For a convenient use, the extra outputs should be stored in a separate list for each model.
-The GVAR model list of inputs must contain:
+The extra inputs should be stored in separate lists for each model for convenience.
+To assist the user in executing this task, one can use the *ListModelInputs* function.
+Otherwise, the *GVARinputs* and *JLLinputs* constructs can be formed manually as specified below.
+
+To GVAR models, the set of supplementary inputs include:
 
@@
 
+### BRWinputs {#S:SectionBRWinputs}
+In an influential paper, BRW (2012) have shown that the estimates from traditional ATSMs can be severely biased due to the typical small-sample size used in these studies.
+As result, such models may produce
+unreasonably stable expected future short-term interest, therefore distorting the term premia estimates for long-maturity bonds.
+To circumvent this problem, BRW (2012) propose a bootstrap-based method that relies on indirect inference estimation that corrects for the bias present in previous works.
+
+From its version $0.3.1$, the *MultiATSM* package accomodates the framework proposed by BRW (2012) to each one of its ATSM types.
+Some additional inputs must be specified, if the user intends to perform the model estimation along the lines of BRW (2012).
+They are:
+
+- **Mean or median of OLS estimates**: the user must decide whether to compute the mean or median of the OLS estimates after each bootstrap iteration.
+To compute the mean (median), the user must set *TRUE* (*FALSE*);
+- **Number of iteration**: total amount of iterations used in the stochatic approximation algorithm after burn-in;
+- **Number of burn-in iteration**:  quantity of the iterations to be discarded in the first stage of the bias-correction estimation process.
+The recommended number is $15\%$ of the total number of iterations;
+- **Number of bootstrap samples**:  quantity of simulated samples used in each burn-in or actual iteration
+- **Adjustment parameter**: this parameter controls the degree of shrinkage between the difference in the OLS estimates and the BRW (2012)'s bootstrap-based one after each iteration.
+The value of this parameter is fixed and must lie in the interval $]0;1[$;
+- **Perform closeness check**: flag whether the user wishes to check the root mean square distance between the model estimates produced by the bias-correction method and those generated by OLS.
+Default is set to *TRUE*;
+- **Number of bootstrap samples used in the closeness check**: default is set to 100,000;
+
+```{r}
+flag_mean <- TRUE # TRUE = compute the mean; FALSE = compute the median
+gamma <- 0.2 # Adjustment parameter
+N_iter <- 500 # Number of iteration to be conserved 
+N_burn <- N_iter*0.15  # Number of iteration to be discarded 
+B <- 50 # Number of bootstrap samples 
+checkBRW <- 1 # Closeness check
+B_check <- 1000 # Number of bootstrap samples used in the closeness check
+```
+
 ## Additional inputs for numerical and graphical outputs
@@
 - Generalized impulse response functions (GIRFs);
-- Generalized forecast error variance decompositions (GFEVDs).
+- Generalized forecast error variance decompositions (GFEVDs);
+- Bond yields decomposition into their expected component and term premia.
 
@@
 ```{r}
-DesiredGraphs <- c("Fit", "GIRF", "GFEVD") # Available options are: "Fit", "IRF", "FEVD", "GIRF", "GFEVD"
+DesiredGraphs <- c("Fit", "GIRF", "GFEVD") # Available options are: "Fit", "IRF", "FEVD", "GIRF", 
+                                          # "GFEVD", "TermPremia".
 ```
@@
 - **methodBS**: the desired bootstrap procedure.
-Available options are *(i)* standard residual bootstrap ('bs'); *(ii)* wild bootstrap ('wild') and block bootstrap ('block').
+Available options are *(i)* standard residual bootstrap ('bs'); *(ii)* wild bootstrap ('wild'), and block bootstrap ('block').
 If the latest is selected, then the block length must be indicated;
@@
 Secondly, it is decided on the form of the risk factor dynamics under the $P$-measure.
-Finally, the user needs to select general features for the model optimization.
+Finally, the user needs to select the general features for the model optimization.
 
@@
 W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, DataPath = NULL, TradeFlows)
+print(W_gvar)
+```
+
+Starting from package version $0.3.3$, the GVAR setups can be estimated incorporating time-varying interdependence weights.
+This entails that the star factors are constructed by applying trade flow weights from the specific year to adjust the risk factors of the corresponding year.
+To employ this feature, users need to specify the transition matrix type as *Full Sample* and select the **same specific date (year)** for both the initial and final years within the transition matrix.
+This implies that the trade weights of this year are used for solving (link matrices) the GVAR system.
+Taking reference from the variable names in the previous example, an illustration could be set as follows:
+
+```{r}
+t_First <- "2019"
+t_Last <-  "2019"
+type <- "Full Sample"
 ```
@@
 
-A separate routine is provided for computing the foreign-specific factors (also commonly referred as the star variables) used in the estimation of the VARX models.
+A separate routine is provided for computing the foreign-specific factors (also commonly referred to as the star variables) used in the estimation of the VARX models.
 ```{r}
@@
 Further on, the functions of this package are used to replicate some of the results presented in the original papers of JPS (2014) and CM (2021).
+See the [Paper Replication Vignette](Paper-Replications.html) of this package for a detailed reproduction.
 
@@
 ```{r, eval=FALSE}
-###############################################################################################################
-#################################### USER INPUTS ##############################################################
-###############################################################################################################
+########################################################################################################
+#################################### USER INPUTS #######################################################
+########################################################################################################
 # A) Load database data
@@
 # B) Decide on general model inputs
-ModelType <- "JLL original" # available options are "JPS", "JPS jointP", "GVAR sepQ", "VAR jointQ", "GVAR jointQ", "JLL original", "JLL NoDomUnit", "JLL jointSigma".
+ModelType <- "VAR jointQ" # available options are "JPS", "JPS jointP", "GVAR sepQ", "VAR jointQ", 
+                          #"GVAR jointQ", "JLL original", "JLL NoDomUnit", "JLL jointSigma".
 
-StationarityUnderQ <- 1 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1, NO:0
+StationarityUnderQ <- 0 # Wish to impose stationary condition for the eigenvalues of each country: 
+                         #YES: 1,NO:0
+BiasCorrection <- 1 # Wish to estimate the model with the bias correction method of BRW (2012): 
+                    #YES: 1, NO:0
+
+WishForwardPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0
+FPmatLim <- c(60,120) #  If the forward premia is desired, then choose the Min and max maturities of the    
+                        # forward premia. Otherwise set NA
 
@@
 DataFrequency <- "Monthly" # Frequency of the data
-UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"
+UnitMatYields <- "Month" # bond yields time-unit. Available options are "Month" or "Year"
 
 # B.1) Decide on specific model inputs
-#################################### GVAR-based models #########################################################
-if (ModelType == 'GVAR sepQ' || ModelType == 'GVAR jointQ'){
-  # Transition matrix inputs:
-  t_First <- "2006" # First year of the sample
-  t_Last <-  "2019" # Last year of the sample
+#################################### GVAR-based models ##################################################
+t_First_Wgvar <- "2004" # First year of the sample (transition matrix)
+t_Last_Wgvar <-  "2019" # Last year of the sample (transition matrix)
 W_type <- 'Sample Mean' # Method to compute the transition matrix
-  # VARX input:
-  VARXtype <- "unconstrained"
-}
-
-#################################### JLL-based models #########################################################
-if (ModelType =="JLL original" || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma"){
-  DominantCountry <- "China" # (i) one of the countries of the system or (ii) "None"
-}
-################################################################################################################
+VARXtype <- "unconstrained" # (i) "unconstrained" or (ii) ""constrained" (VARX)
+#################################### JLL-based models ###################################################
+DomUnit <- "China"
+WishSigmas <- 1 # Sigma matrix is estimated within the "InputsForMLEdensity" function
+SigmaNonOrtho <- NULL
+JLLModelType <- ModelType
+###################################### BRW inputs  ######################################################
+flag_mean <- TRUE # TRUE = compute the mean; FALSE = compute the median
+gamma <- 0.2 # Adjustment parameter
+N_iter <- 500 # Number of iteration to be conserved 
+N_burn <- N_iter*0.15  # Number of iteration to be discarded 
+B <- 50 # Number of bootstrap samples 
+checkBRW <- 1
+B_check <- 1000 # 
+#########################################################################################################
 
@@
 Horiz <- 50
-DesiredGraphs <- c("IRF", "FEVD") #c("Fit", "IRF", "FEVD", "GIRF", "GFEVD")
+DesiredGraphs <- c("Fit", "IRF", "FEVD") # "Fit", "IRF", "FEVD", "GIRF", "GFEVD", "TermPremia"
 WishGraphRiskFac <- 0
@@
 Bootlist$BlockLength <- 4 # necessary input if one chooses the block bootstrap
-Bootlist$ndraws <-  30
+Bootlist$ndraws <-  5
 Bootlist$pctg   <-  95 # confidence level
@@
 
-###############################################################################################################
-############################### NO NEED TO MAKE CHANGES FROM HERE #############################################
-###############################################################################################################
+#########################################################################################################
+############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
+#########################################################################################################
 
@@
 FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) # Generate the set of labels
-mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields) # Vector of maturities expressed in years
-# (alternatively, it can be completed manually)
 ZZ <- RiskFactors
-# 2.1) Build the GVARinputs
-if (ModelType == 'GVAR sepQ' || ModelType == 'GVAR jointQ'){
-  GVARinputs <- list()
-  GVARinputs$Economies <- Economies
-  GVARinputs$GVARFactors <- FactorsGVAR
-  GVARinputs$VARXtype <- VARXtype
-  GVARinputs$Wgvar <- Transition_Matrix(t_First, t_Last, Economies, W_type, DataPath = NULL, Data = TradeFlows)
-} else { GVARinputs <- NULL }
+if(any(ModelType == c("GVAR sepQ", "GVAR jointQ"))){
+Data <- list()
+Data$GVARFactors <- FactorsGVAR}
+mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields)
 
-# 2.2) Build the JLLinputs
-if (ModelType =="JLL original" || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma"){
-  JLLinputs <- list()
-  JLLinputs$Economies <- Economies
-  JLLinputs$DomUnit <- DominantCountry
-  JLLinputs$WishSigmas <- 1 # Sigma matrix is estimated within the "InputsForMLEdensity" function
-  JLLinputs$SigmaNonOrtho <- NULL
-  JLLinputs$JLLModelType <- ModelType
-}else{
-  JLLinputs <- NULL
-}
+# 2.1) Generate GVARinputs, JLLinputs and BRWinputs
+ModInputs <- ListModelInputs(ModelType, Data, Economies, VARXtype, t_First_Wgvar, t_Last_Wgvar, W_type,
+                             DomUnit, WishSigmas, SigmaNonOrtho, BiasCorrection, flag_mean, gamma, N_iter,
+                             N_burn, B, checkBRW, B_check)
+
+GVARinputs <- ModInputs$GVARinputs
+JLLinputs <- ModInputs$JLLinputs
+BRWinputs <- ModInputs$BRWinputs
 
@@
 for (i in 1:C){
-  if (( ModelType == "GVAR jointQ" || ModelType == "VAR jointQ" || ModelType == "JLL original"
-        || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma" )   & i >1 ){break} # the break avoids the models with joint estimation under the Q to be estimated C times
+  if (( any(ModelType ==c("GVAR jointQ", "VAR jointQ","JLL original", "JLL NoDomUnit","JLL jointSigma"))) 
+      & i >1 ){break}
 
   # 3.1) Compute the inputs that go directly into the log-likelihood function
-  ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, JLLinputs,   GVARinputs)
+  ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency,
+                                    JLLinputs, GVARinputs, BRWinputs)
 
@@
   K1XQ <- ATSMInputs$K1XQ
-  if (ModelType == "JLL original" || ModelType == "JLL NoDomUnit" ){ SSZ <- NULL} else{SSZ <- ATSMInputs$SSZ}
+  if (any(ModelType == c("JLL original", "JLL NoDomUnit"))){ SSZ <- NULL} else{SSZ <- ATSMInputs$SSZ}
 
@@
   if (ModelType == 'JPS' || ModelType == 'JPS jointP' || ModelType == "GVAR sepQ"){
-    ModelParaList[[ModelType]][[Economies[i]]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs, GVARinputs)$Summary
-  }else{ ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs, GVARinputs)$Summary}
-
+ModelParaList[[ModelType]][[Economies[i]]] <- Optimization(f, tol, varargin, FactorLabels,
+                                                           Economies, ModelType)$Summary
+  }else{
+ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType,
+                                           JLLinputs, GVARinputs)$Summary}
 }
@@
 # 7) Numerical and graphical outputs
-InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,                  WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishBootstrap, Bootlist, WishForecast, ForecastList)
-# A) Fit, IRF, FEVD, GIRF and GFEVD
+InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,
+                                     UnitMatYields, WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs,
+                                     WishForwardPremia, FPmatLim, WishBootstrap, Bootlist, WishForecast,
+                                     ForecastList)
+# A) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
 NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)
@@
 # B) Bootstrap
-Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs, FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs)
+Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs,
+                       FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs, BRWinputs)
 
 # C) Out-of-sample forecasting
-Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies, DataFrequency, JLLinputs, GVARinputs)
+Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
+                            DataFrequency, JLLinputs, GVARinputs, BRWinputs)
 
@@
 
-## Paper replications
-### Joslin, Priebisch and Singleton (2014)
-To the best of my knowledge, the original data-set used in @JoslinPriebschSingleton2014 is not publicly available.
-Nevertheless, there exists an alternative source of data that closely matches JPS (2014) in terms of the construction method and the sample period.
-This database is constructed by @BauerRudebusch2017 (BR, 2017) and is available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).
-In that paper, BR (2017) investigate whether macro-finance term structure models better suit the unspanned macro risk framework by JPS (2014) or the earlier traditional spanned settings as the one by @AngPiazzesi2003JME.
-Accordingly, BR (2017) intend to replicate some of the empirical results reported in JPS (2014).
-The R-code used by BR (2017) is also available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).
-
-Supported by BR (2017)'s data-set, the code below uses the *MultiATSM* package to estimate the key model parameters from the ATSM along the lines of JPS (2014).
-
-```{r, eval=FALSE} 
-###############################################################################################################
-#################################### USER INPUTS ##############################################################
-###############################################################################################################
-# A) Load database data
-data("BR_jps_gro_R3")
-
-# B) Decide on general model inputs
-ModelType <- "JPS"
-
-StationarityUnderQ <- 0 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1, NO:0
-
-Economies <- c("U.S.") # Names of the economies from the economic system
-GlobalVar <- c()
-DomVar <- c("GRO", "INF") # Country-specific variables
-
-N <- 3 # Number of spanned factors per country
-
-DataFrequency <- "Monthly" # Frequency of the data
-UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"
-
-mat <- c(0.25, 0.5, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10) # maturities expressed in years
-
-###############################################################################################################
-############################### NO NEED TO MAKE CHANGES FROM HERE #############################################
-###############################################################################################################
-# 2) Minor preliminary work
-C <- length(Economies)
-M <- length(DomVar)
-FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType) # Generate the set of labels
-Yields <- t(BR_jps_out$Y)
-W <- BR_jps_out$W[1:N,] # Use the Wpca matrix from BR (2017)
-SpaFac <- W%*%Yields
-rownames(SpaFac) <- FactorLabels$Tables$U.S.[-(1:M)]
-ZZ <- rbind(t(BR_jps_out$M.o), SpaFac) # Complete set of risk factors
-
-# 3) Prepare the inputs of the likelihood function
-for (i in 1:C){
-# 3.1) Compute the inputs that go directly into the log-likelihood function
-ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency)
-# 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
-K1XQ <- ATSMInputs$K1XQ
-SSZ <- ATSMInputs$SSZ
-# 3.3) Adjust the inputs which are funtion of the W matrix
-ATSMInputs$Wpca <- W
-ATSMInputs$We <- t(pracma::null(W))
-ATSMInputs$WpcaFull <- rbind(ATSMInputs$Wpca, ATSMInputs$We)
-ATSMInputs$PP <- SpaFac
-
-# 4) Build the objective function
-f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)
-
-# 5) Set the optimization settings
-VarLab <- ParaLabels(ModelType, StationarityUnderQ)
-
-varargin <- list()
-varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
-varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
-varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
-varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
-varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
-varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
-varargin$OptRun <-  c("iter off")
-
-LabelVar<- c('Value', 'Label', 'LB', 'UB') # Elements of each parameter
-for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}
-
-tol <- 1e-4
-
-# 6) Optimization of the model
-ModelPara <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType)$Summary
-}
-```
-
-The tables below compare the ATSM parameter estimates generated from BR (2017) and the *MultiATSM*. Overall, one can note that the differences in the estimates are economically modest.
-
-```{r, echo= FALSE}
-options(scipen = 100) # eliminate the scientific notation
-data("BR_jps_gro_R3")
-data("JPSrep")
-
-RowsQ <- c("$r0$", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_3$" )
-TableQ <- data.frame(matrix(NA, ncol = 0, nrow =length(RowsQ)))
-row.names(TableQ) <- RowsQ
-
-PackageQ<- c(ModelPara$ests$r0, diag(ModelPara$ests$K1XQ))
-BRq <- c(BR_jps_out$est.llk$rho0.cP, diag(BR_jps_out$est.llk$KQ.XX))
-TableQ$MultiATSM <- PackageQ
-TableQ$'BR (2017)' <- BRq
-
-TableQ <- round(TableQ, digits = 5)
-
-kableExtra::kbl(TableQ, align = "c", caption = "$Q$-dynamics parameters") %>%
-  kableExtra::kable_classic("striped", full_width = F)  %>%
-  kableExtra::row_spec(0, font_size = 14) %>%
-  kableExtra::footnote(general = " $\\lambda$'s are the eigenvalues from the risk-neutral feedback matrix and $r0$ is the long-run mean of the short rate under Q.")
-```
-
-```{r, echo= FALSE}
-data("BR_jps_gro_R3")
-data("JPSrep")
-
-RowsP <- c("PC1", "PC2", "PC3", "GRO", "INF")
-ColP <- c(" ", RowsP)
-
-# 1) K0Z and K1Z
-# Bauer and Rudebusch coefficients
-TablePbr <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
-row.names(TablePbr) <- RowsP
-colnames(TablePbr) <- ColP
-
-TablePbr[[ColP[1]]] <- BR_jps_out$est.llk$KP.0Z
-for(j in 1:length(RowsP) ){TablePbr[[RowsP[j]]] <- BR_jps_out$est.llk$KP.ZZ[,j]}
-
-TablePbr <- round(TablePbr, digits = 5)
-
-# MultiATSM coefficients
-TablePMultiATSM <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
-row.names(TablePMultiATSM) <- RowsP
-colnames(TablePMultiATSM) <- ColP
-
-IdxVar <- c(3:5, 1:2) # indexes to flip the order of the spanned and unspanned factors
-TablePMultiATSM[[ColP[1]]] <- ModelPara$ests$K0Z[IdxVar]
-ModelPara$ests$K1Z <- ModelPara$ests$K1Z[IdxVar,IdxVar]
-for(j in 1:length(RowsP) ){ TablePMultiATSM[[RowsP[j]]] <- ModelPara$ests$K1Z[,j]}
-
-TablePMultiATSM <- round(TablePMultiATSM, digits = 5)
-TableP <- rbind(TablePbr,TablePMultiATSM)
-row.names(TableP) <- c(RowsP,paste(RowsP," ",sep="")) # trick to avoid rows to have the same name
-
-kableExtra::kbl(TableP, align = "c", caption = "$P$-dynamics parameters") %>%
-  kableExtra::kable_classic("striped", full_width = F)  %>%
-  kableExtra::row_spec(0, font_size = 14) %>%
-  kableExtra::add_header_above(c(" "= 1, "K0Z"=1, "K1Z" = 5), bold = T) %>%
-  kableExtra::pack_rows("BR (2017)", 1, 5) %>%
-  kableExtra::pack_rows("MultiATSM", 6, 10) %>%
-  kableExtra::footnote(general = " $K0Z$ is the intercept and $K1Z$ is feedback matrix from the $P$-dynamics.")
-```
-
-```{r, echo= FALSE}
-data("BR_jps_gro_R3")
-data("JPSrep")
-
-se <- data.frame(BR_jps_out$est.llk$sigma.e, ModelPara$ests$se )
-rownames(se) <- "se"
-colnames(se) <- c("MultiATSM","BR (2017)")
-se <- round(se, digits = 7)
-
-kableExtra::kbl(se, align = "c", caption ="Portfolio of yields with errors") %>%
-  kableExtra::kable_classic("striped", full_width = F)  %>%
-  kableExtra::row_spec(0, font_size = 14) %>%
-  kableExtra::footnote(general = " $se$ is the standard deviation of the portfolio of yields observed with errors.")
-```
-
-It is relevant to highlight that the script above makes use of the principal component weights provided by BR (2017).
-Such a matrix is simply a scaled-up version of the one provided by the function *pca_weights_one_country* of this package.
-Accordingly, despite the numerical differences on the weight matrices, both methods generate time series of spanned factors which are perfectly correlated.
-Another difference between the two approaches relates to the construction form of the log-likelihood function (llk): while in the BR (2017) code the llk is expressed in terms of a portfolio of yields, the *MultiATSM* package generates this same input directly as a function of observed yields (i.e.
-both procedures lead to equivalent llk up to the Jacobian term).
-
-### Candelon and Moura (2021)
-```{r, eval=FALSE}
-###############################################################################################################
-#################################### USER INPUTS ##############################################################
-###############################################################################################################
-# A) Load database data
-data("CM_Factors")
-data('CM_Factors_GVAR')
-data('CM_Trade')
-data('CM_Yields')
-
-# B) Decide on general model inputs
-ModelType <- "GVAR jointQ" # Set "GVAR jointQ" for the GVAR-ATSM and "JLL original" for the JLL-ATSM
-
-StationarityUnderQ <- 0
-
-Economies <- c("China", "Brazil", "Mexico", "Uruguay") 
-GlobalVar <- c("GBC", "CPI_OECD") # Global Variables
-DomVar <- c("Eco_Act", "Inflation") # Country-specific variables
-
-N <- 3 # Number of spanned factors per country
-
-OutputLabel <- "CM_2021" # label of the model for saving the file
-DataFrequency <- "Monthly" # Frequency of the data
-UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"
-
-# B.1) Decide on specific model inputs
-#################################### GVAR-based models #########################################################
-if (ModelType == 'GVAR jointQ'){
-# Transition matrix inputs:
-t_First <- "2006" # First year of the sample
-t_Last <-  "2019" # Last year of the sample
-W_type <- 'Sample Mean' # Method to compute the transition matrix
-# VARX input:
-VARXtype <- "unconstrained"
-}
-
-#################################### JLL-based models #########################################################
-if (ModelType =="JLL original"){
-    DominantCountry <- "China"
-}
-################################################################################################################
-
-# C) Decide on Settings for numerical outputs
-Horiz <- 25
-DesiredGraphs <- c("Fit", "GIRF", "GFEVD")
-WishGraphRiskFac <- 0
-WishGraphYields <- 1
-WishOrthoJLLgraphs <- 1
-
-# D) Bootstrao settings
-WishBootstrap <- 0 #  If bootstrap outputs are desired set this variable equal to 1 (it may take several hours/days).
-Bootlist <- list()
-Bootlist$methodBS <- 'bs'
-Bootlist$BlockLength <- c()
-Bootlist$ndraws <-  1000
-Bootlist$pctg   <-  95
-
-# E) Out-of-sample forecast
-WishForecast <- 0 #  If out-of-sample forecasts outputs are desired set this variable equal to 1
-ForecastList <- list()
-ForecastList$ForHoriz <- 12
-ForecastList$t0Sample <- 1
-ForecastList$t0Forecast <- 90
-
-###############################################################################################################
-############################### NO NEED TO MAKE CHANGES FROM HERE #############################################
-###############################################################################################################
-
-# 2) Minor preliminary work
-C <- length(Economies)
-FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType) # Generate the set of labels
-mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields) # Vector of maturities expressed in years
-
-ZZ <- RiskFactors
-# 2.1) Build the GVARinputs
-if (ModelType == "GVAR jointQ"){
-  GVARinputs <- list()
-  GVARinputs$Economies <- Economies
-  GVARinputs$GVARFactors <- FactorsGVAR
-  GVARinputs$VARXtype <- VARXtype
-  GVARinputs$Wgvar <- Transition_Matrix(t_First, t_Last, Economies, W_type, DataPath = NULL, Data = TradeFlows)
-  } else { GVARinputs <- NULL }
-
-# 2.2) Build the JLLinputs
-if (ModelType =="JLL original"){
-  JLLinputs <- list()
-  JLLinputs$Economies <- Economies
-  JLLinputs$DomUnit <- DominantCountry
-  JLLinputs$WishSigmas <- 1
-  JLLinputs$SigmaNonOrtho <- NULL
-  JLLinputs$JLLModelType <- ModelType
-}else{  JLLinputs <- NULL }
-
-# 3) Prepare the inputs of the likelihood function
-ModelParaList <- list()
-
-# 3.1) Compute the inputs that go directly into the log-likelihood function
-ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, JLLinputs,
-GVARinputs)
-
-  # 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
-K1XQ <- ATSMInputs$K1XQ
-if (ModelType == "JLL original"){ SSZ <- NULL} else{SSZ <- ATSMInputs$SSZ}
-
-# 4) Build the objective function
-f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)
-
-# 5) Set the optimization settings
-VarLab <- ParaLabels(ModelType, StationarityUnderQ)
-
-varargin <- list()
-varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]], NULL , NULL)
-varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
-varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
-varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
-varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
-varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
-varargin$OptRun <-  c("iter off")
-
-LabelVar<- c('Value', 'Label', 'LB', 'UB')
-for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}
-
-tol <- 1e-4
-
-# 6) Optimization of the model
-ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, JLLinputs, GVARinputs)$Summary
-
-# 7) Numerical and graphical outputs
-InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ, WishGraphYields,  WishGraphRiskFac, WishOrthoJLLgraphs, WishBootstrap, Bootlist, WishForecast, ForecastList)
-
-# A) Fit, IRF, FEVD, GIRF and GFEVD
-NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)
-
-# B) Bootstrap
-Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs, FactorLabels,
-DataFrequency, varargin, JLLinputs, GVARinputs)
-
-# C) Out-of-sample forecasting
-Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies, DataFrequency, JLLinputs, GVARinputs)
-```
-
 # References

### 2023-08-16 inst/doc/Paper-Replications.R

@@ -1,0 +1,428 @@
+## -----------------------------------------------------------------------------
+library(MultiATSM)
+
+## ---- eval=FALSE--------------------------------------------------------------
+#  #########################################################################################################
+#  #################################### USER INPUTS ########################################################
+#  #########################################################################################################
+#  # A) Load database data
+#  data("BR_jps_gro_R3")
+#  
+#  # B) Decide on general model inputs
+#  ModelType <- "JPS"
+#  
+#  StationarityUnderQ <- 0
+#  BiasCorrection <- 0
+#  WishForwardPremia <- 1
+#  
+#  
+#  Economies <- c("U.S.") # Names of the economies from the economic system
+#  GlobalVar <- c()
+#  DomVar <- c("GRO", "INF") # Country-specific variables
+#  
+#  N <- 3 # Number of spanned factors per country
+#  
+#  DataFrequency <- "Monthly" # Frequency of the data
+#  UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"
+#  
+#  mat <- c(0.25, 0.5, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10) # maturities expressed in years
+#  
+#  #########################################################################################################
+#  ############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
+#  #########################################################################################################
+#  # 2) Minor preliminary work
+#  C <- length(Economies)
+#  M <- length(DomVar)
+#  FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType) # Generate the set of labels
+#  Yields <- t(BR_jps_out$Y)
+#  W <- BR_jps_out$W[1:N,] # Use the Wpca matrix from BR (2017)
+#  SpaFac <- W%*%Yields
+#  rownames(SpaFac) <- FactorLabels$Tables$U.S.[-(1:M)]
[…]

### 2024-04-29 inst/doc/Paper-Replications.R
version 0.3.6

@@ -1,7 +1,7 @@
 ## -----------------------------------------------------------------------------
 library(MultiATSM)
 
-## ---- eval=FALSE--------------------------------------------------------------
+## ----eval=FALSE---------------------------------------------------------------
 #  #########################################################################################################
 #  #################################### USER INPUTS ########################################################
 #  #########################################################################################################
@@ -78,7 +78,7 @@ library(MultiATSM)
 #    ModelPara <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType)$Summary
 #  
 
-## ---- echo= FALSE-------------------------------------------------------------
+## ----echo= FALSE--------------------------------------------------------------
 options(scipen = 100) # eliminate the scientific notation
 data("BR_jps_gro_R3")
 data("JPSrep")
@@ -101,7 +101,7 @@ kableExtra::kbl(TableQ, align = "c", caption = "$Q$-dynamics parameters") %>%
   kableExtra::row_spec(0, font_size = 14) %>%
   kableExtra::footnote(general = " $\\lambda$'s are the eigenvalues from the risk-neutral feedback matrix and $r0$ is the long-run mean of the short rate under Q.")
 
-## ---- echo= FALSE-------------------------------------------------------------
+## ----echo= FALSE--------------------------------------------------------------
 data("BR_jps_gro_R3")
 data("JPSrep")
 
@@ -142,7 +142,7 @@ kableExtra::kbl(TableP, align = "c", caption = "$P$-dynamics parameters") %>%
   kableExtra::pack_rows("MultiATSM", 6, 10) %>%
   kableExtra::footnote(general = " $K0Z$ is the intercept and $K1Z$ is feedback matrix from the $P$-dynamics.")
 
-## ---- echo= FALSE-------------------------------------------------------------
+## ----echo= FALSE--------------------------------------------------------------
 data("BR_jps_gro_R3")
 data("JPSrep")
 
@@ -156,7 +156,7 @@ kableExtra::kbl(se, align = "c", caption ="Portfolio of yields with errors") %>%
[…]

### 2024-10-15 inst/doc/Guidelines.Rmd
version 1.0.0

---
title: "MultiATSM package - General Guidelines"
author: Rubens Moura
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    number_sections: true
    df_print: paged
    toc: yes
    toc_depth: 2
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{General Guidelines}
  \usepackage[utf8]{inputenc} 
  %\VignetteEngine{knitr::rmarkdown}
bibliography: references.bib
editor_options: 
  markdown: 
    wrap: 72
---

This document aims at providing general guidance on the use of the
MultiATSM package available at the [CRAN
repository](https://cran.r-project.org/).

```{r}
library(MultiATSM)
```

The *MultiATSM* package offers estimation routines and various outputs
for single and multi-country affine term structure models (ATSMs). All
the frameworks of this package are based on the unspanned economic risk
framework from **(JPS, 2014)**. In essence, these models assume the
absence of arbitrage opportunities, consider a linear state space
representation of the yield curve dynamics and offer a tractable
approach to simultaneously combine the traditional yield curve factors
(spanned factors) along with macroeconomic and financial variables
(unspanned factors).

The *MultiATSM* package currently supports the generation of outputs for
eight distinct classes of ATSMs. Specifically, the package accommodates
models estimated either on a *country-by-country* basis, following the
approach of JPS (2014), or *jointly* across all countries within the
economic system, as developed in **JLL (2015)** by
@JotikasthiraLeLundblad2015 and **CM (2024)** by @CandelonMoura2024.

Due to peculiar features of JPS-based specifications, an efficient
estimation of the parameters governing the risk-neutral ($Q$-measure)
and the physical ($P$-measure) probability measures can be carried out
rather independently. The only exception is the variance-covariance
matrix (sigma) term which is a common element to both the $P$ and the
$Q$ density functions. In all cases, the risk factor dynamics under the
$P$-measure follow a VAR(1) model of some sort. The table below
summarizes the general features of each one of the models available at
this package.

```{r ModFea, message=FALSE, echo=FALSE}
ModelLabels <- c("JPS original", "JPS global", "JPS multi", "GVAR single", "GVAR multi", 
                 "JLL original", "JLL No DomUnit", "JLL joint Sigma")

# Rows
Tab <- data.frame(matrix(nrow = length(ModelLabels), ncol = 0)) 
rownames(Tab) <- ModelLabels

# Empty columns
EmptyCol <- c("", "", "", "", "", "", "", "") 
Tab$EmptyCol0 <- EmptyCol
# P-dynamics + 2 empty spaces
Tab$PdynIndUnco <- c("x", "", "", "", "", "", "", "")
Tab$PdynIndCo <- c("", "", "", "", "", "", "", "")
Tab$PdynJointUnco <- c("", "x", "x", "", "", "", "", "")
Tab$PdynJointJLL <- c("", "", "", "", "", "x", "x", "x")
Tab$PdynJointGVAR <- c("", "", "", "x", "x", "", "", "")
Tab$EmptyCol1 <- EmptyCol
Tab$EmptyCol2 <- EmptyCol
# Q-dynamics + 2 empty spaces
Tab$QdynInd <- c("x", "x", "", "x", "", "", "", "")   
Tab$QdynJoint <- c("", "", "x", "", "x", "x", "x", "x") 
Tab$EmptyCol3 <- EmptyCol
Tab$EmptyCol4 <- EmptyCol
# Sigma + 2 empty spaces
Tab$Ponly <-  c("", "", "", "", "", "x", "x", "")
Tab$PandQ <- c("x", "x", "x", "x", "x", "", "", "x")
Tab$EmptyCol5 <- EmptyCol
Tab$EmptyCol6 <- EmptyCol
# Dominant Unit
Tab$DomUnit <- c("", "", "", "", "", "x", "", "x")

# Adjust column names
ColNames <- c("","","","","JLL", "GVAR", "", "", "", "", "", "", "","", "", "","")
colnames(Tab) <- ColNames

# Generate the table
suppressWarnings(library(magrittr))

kableExtra::kbl(Tab, align = "c", caption = "Model Features") %>%
  kableExtra::kable_classic("striped", full_width = F)  %>%
  kableExtra::row_spec(0, font_size = 10) %>%
  kableExtra::add_header_above(c(" "=2, "Unrestricted" = 1, "Restricted" = 1, "Unrestricted" = 1, "Restricted" = 2, " " = 11)) %>%
 kableExtra::add_header_above(c(" "=2, "Individual" = 2, "Joint" = 3, " "=2, "Individual" = 1, "Joint" = 1, " "=2, "P only" = 1, "P and Q" = 1, " " = 3))  %>%
  kableExtra::add_header_above(c( " "=2, "P-dynamics"= 5, " "=2, "Q-dynamics"= 2, " "=2, "Sigma matrix estimation" = 2, " "=2, "Dominant Country"=1), bold = T) %>%
kableExtra::pack_rows("Unrestricted VAR", 1, 3 , label_row_css = "background-color: #666; color: #fff;")  %>%
kableExtra::pack_rows("Restricted VAR (GVAR)", 4, 5, label_row_css = "background-color: #666; color: #fff;") %>%
kableExtra::pack_rows("Restricted VAR (JLL)", 6, 8, label_row_css = "background-color: #666; color: #fff;")
```

Some aspects of the multi-country frameworks are worth highlighting. As
for the models based on the setup of JLL (2015), the version "*JLL
original*" follows closely the seminal work of JLL (2015), i.e., it is
assumed an economic cohort containing a worldwide dominant economy and a
set of smaller countries, in addition to the estimation of the sigma
matrix be performed exclusively under the $P$-measure. The two other
alternative versions assume the absence of a dominant country ("*JLL No
DomUnit"*) and the estimation of sigma under both the $P$ and $Q$
measures ("*JLL joint Sigma*"), as in the standard single-country JPS
(2014) model. As for the remaining multi-country ATSMs, it is considered
that the dynamics of the risk factors under the $P$-measure evolves
according to a GVAR model. The version labeled *"GVAR multi"* is the one
presented in CM (2024).

In what follows, the focus will be devoted to describe the various
pieces that are necessary to implement ATSMs. Specifically, Section
\@ref(S:SectionData) describes the data-set available at this package
and the set of functions that are useful to retrieve data from Excel
files. Section \@ref(S:SectionInputs) exposes the necessary inputs that
have to be specified by the user. In Section \@ref(S:SectionEstimation)
the estimation procedure is detailed. Section
\@ref(S:SectionReplication) shows how to use the *MultiATSM* package to
estimate ATSMs from scratch.

# Data {#S:SectionData}

## Package data-set {#S:SectionDataPackage}

The *MultiATSM* package includes modified datasets compared to those
originally used in @CandelonMoura2023 (CM, 2023) and @CandelonMoura2024
(CM, 2024) due to limited availability of certain data sources. In some
cases, simulated data replicating the original methodology is employed.
To simplify the exposition, the database of CM (2024) is used throughout
the following sections to demonstrate the package's functions. These
datasets can be accessed using the `LoadData()` function with the
argument *CM_2024*, and similarly, the CM (2023) data with *CM_2023*.

```{r}
LoadData("CM_2024")
```

The previous function loads four types of datasets. The first dataset
consists of time series for zero-coupon bond yields from four emerging
markets: China, Brazil, Mexico, and Uruguay. For the purpose of the ATSM
estimation, it is important to note two requirements: *(i)* the bond
yield maturities must be consistent across all countries (although the
`DataForEstimation()` function can take different maturities as inputs,
it outputs bond yields with uniform maturities for all economies); and
*(ii)* the bond yields should be expressed in percentage terms per
annum, not in basis points. The *MultiATSM* package does not include
routines for bootstrapping zero-coupon yields from coupon bonds, so
users must handle this data manipulation themselves if needed.

```{r}
data('CM_Yields')
```

The second dataset focuses on the time series of risk factors. Using the
terminology of JPS (2014), this dataset includes *(i)* country-specific
**spanned factors** (commonly represented by level, slope, and curvature
-- see Section \@ref(S:SpaFac) for an intuitive explanation of these
factors) and *(ii)* a set of country-specific and global **unspanned
factors**, such as economic growth and inflation. Similar to the bond
yield data, these economic indicators must be constructed by the user.

```{r}
data('CM_Factors')
```

The final two datasets are essential for estimating GVAR-based models.
The trade flows data, sourced from the International Monetary Fund
(IMF), presents the sum of the value of all goods imports and exports
between any two countries of the sample on a yearly basis since 1948.
All the values are free on board and are expressed in U.S. dollars.
These data are used to construct the transition matrix of the GVARs
models.

```{r}
data('CM_Trade')
```

The GVAR factors database casts country-specific lists of all the
factors that are used in the estimation of each country's VARX(1,1,1).
The function titled `DatabasePrep()` provides assistance to structure
the data in a similar fashion to this form. Moreover, a specific
function for computing the star risk factors is detailed in the Section
\@ref(S:SectionGVAR).

```{r}
data('CM_Factors_GVAR')
```

## Import data from Excel files {#S:SectionImport}

This package also offers an automatized procedure to extract data from
Excel files and to, subsequentially, prepare the risk factor databases
that are directly used in the model estimation. The use of the package
functions requires that the databases *(i)* are constructed in separate
Excel files for the unspanned factors, the term structure data, and the
measures of interdependence (for the GVAR-based models); *(ii)* contain,
in each Excel file, one separate tab per country (in addition to the
global variables, in the case of the unspanned factors' database); and
*(iii)* have identical variable labels across all the tabs within each
file. For illustration, see the Excel file available at the package. One
example of list of inputs to be provided is, for instance,

```{r}
Initial_Date <- "2006-09-01" # Format "yyyy-mm-dd"
Final_Date <- "2019-01-01" # Format "yyyy-mm-dd"
DataFrequency <- "Monthly"
GlobalVar <- c("GBC", "VIX") # Global Variables
DomVar <- c("Eco_Act", "Inflation", "Com_Prices", "Exc_Rates") #  Domestic variables
N <-  3 # Number of spanned factors per country
Economies <- c("China", "Mexico", "Uruguay", "Brazil", "Russia")
ModelType <- "JPS original"
```

Using these inputs, the variable *RiskFac_TS* can be constructed,
containing the full time series of risk factors required for the model.

```{r}
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
RiskFac_TS <- DataForEstimation(Initial_Date, Final_Date, Economies, N, FactorLabels, ModelType,
                                DataFrequency)
```

# Required user inputs {#S:SectionInputs}

## Fundamental inputs required for estimation

In order to estimate any ATSM of this package, the user needs to specify
several general model inputs, namely:

**1. Desired ATSM class**

-   *Model type*: a character vector indicating the model type to be
    estimated, as presented in Table \@ref(tab:ModFea).

**2. Risk factor set**

-   *Economies*: a character vector containing the names of the
    economies included in the system;

-   *Global variables*: a character vector containing the labels of the
    desired global unspanned factors used in the model estimation;

-   *Domestic variables*: a character vector containing the labels of
    the desired domestic unspanned factors used in the model estimation;

-   *Number of spanned factors (N)*: a scalar that represents the number
    of country-specific spanned factors. Note that *N* is equal across
    countries.

**3. Sample span**

-   *Initial sample date*: Start date of the sample period in the format
    "dd-mm-yyyy";

-   *End sample date*: End date of the sample period in the format
    "dd-mm-yyyy".

**4. Frequency of the data**:

-   a character vector that specifies the frequency of the time series
    data. The available options are: *"Annually"*, *"Quarterly"*,
    *"Monthly"*, *"Weekly"*, *"Daily Business Days"*, and *"Daily All
    Days"*.

**5. Stationary constraint under the** $Q$-**dynamics**:

-   A binary variable that equals $1$ if the user wishes to impose the
    constraint that the largest eigenvalue under the $Q$-measure is
    strictly less than $1$ otherwise, it should be set to $0$. While
    enforcing this stationary constraint may prolong model estimation,
    it can improve the convergence of the model.

**6. Output label**

-   Character vector of length one, specifying the name used in the file
    that stores the model outputs.

One possible example of the basic model inputs is

```{r}
# 1) Model type
ModelType <- "JPS original"

# 2) Risk factor set
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
GlobalVar <- c("GBC", "CPI_OECD") # Global variables
DomVar <- c("Eco_Act", "Inflation") # Domestic variables 
N <- 3 # number of spanned factors per country

# 3) Sample span
Initial_Date <- "01-05-2005" # Format: "dd-mm-yyyy"
Final_Date <- "01-12-2019" # Format: "dd-mm-yyyy"

# 4) Frequency of the data
DataFrequency <- "Monthly"

# 5) Risk-neutral stationary constraint
StationarityUnderQ <- 0 # 1 = set stationary condition under the Q; 0 = no stationary condition

# 6) Output label
OutputLabel <- "Model_demo"
```

## Model-specific inputs required for estimation

### GVARlist and JLLlist {#S:SectionGVARJLLinputs}

Additional inputs are required for estimating GVAR-ATSM or JLL-ATSM
models. These inputs should be organized into separate lists, named
*GVARlist* and *JLLlist* for clarity. This section outlines the general
structure of both lists, while Sections \@ref(S:SectionGVAR) and
\@ref(S:SectionJLL) provide detailed explanations and available options.

For GVAR models, the nature of the inputs is twofold. First, the dynamic
structure of each country's VARX model must be specified. For example:

```{r}
VARXtype <- "unconstrained"
```

Second, the estimation of the GVAR requires specifying the necessary
inputs to build the `transition matrix()` . An example of this list of
inputs is

```{r}
W_type <- 'Sample Mean' # Method to compute the transition matrix
t_First_Wgvar <- "2000" # First year of the sample
t_Last_Wgvar <-  "2015" # Last year of the sample
```

Therefore, a complete *GVARlist* is

```{r}
GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean",
                  t_First_Wgvar = "2000", t_Last_Wgvar = "2015") 
```

Regarding the JLL-ATSM frameworks, it is sufficient to specify the name
of the dominant economy (if applicable) within the economic system:

```{r}
JLLlist <- list(DomUnit =  "China")
```

### BRWlist {#S:SectionBRWinputs}

In an influential paper, @BauerRudebuschWu2012 **(BRW, 2012)** have
shown that the estimates from traditional ATSMs can be severely biased
due to the typical small-sample size used in these studies. As result,
such models may produce unreasonably stable expected future short-term
interest, therefore distorting the term premia estimates for
long-maturity bonds. To circumvent this problem, they propose a
bootstrap-based method that relies on indirect inference estimation that
corrects for the bias present in previous works.

From its version $0.3.1$, the *MultiATSM* package accommodates the
framework proposed by BRW (2012) to each one of its ATSM types. Some
additional inputs must be specified, if the user intends to perform the
model estimation along the lines of BRW (2012). They are:

-   *Mean or median of OLS estimates* *(flag_mean)*: the user must
    decide whether to compute the mean or median of the OLS estimates
    after each bootstrap iteration. To compute the mean (median), the
    user must set *TRUE* (*FALSE*);

-   *Adjustment parameter* *(gamma)*: this parameter controls the degree
    of shrinkage between the difference in the OLS estimates and the BRW
    (2012)'s bootstrap-based one after each iteration. The value of this
    parameter is fixed and must lie in the interval $]0;1[$;

-   *Number of iteration* *(N_iter)* : total amount of iterations used
    in the stochastic approximation algorithm after burn-in;

-   *Number of bootstrap samples (B)*: quantity of simulated samples
    used in each burn-in or actual iteration;

-   *Perform closeness check (checkBRW)*: flag whether the user wishes
    to check the root mean square distance between the model estimates
    produced by the bias-correction method and those generated by OLS.
    Default is set to *TRUE*;

-   *Number of bootstrap samples used in the closeness check (B_check)*:
    default is set to 100,000;

-   *Number of burn-in iteration (N_burn)*: quantity of the iterations
    to be discarded in the first stage of the bias-correction estimation
    process. The recommended number is $15\%$ of the total number of
    iterations.

```{r}
BRWlist <- within(list(flag_mean = TRUE, gamma = 0.2, N_iter = 500, B = 50, 
                       checkBRW = TRUE, B_check = 1000), N_burn <- round(N_iter * 0.15))
```

## The `InputsForOpt()` function

The `InputsForOpt()` function plays a crucial role in the functioning of
the *MultiATSM* package. Specifically, it gathers, transforms, and
generates the necessary inputs for building the likelihood function
based on the user-provided inputs described in the previous sections,
which are then directly utilized in the model optimization process. One
possible use of `InputsForOpt()` is:

```{r,, eval=FALSE}
data(CM_GlobalMacroFactors)
data(CM_DomMacroFactors)
data(CM_Yields)

ModelType <- "JPS original"
Economies <- "Mexico"
t0 <- "01-05-2007" # Initial Sample Date (Format: "dd-mm-yyyy")
tF <- "01-12-2018" # Final Sample Date (Format: "dd-mm-yyyy")
N <- 3
GlobalVar <- c("Gl_Eco_Act") # Global Variables
DomVar <- c("Eco_Act") # Domestic Variables
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)

DataFreq <- "Monthly"

ATSMInputs <- InputsForOpt(t0, tF, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,
                         FactorLabels, Economies, DataFreq,CheckInputs = FALSE)
```

## Additional inputs for numerical and graphical outputs

Once the model parameters from the ATSM have been estimated, the
`NumOutputs()` function enables the compilation of numerical and
graphical of the following additional outputs:

-   Model fit of the bond yields;

-   Impulse response functions (IRFs);

-   Forecast error variance decompositions (FEVDs);

-   Generalized impulse response functions (GIRFs);

-   Generalized forecast error variance decompositions (GFEVDs);

-   Bond yields decomposition into their expected component and term
    premia.

For the numerical computation of the IRFs, GIRFs, FEVDs or GFEVDs, a
horizon of analysis has to be specified, e.g.:

```{r}
Horiz <- 100
```

For the graphical outputs, the user must indicate the desired types of
graphs in a string-based vector:

```{r}
DesiredGraphs <- c("Fit", "GIRF", "GFEVD") # Available options are: "Fit", "IRF", "FEVD", "GIRF", 
                                          # "GFEVD", "TermPremia".
```

Moreover, the user must select the types of variables of interest
(yields, risk factors or both) and, for the JLL type of models, whether
the orthogonalized version should be additionally included

```{r}
WishGraphRiskFac <- 0 #   YES: 1; No = 0.
WishGraphYields <- 1 #    YES: 1; No = 0.
WishOrthoJLLgraphs <- 0 # YES: 1; No = 0.
```

The function `InputsForOutputs()` can provide some guidance for
customizing the features of the wished outputs. Conditional to these
settings, individual folders are created at the user's *temporary
directory* (which can be accessed using `tempdir()` ) to store the
different types of the desired graphical outputs.

### Bond yield decomposition

In its current form, the *MultiATSM* package allows for the calculation
of two risk compensation measures: term premia and forward premia. In
particular, a $n$-maturity bond yield can be decomposed as the sum of
its expected ($Exp_t^{(n)}$) and term premia ($TP_t^{(n)}$) components.
Technically: $$
y_t^{(n)} = Exp_t^{(n)} + TP_t^{(n)}
$$ In the package's standard form, the expected short rate term is
computed from time $t$ to $t+n$, which represents the bond's maturity:
$Exp_t^{(n)} = \sum_{h=0}^{n} E_t[y_{t+h}^{(1)}]$. Alternatively, one
could perfom such a decomposition at the forward rate level
($f_t^{(n)}$), specifically
$f_t^{(n)} = \sum_{h=m}^{n} E_t[y_{t+h}^{(1)}] + FP_t^{(n)}$ where
$FP_t^{(n)}$ corresponds to the forward premia. In this case, the user
must specify a binary variable set to $1$ if the computation of forward
premia is desired, or $0$ otherwise. If set to $1$, the user must also
provide a two-element numerical vector containing the maturities
corresponding to the starting and ending dates of the bond maturity.
Example:

```{r}
    WishFPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0 
    FPmatLim <- c(60,120)
```

### Bootstrap settings

If the user intends to generate confidence bands around the point
estimates of the numerical outputs previously described using a
bootstrap procedure, an additional list of inputs is required.
Specifically:

-   The desired bootstrap procedure (*methodBS*): Available options are
    *(i)* standard residual bootstrap (*bs*); *(ii)* wild bootstrap
    (*wild*), and block bootstrap (*block*). If the latest is selected,
    then the block length (*BlockLength*) must be indicated;
-   The number of bootstrap draws (*ndraws*);
-   The confidence level expressed in percentage points (*pctg*).

```{r}
Bootlist <- list(methodBS = 'block', BlockLength = 4, ndraws =  50, pctg   =  95)
```

### Out-of-sample forecast settings

To perform out-of-sample forecasts of bond yields, the following
list-based features have to be detailed:

-   Number of periods-ahead that the forecasts are to be generated
    (*ForHoriz*);
-   Time-dimension index of the first observations belonging to the
    information set (*t0Sample*);
-   Time-dimension index of the last observation of the information set
    used to perform the first forecast set (*t0Forecast*);
-   Method used in the forecast computation (*ForType*): forecasts can
    be generated using rolling or expanding windows. Rolling window
    forecasting to be done by setting this parameter as *Rolling*. In
    such a case, the forecast length used in the forecasts is the one
    specified as in *t0Sample*. Should the user wishes the forecasts to
    be performed on a expanding window basis, then this input should be
    set as *Expanding*.

```{r}
ForecastList <- list(ForHoriz = 12, t0Sample = 1, t0Forecast = 70, ForType = "Rolling")
```

# Model Estimation {#S:SectionEstimation}

Once bond yields and economic time series data are gathered and model
features are selected, ATSM estimation proceeds in three steps. First,
estimate the country-specific spanned factors for inclusion in the
global ATSM. Next, estimate the parameters governing risk factor
dynamics under the $P-$measure. Finally, optimize the entire selected
ATSM, including the $Q-$measure parameters.

## Spanned Factors {#S:SpaFac}

The spanned factors are yield-related variables that are used to fit the
cross-section dimensions of the term structures. Typically, the spanned
factors of country $i$, $P_{i,t}$, are computed as the *N* first
principal components (PCs) of the set of observed bond yields. Formally,
$P_{i,t}$ is constructed as $P_{i,t} = w_i Y_{i,t}$ where $w_i$ is the
PC weight matrix and $Y_{i,t}$ is a country-specific column-vector of
yields with increasing maturities.

For *N=3*, the spanned factors are traditionally interpreted as level,
slope, and curvature. The nature of such interpretability results from
the features of the PC weight matrix as illustrated below:

```{r}
w <- pca_weights_one_country(Yields, Economy = "Uruguay") 
```

In matrix *w*, each row stores the weights used for constructing each
spanned factor. The entries of the first row are linked to the
composition of the *level* factor in that they load rather equally on
all yields. Accordingly, high (low) values of the level factor indicate
an overall high (low) value of yields across all maturities. In the
second row, the weights monotonically increase with the maturities and,
therefore, they capture the degree of steepness (*slope*) of the term
structure. High slope factor values imply a steep yield curve, whereas
low ones entail flat (or, possibly, downward) curves. In the third row,
the weights of the *curvature* factor are presented. The name of this
factor follows from the fact that the weights have a more pronounced
effect on the middle range maturities of the curve. These concepts are
graphically illustrated below.

```{r, fig.cap = "Yield loadings on the spanned factors", echo=FALSE}
LabSpaFac <- c("Level", "Slope", "Curvature")
N <- length(LabSpaFac)

w_pca <- data.frame(t(w[1:N,]))
colnames(w_pca) <- LabSpaFac
w_pca$mat <- c(0.25, 0.5, 1, 3, 5, 10) # vector of maturitie

# Prepare plots
colors <- c("Level" = "blue", "Slope" = "green", "Curvature" = "red")

g <-  ggplot2::ggplot(data = w_pca, ggplot2::aes(x=  mat)) +  
    ggplot2::geom_line(ggplot2::aes(y = Level, color = "Level"), linewidth = 0.7) + 
    ggplot2::geom_line(ggplot2::aes(y = Slope, color = "Slope"), linewidth = 0.7) +
    ggplot2::geom_line(ggplot2::aes(y = Curvature, color = "Curvature"),  linewidth = 0.7) +
      ggplot2::labs(color = "Legend") + ggplot2::scale_color_manual(values = colors) + ggplot2::theme_classic() +
    ggplot2::theme(axis.title.y= ggplot2::element_blank(), legend.position="top", legend.title=ggplot2::element_blank(), legend.text= ggplot2::element_text(size=8) ) + 
   ggplot2::xlab("Maturity (Years)") + ggplot2::geom_hline(yintercept=0)

print(g)
```

To directly obtain the time series of the country-specific spanned
factors, the user can simply use the `Spanned_Factors()` function as
follows:

```{r}
data('CM_Yields')
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
N <- 3
SpaFact <- Spanned_Factors(Yields, Economies, N)
```

## The P-dynamics estimation

As presented in Table \@ref(tab:ModFea), the dynamics of the risk
factors under the $P$-measure evolve according to a VAR(1) model, which
may be fully unrestricted (as in the case of the JPS-related models) or
partially restricted (as in the GVAR and JLL frameworks). While the
focus here is their integration into an ATSM setup, each specification
can also be estimated independently to study dynamic systems. The usage
of these models is illustrated below.

### VAR {#S:SectionVAR}

Using the `VAR()` function of this package requires simply selecting the
set of risk factors for the desired estimated model. For instance, for
the models *JPS global* and *JPS multi*, the estimation of the
$P$-dynamics parameters is obtained as

```{r}
data("CM_Factors")
PdynPara <- VAR(RiskFactors, VARtype= "unconstrained")
```

whereas the estimation of a *JPS original* model for China is

```{r}
FactorsChina <- RiskFactors[1:7,]
PdynPara <- VAR(FactorsChina, VARtype= "unconstrained")
```

In both cases, the outputs generated are the vector of intercepts and
both the feedback and variance-covariance matrices of an unconstrained
VAR model.

### GVAR {#S:SectionGVAR}

The `GVAR()` function is designed to estimate a first-order GVAR which
is built from country-specific VARX(1,1,1) models. For a comprehensive
discussion of GVAR models, see @ChudikPesaran2016.

The `GVAR()` function requires two key inputs: the number of
country-specific spanned factors (N) and a set of supplementary inputs
stored in the *GVARinputs* list. The former is composed by four
elements:

1.  *Economies*: A character vector containing the names of the
    economies included in the system;

2.  *GVAR list of risk factors*: a list of risk factors sorted by
    country in addition to the global variables. See the outputs of the
    `DatabasePrep()` function;

3.  *VARX type*: a string-vector containing the desired estimated form
    of the VARX(1,1,1). Three possibilities are available. The
    *"unconstrained"* form estimates the model without any constrains,
    using standard OLS regressions for each model equation. The
    *"constrained: Spanned Factors"* form prevents
    foreign-spanned-factors to impact any domestic factor in the
    feedback matrix, whereas *"constrained: '* followed by the name of
    the risk factor restricts this same factor to be influenced only by
    its own lagged values and the lagged values of its own star
    variables. In the last two cases, the VARX(1,1,1) is estimated by
    restricted least squares.

```{r}
GVARinputs <- list(Economies = Economies, GVARFactors = FactorsGVAR, VARXtype ="constrained: Inflation")
```

4.  *Transition matrix*: The estimation of a GVAR model requires
    defining a measure of interdependence among the countries of the
    economic system. This information is reported in the transition
    matrix, the entries of which reflect the degree of interconnection
    across two entities of this same economic system. The GVAR setups
    can be estimated using fixed or, starting from package's version
    $0.3.3$, time-varying interdependence weights.

Time-varying interdependence implies that the star factors are
constructed by applying trade flow weights from the specific year to
adjust the risk factors of the corresponding year. To employ this
feature, users need to specify the transition matrix type as
*Time-varying* and select the *same specific date (year)* for both the
initial and final years within the transition matrix. This implies that
the trade weights of this year are used for solving (link matrices) the
GVAR system.

Below the illustration of the transition matrix is computed as the
average of the cross-border trade flow weights for the period spanning
the years from 2006 to 2019. Therefore, the degree of dependence across
countries is assumed to be fixed over the entire sample span. Note that
each row sums up to $1$.

```{r}
data("CM_Trade")
t_First <- "2006"
t_Last <-  "2019"
Economies <- c("China", "Brazil", "Mexico", "Uruguay")
type <- "Sample Mean"
W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, DataPath = NULL, TradeFlows)
print(W_gvar)
```

Having defined the form of the transition matrix, one can complete the
*GVARinputs* list to estimate a GVAR(1) model. Note that the
*CheckInputs* parameter is set as *TRUE* to perform a prior consistency
check on the inputs provided in *GVARinputs*.

```{r}
data("CM_Factors_GVAR")

GVARinputs <- list(Economies = Economies, GVARFactors = FactorsGVAR, VARXtype = "unconstrained", 
                   Wgvar = W_gvar)
N <- 3

GVARpara <- GVAR(GVARinputs, N, CheckInputs = TRUE)
```

A separate routine is provided for computing the foreign-specific
factors (also commonly referred to as the star variables) used in the
estimation of the VARX models.

```{r}
data('CM_Factors')
StaFac <- StarFactors(RiskFactors, Economies, W_gvar)
```

### JLL {#S:SectionJLL}

Calculating the $P$-dynamics parameters in the form proposed by JLL
(2015) requires the following inputs: *(i)* the time series of the risk
factors in non-orthogonalized form; *(ii)* the number of
country-specific spanned factors and *(iii)* the specification of the
*JLLinputs* list, which includes the following elements:

1.  *Economies*: A character vector containing the names of the
    economies included in the system;

2.  *Dominant Country*: a string-vector containing the name of the
    economy which is assigned as the dominant country (applicable for
    the *JLL original* and *JLL joint Sigma* models) or *None*
    (applicable for the *JLL No DomUnit*);

3.  *Wish the estimation of the sigmas matrices*: this is a binary
    variable which assumes value equal to $1$ if the user wishes the
    estimation of all JLL sigma matrices (i.e. variance-covariance and
    the Cholesky factorization matrices) and, $0$ otherwise. Since this
    estimation is done numerically, its estimation can take several
    minutes;

4.  *Sigma of the non-orthogonalized variance-covariance matrix*: to
    save time, the user may provide the variance-covariance matrix from
    the non-orthogonalized dynamics. Otherwise, this input should be set
    as *NULL*.

5.  *JLL type*: a string-vector containing the label of the JLL model to
    be estimated according to the features described in Table
    \@ref(tab:ModFea).

One possible list of *JLLinputs* is as follows:

```{r}
ModelType <- "JLL original"   
JLLinputs <- list(Economies = Economies, DomUnit = "China", WishSigmas = 1,  SigmaNonOrtho = NULL,
                  JLLModelType = ModelType)
```

As such, the $P-$dynamics dimension of a JLL-based setup can be computed
as:

```{r, eval=FALSE}
data("CM_Factors")
N <- 3
JLLpara <- JLL(RiskFactors, N, JLLinputs, CheckInputs = TRUE)
```

## ATSM estimation

ATSM estimation involves optimizing the log-likelihood function by
optimally selecting the model parameters. The optimization structure in
this package is based on routines from the term structure package by
@LeSingleton2018. The ATSM otimization is carried out using the
`Optimization()` function.

Broadly, the unspanned risk framework from JPS (2014) (and, therefore,
all its multi-country extensions) requires the estimation parameter set
consisting of 6 elements, namely: the risk-neutral long-run mean of the
short rate (*r0*), the risk-neutral feedback matrix (*K1XQ*), the
variance-covariance matrix (*SSZ*) from the VAR processes, the standard
deviation of the errors from the portfolios of yields observed with
error (*se*), in addition to the intercept (*K0Z*) and the feedback
(*K1Z*) matrices of the physical dynamics.

The parameters *K0Z* and *K1Z* are estimated as part of the model's
$P-$dynamics and are known in closed form. Likewise, *r0* and *se* have
analytical solutions and can be factored out of the log-likelihood
function. However, the remaining parameters (*K1XQ* and *SSZ*) require
numerical estimation. @LeSingleton2018 provides standardized routines
(integrated into this package) to set effective initial values,
facilitating the optimization process.

The *MultiATSM* package offers two optimization algorithms: *fminunc*
and *fminsearch*. By default, both algorithms are run sequentially twice
to ensure the model reaches the global optimum. Specifically in the
default bootstrap setup, only *fminunc* is used to save time.

# Examples of full implementation of ATSMs {#S:SectionReplication}

This section presents some examples on how to use the *MultiATSM*
package to fully implement ATSMs. The console displays messages to help
the user track the progress of the code execution at each step. Further
on, the functions of this package are used to replicate some of the
results presented in the original papers of JPS (2014), CM (2023) and CM
(2024). See the [Paper Replication Vignette](Paper-Replications.html) of
this package for a detailed reproduction.

## General template

```{r, eval=FALSE}
########################################################################################################
#################################### USER INPUTS #######################################################
########################################################################################################
# A) Load database data
LoadData("CM_2024")

# B) GENERAL model inputs
ModelType <- "JPS multi" # available options: "JPS original", "JPS global", "GVAR single", "JPS multi",
                          #"GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".

Economies <- c("China", "Brazil", "Mexico") # Names of the economies from the economic system
GlobalVar <- c("Gl_Eco_Act")# Global Variables
DomVar <- c("Eco_Act", "Inflation") # Country-specific variables
N <- 2  # Number of spanned factors per country

t0_sample <- "01-05-2005" # Format: "dd-mm-yyyy"
tF_sample <- "01-12-2019" # Format: "dd-mm-yyyy"

OutputLabel <- "Test" # label of the model for saving the file
DataFreq <-"Monthly" # Frequency of the data

StatQ <- 0 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1,NO:0

# B.1) SPECIFIC model inputs
#################################### GVAR-based models ##################################################
GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2005",
                  t_Last_Wgvar = "2019")
# VARXtype: Available options "unconstrained" or "constrained" (VARX)
# W_type: Method to compute the transition matrix. Options:"Time-varying" or "Sample Mean"
# t_First_Wgvar: First year of the sample (transition matrix)
# t_Last_Wgvar:  Last year of the sample (transition matrix)
#################################### JLL-based models ###################################################
JLLlist <- list(DomUnit =  "China")
# DomUnit: name of the economy of the economic system, or "None" for the model "JLL No DomUnit"
###################################### BRW inputs  ######################################################
WishBC <- 0 # Wish to estimate the model with the bias correction method of BRW (2012): #YES: 1, NO:0
BRWlist <- within(list(flag_mean = TRUE, gamma = 0.05, N_iter = 250, B = 50, checkBRW = TRUE,
                       B_check = 1000),  N_burn <- round(N_iter * 0.15))
# flag_mean: TRUE = compute the mean; FALSE = compute the median
# gamma: Adjustment parameter
# N_iter:  Number of iteration to be conserved
# N_burn:  Number of iteration to be discarded
# B: Number of bootstrap samples
# checkBRW: wishes to perform closeness check: TRUE or FALSE
# B_check: If checkBRW is chosen as TRUE, then choose number of bootstrap samples used in the check

# C) Decide on Settings for numerical outputs
WishFPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0
FPmatLim <- c(60,120) #  If the forward premia is desired, then choose the Min and max maturities of the
                      # forward premia. Otherwise set NA
Horiz <- 30
DesiredGraphs <- c("Fit", "IRF", "TermPremia") # "Fit", "IRF", "FEVD", "GIRF", "GFEVD", "TermPremia"
WishGraphRiskFac <- 0
WishGraphYields <- 1
WishOrthoJLLgraphs <- 0

# D) Bootstrap settings
WishBootstrap <- 1 #  YES: 1; No = 0.
BootList <- list(methodBS = 'bs', BlockLength = 4, ndraws = 5, pctg =  95)
# methodBS: bootstrap method. Available options: (i) 'bs' ; (ii) 'wild'; (iii) 'block'
# BlockLength: Block length, necessary input for the block bootstrap method
# ndraws: number of bootstrap draws
# pctg: confidence level

# E) Out-of-sample forecast
WishForecast <- 1 #  YES: 1; No = 0.
ForecastList <- list(ForHoriz = 12,  t0Sample = 1, t0Forecast = 162, ForType = "Rolling")
# ForHoriz: forecast horizon
# t0Sample:   initial sample date
# t0Forecast:  last sample date for the first forecast
# ForType: Available options are "Rolling" or "Expanding"

#########################################################################################################
############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
#########################################################################################################

# 2) Minor preliminary work: get the sets of factor labels and  a vector of common maturities
FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
if(any(ModelType == c("GVAR single", "GVAR multi"))){ GVARlist$DataConnectedness <- TradeFlows}

# 3) Prepare the inputs of the likelihood function
ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,
                           FactorLabels, Economies, DataFreq, GVARlist, JLLlist, WishBC, BRWlist)

# 4) Optimization of the ATSM (Point Estimates)
ModelParaList <- Optimization(ATSMInputs, StatQ, DataFreq, FactorLabels, Economies, ModelType)

# 5) Numerical and graphical outputs
# a) Prepare list of inputs for graphs and numerical outputs
InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
                                    WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishFPremia,
                                    FPmatLim, WishBootstrap, BootList, WishForecast, ForecastList)

# b) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)

# c) Confidence intervals (bootstrap analysis)
BootstrapAnalysis <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, Economies, InputsForOutputs,
                               FactorLabels, JLLlist, GVARlist, WishBC, BRWlist)

# 6) Out-of-sample forecasting
Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
                            JLLlist, GVARlist, WishBC, BRWlist)
```

# References

### 2024-10-15 inst/doc/Paper-Replications.Rmd
version 1.0.0

@@
 ---
-title: "Paper Replication"
+title: "Paper Replications"
 author: Rubens Moura
@@
 vignette: >
-  %\VignetteIndexEntry{Vignette for package MultiATSM_v2}
+  %\VignetteIndexEntry{Paper Replications}
   %\VignetteEngine{knitr::rmarkdown}
@@
 economy.
-Subsequently, I present the multi-country model extension proposed by @CandelonMoura2021, employing a GVAR approach to capture risk factor dynamics.
+Subsequently, I present the multi-country model extension proposed by @CandelonMoura2024, employing a GVAR approach to capture risk factor dynamics.
 Lastly, I introduce the program designed to replicate results detailed in @CandelonMoura2023, an extension of the previous framework to investigate the term structure behavior of major emerging markets during the unprecedented COVID$-19$ pandemic.
+It is worth highlighting that, due to limited public availability of bond yield data, I employ, in certain instances, simulated data closely matching the original papers' construction method.
+Consequently, while the outputs generated by the scripts below may not exactly mirror those from the original papers, the main features of these results are preserved.
 
-A thorough explanation on the usage of the additional functionalities of the *MultiATSM* package is available at the [package vignette](my-vignette.html).
+A thorough explanation on the usage of the additional functionalities of the *MultiATSM* package is available at the [package vignette](Guidelines.html).
 
@@
 
-# Paper replications
-## Joslin, Priebisch and Singleton (2014)
-To the best of my knowledge, the original data-set used in @JoslinPriebschSingleton2014 is not publicly available.
-Nevertheless, there exists an alternative source of data that closely matches JPS (2014) in terms of the construction method and the sample period.
-This database is constructed by @BauerRudebusch2017 (BR, 2017) and is available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).
+# Joslin, Priebisch and Singleton (2014)
+
+The database used in this replication was constructed by @BauerRudebusch2017 (BR, 2017) and is available at [Bauer's website](https://www.michaeldbauer.com/publication/spanning-puzzle/).
 In that paper, BR (2017) investigate whether macro-finance term structure models better suit the unspanned macro risk framework by JPS (2014) or the earlier traditional spanned settings as the one by @AngPiazzesi2003JME.
@@
 Supported by BR (2017)'s data-set, the code below uses the *MultiATSM* package to estimate the key model parameters from the ATSM along the lines of JPS (2014).
-The model estimation is performed using the *JPS* modeling setup.
+The model estimation is performed using the *JPS original* modeling setup.
 
 ```{r, eval=FALSE}
-#########################################################################################################
-#################################### USER INPUTS ########################################################
-#########################################################################################################
 # A) Load database data
-data("BR_jps_gro_R3")
+LoadData("BR_2017")
 
-# B) Decide on general model inputs
-ModelType <- "JPS"
-
-StationarityUnderQ <- 0
-BiasCorrection <- 0
-WishForwardPremia <- 1
+# B) GENERAL model inputs
+ModelType <- "JPS original"
 
 Economies <- c("U.S.") # Names of the economies from the economic system
-GlobalVar <- c()
+GlobalVar <- c() # Global Variables
 DomVar <- c("GRO", "INF") # Country-specific variables
-
 N <- 3 # Number of spanned factors per country
 
-DataFrequency <- "Monthly" # Frequency of the data
-UnitMatYields <- "Month" # time-unit in which yields are expressed. Available options are "Month" or "Year"
+t0_sample <- "January-1985"
+tF_sample <- "December-2007"
 
-mat <- c(0.25, 0.5, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10) # maturities expressed in years
+DataFreq <- "Monthly" # Frequency of the data
+
+StatQ <- 0 # Stationary condition
 
@@
 # 2) Minor preliminary work
-C <- length(Economies)
-M <- length(DomVar)
-FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType) # Generate the set of labels
+FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+
 Yields <- t(BR_jps_out$Y)
-W <- BR_jps_out$W[1:N,] # Use the Wpca matrix from BR (2017)
-SpaFac <- W%*%Yields
-rownames(SpaFac) <- FactorLabels$Tables$U.S.[-(1:M)]
-ZZ <- rbind(t(BR_jps_out$M.o), SpaFac) # Complete set of risk factors
+DomesticMacroVar <- t(BR_jps_out$M.o)
+GlobalMacroVar <- c()
 
 # 3) Prepare the inputs of the likelihood function
-i <- length(Economies)
-  # 3.1) Compute the inputs that go directly into the log-likelihood function
-  ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency)
-  # 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
-  K1XQ <- ATSMInputs$K1XQ
-  SSZ <- ATSMInputs$SSZ
-  # 3.3) Adjust the inputs which are funtion of the W matrix
-  ATSMInputs$Wpca <- W
-  ATSMInputs$We <- t(pracma::null(W))
-  ATSMInputs$WpcaFull <- rbind(ATSMInputs$Wpca, ATSMInputs$We)
-  ATSMInputs$PP <- SpaFac
+ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,
+                           FactorLabels, Economies, DataFreq)
 
-  # 4) Build the objective function
-  f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)
-
-  # 5) Set the optimization settings
-  VarLab <- ParaLabels(ModelType, StationarityUnderQ)
-
-  varargin <- list()
-  varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
-  varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
-  varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
-  varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
-  varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
-  varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
-  varargin$OptRun <-  c("iter off")
-
-  LabelVar<- c('Value', 'Label', 'LB', 'UB') # Elements of each parameter
-  for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}
-
-  tol <- 1e-4
-
-  # 6) Optimization of the model
-  ModelPara <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType)$Summary
-
+# 4) Optimization of the model
+ModelPara <- Optimization(ATSMInputs, StatQ, DataFreq, FactorLabels, Economies, ModelType)
 ```
@@
 
-```{r, echo= FALSE}
+```{r PdynTab, echo= FALSE}
 data("BR_jps_gro_R3")
@@
 
-It is relevant to highlight that the script above makes use of the principal component weights provided by BR (2017).
-Such a matrix is simply a scaled-up version of the one provided by the function *pca_weights_one_country* of this package.
+For replicability, it is important to note that the $P-$dynamics results reported in Table \@ref(tab:PdynTab) for the *MultiATSM* package make use of the principal component weights provided by BR (2017).
+Such a matrix is simply a scaled-up version of the one provided by the function `pca_weights_one_country()` of this package.
 Accordingly, despite the numerical differences on the weight matrices, both methods generate time series of spanned factors which are perfectly correlated.
@@
 
-## Candelon and Moura (2021)
-The multi-country framework introduced in @CandelonMoura2021 was meticulously designed to enhance the tractability of large-scale ATSMs, providing a deeper understanding of the intricate global economic mechanisms that underlie yield curve fluctuations.
+# Candelon and Moura (forthcoming)
+
+The multi-country framework introduced in @CandelonMoura2024 was designed to enhance the tractability of large-scale ATSMs, providing a deeper understanding of the intricate global economic mechanisms that underlie yield curve fluctuations.
 This framework also reinforces the statistical validity of inference and enhances the predictive capabilities of these models.
-This novel setup, embodied by the *GVAR jointQ* model class, is benchmarked against the findings of @JotikasthiraLeLundblad2015, which are captured by the *JLL original* model class.
+This novel setup, embodied by the *GVAR multi* model class, is benchmarked against the findings of @JotikasthiraLeLundblad2015, which are captured by the *JLL original* model class.
 The paper showcases an empirical illustration involving China, Brazil, Mexico, and Uruguay.
@@
 ```{r, eval=FALSE}
-#########################################################################################################
-#################################### USER INPUTS ########################################################
-#########################################################################################################
 # A) Load database data
-data("CM_Factors")
-data('CM_Factors_GVAR')
-data('CM_Trade')
-data('CM_Yields')
+LoadData("CM_2024")
 
-# B) Decide on general model inputs
-ModelType <- "GVAR jointQ" # Options: "GVAR jointQ", "JLL original"
+# B) GENERAL model inputs
+ModelType <- "GVAR multi" # Options: "GVAR multi" or "JLL original".
 
-StationarityUnderQ <- 0  
-BiasCorrection <- 0 
-WishForwardPremia <- 0 
-FPmatLim <- c(47,48)
+Economies <- c("China", "Brazil", "Mexico", "Uruguay")
+GlobalVar <- c("Gl_Eco_Act", "Gl_Inflation")
+DomVar <- c("Eco_Act", "Inflation")
+N <- 3
 
-Economies <- c("China","Brazil","Mexico", "Uruguay") 
-GlobalVar <- c("GBC", "CPI_OECD") 
-DomVar <- c("OECD_Det", "Inflation")
+t0_sample <- "01-06-2004"
+tF_sample <- "01-01-2020"
 
-N <- 3 # Number of spanned factors per country
+OutputLabel <- "CM_jfec"
+DataFreq <-"Monthly"
 
-OutputLabel <- "CM_2021"
-DataFrequency <- "Monthly"
-UnitMatYields <- "Month"
+StatQ <- 0
 
-# B.1) Decide on specific model inputs
+# B.1) SPECIFIC model inputs
 #################################### GVAR-based models ##################################################
-t_First_Wgvar <- "2004"
-t_Last_Wgvar <-  "2019"
-W_type <- 'Sample Mean'
-VARXtype <- "unconstrained"
+GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2004",
+                  t_Last_Wgvar = "2019")
 #################################### JLL-based models ###################################################
-DomUnit <- "China"
-WishSigmas <- 0
-SigmaNonOrtho <- NULL
-JLLModelType <- ModelType
+JLLlist <- list(DomUnit =  "China")
 ###################################### BRW inputs  ######################################################
-flag_mean <- TRUE
-gamma <- 0.1
-N_iter <- 500
-N_burn <- N_iter*0.15
-B <- 50
-checkBRW <- 1
-B_check <- 1000
-#########################################################################################################
+WishBC <- 1
+BRWlist <- within(list(flag_mean = TRUE, gamma = 0.001, N_iter = 200, B = 50, checkBRW = TRUE,
+                       B_check = 1000),  N_burn <- round(N_iter * 0.15))
 
 # C) Decide on Settings for numerical outputs
+WishFPremia <- 1
+FPmatLim <- c(24,36)
+
 Horiz <- 25
-DesiredGraphs <- c("Fit", "GIRF", "GFEVD")
+DesiredGraphs <- c("GIRF", "GFEVD", "TermPremia")
 WishGraphRiskFac <- 0
@@
 # D) Bootstrap settings
-WishBootstrap <- 0 
-Bootlist <- list()
-Bootlist$methodBS <- 'bs' 
-Bootlist$BlockLength <- 4 
-Bootlist$ndraws <-  1000
-Bootlist$pctg   <-  95
+WishBootstrap <- 0 #  Set it to 1, if bootstrap is desired
+BootList <- list(methodBS = 'bs', BlockLength = 4, ndraws = 3, pctg =  95)
 
 # E) Out-of-sample forecast
-WishForecast <- 0 
-ForecastList <- list()
-ForecastList$ForHoriz <- 12 
-ForecastList$t0Sample <- 1 
-ForecastList$t0Forecast <- 90
+WishForecast <- 1
+ForecastList <- list(ForHoriz = 12,  t0Sample = 1, t0Forecast = 100, ForType = "Rolling")
 
@@
 
-# 2) Minor preliminary work
-C <- length(Economies)
-FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType)
-ZZ <- RiskFactors
-
-Data <- list()
-Data$GVARFactors <- FactorsGVAR
-mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields)
-
-# 2.1) Generate GVARinputs, JLLinputs and BRWinputs
-ModInputs <- ListModelInputs(ModelType, Data, Economies, VARXtype, t_First_Wgvar, t_Last_Wgvar, W_type,
-                             DomUnit, WishSigmas, SigmaNonOrtho, BiasCorrection, flag_mean, gamma, 
-                             N_iter, N_burn, B, checkBRW, B_check)
-
-GVARinputs <- ModInputs$GVARinputs
-JLLinputs <- ModInputs$JLLinputs
-BRWinputs <- ModInputs$BRWinputs
+# 2) Minor preliminary work: get the sets of factor labels and  a vector of common maturities
+FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+if(any(ModelType == c("GVAR single", "GVAR multi"))){ GVARlist$DataConnectedness <- TradeFlows}
 
 # 3) Prepare the inputs of the likelihood function
-ModelParaList <- list()
-for (i in 1:C){
-  if (( ModelType == "GVAR jointQ" || ModelType == "VAR jointQ" || ModelType == "JLL original"
-        || ModelType == "JLL NoDomUnit" || ModelType == "JLL jointSigma" )   & i >1 ){break}
+ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,
+                           FactorLabels, Economies, DataFreq, GVARlist, JLLlist, WishBC, BRWlist)
 
-# 3.1) Compute the inputs that go directly into the log-likelihood function
-ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, 
-                                  JLLinputs, GVARinputs, BRWinputs)
+# 4) Optimization of the ATSM (Point Estimates)
+ModelParaList <- Optimization(ATSMInputs, StatQ, DataFreq, FactorLabels, Economies, ModelType)
 
-# 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
-K1XQ <- ATSMInputs$K1XQ
-if (ModelType == "JLL original" || ModelType == "JLL NoDomUnit"){ SSZ <- NULL}else{SSZ <- ATSMInputs$SSZ}
+# 5) Numerical and graphical outputs
+# a) Prepare list of inputs for graphs and numerical outputs
+InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
+                                    WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishFPremia,
+                                    FPmatLim, WishBootstrap, BootList, WishForecast, ForecastList)
 
-# 4) Build the objective function
-f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)
-
-# 5) Set the optimization settings
-VarLab <- ParaLabels(ModelType, StationarityUnderQ)
-
-varargin <- list()
-varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
-varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
-varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
-varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
-varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
-varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
-varargin$OptRun <-  c("iter off")
-
-LabelVar<- c('Value', 'Label', 'LB', 'UB') 
-for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}
-
-tol <- 1e-4
-# 6) Optimization of the model
-ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType,
-                                           JLLinputs, GVARinputs)$Summary
-}
-
-# 7) Numerical Outputs
-InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,
-                                     UnitMatYields, WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, 
-                                     WishForwardPremia, FPmatLim, WishBootstrap, Bootlist, WishForecast,
-                                     ForecastList)
-# A) Fit, IRF, FEVD, GIRF, GFEVD and Risk premia
+# b) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
 NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)
 
-# B) Bootstrap
-Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs,
-                       FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs, BRWinputs)
+# c) Confidence intervals (bootstrap analysis)
+BootstrapAnalysis <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, Economies, InputsForOutputs,
+                               FactorLabels, JLLlist, GVARlist, WishBC, BRWlist)
 
-# C) Out-of-sample forecasting
+# 6) Out-of-sample forecasting
 Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
-                            DataFrequency, JLLinputs, GVARinputs, BRWinputs)
-
+                            JLLlist, GVARlist, WishBC, BRWlist)
 ```
 
-## Candelon and Moura (2023)
+# Candelon and Moura (2023)
+
 In this paper, @CandelonMoura2023 delve into an exploration of the underlying factors that shape the sovereign yield curves of Brazil, India, Mexico, and Russia in the midst of the COVID$-19$ pandemic crisis.
-To comprehensively address the intricate web of global macro-financial and particularly health-related interdependencies, the study employs a modeling approach based on a *GVAR jointQ* type framework.
+To comprehensively address the intricate web of global macro-financial and particularly health-related interdependencies, the study employs a modeling approach based on a *GVAR multi* type framework.
 By delving into these aspects, the authors aim to provide a nuanced understanding of the dynamics influencing sovereign yield curves in these countries during the challenging circumstances of the pandemic.
@@
 # A) Load database data
-data("CM_Factors_2023")
-data('CM_Factors_GVAR_2023')
-data('CM_Trade_2023')
-data('CM_Yields_2023')
+LoadData("CM_2023")
 
-# B) Decide on general model inputs
-ModelType <- "GVAR jointQ"
+# B) GENERAL model inputs
+ModelType <- "GVAR multi"
 
-StationarityUnderQ <- 0 
-BiasCorrection <- 0
+Economies <- c("Brazil", "India", "Russia", "Mexico")
+GlobalVar <- c("US_Output_growth", "China_Output_growth", "SP500")
+DomVar <- c("Inflation","Output_growth", "CDS", "COVID")
+N <- 2
 
-WishForwardPremia <- 1 
+t0_sample <- "22-03-2020"
+tF_sample <- "26-09-2021"
+
+OutputLabel <- "CM_EM"
+DataFreq <-"Weekly"
+
+StatQ <- 0
+
+# B.1) SPECIFIC model inputs
+#################################### GVAR-based models ##################################################
+GVARlist <- list( VARXtype = "constrained: COVID", W_type = "Sample Mean", t_First_Wgvar = "2015",
+                  t_Last_Wgvar = "2020")
+###################################### BRW inputs  ######################################################
+WishBC <- 0
+
+# C) Decide on Settings for numerical outputs
+WishFPremia <- 1
 FPmatLim <- c(47,48)
 
-Economies <- c("Brazil", "India", "Russia", "Mexico") # Names of the economies from the economic system
-GlobalVar <- c("US_Output_growth", "China_Output_growth", "SP500") # Global Variables
-DomVar <- c("Inflation","Output_growth", "CDS", "COVID") # Country-specific variables
-
-N <- 2 # Number of spanned factors per country
-
-OutputLabel <- "CM_2023" 
-DataFrequency <- "Weekly" 
-UnitMatYields <- "Month"
-
-# C.1) Decide on specific model inputs
-#################################### GVAR-based models ##################################################
-t_First_Wgvar <- "2015" 
-t_Last_Wgvar <-  "2020" 
-W_type <- 'Sample Mean' 
-VARXtype <- "constrained: COVID" 
-###################################### BRW inputs  ######################################################
-
-# D) Decide on Settings for numerical outputs
 Horiz <- 12
@@
 
-# E) Bootstrap settings
-WishBootstrap <- 0 #  YES: 1; No = 0.
-Bootlist <- list()
-Bootlist$methodBS <- 'bs' 
-Bootlist$BlockLength <- 4 # necessary input if one chooses the block bootstrap
-Bootlist$ndraws <-  100
-Bootlist$pctg   <-  95 # confidence level
+# D) Bootstrap settings
+WishBootstrap <- 1 #  YES: 1; No = 0.
+BootList <- list(methodBS = 'bs', BlockLength = 4, ndraws = 100, pctg =  95)
 
-# F) Out-of-sample forecast
-WishForecast <- 0 
-ForecastList <- list()
-ForecastList$ForHoriz <- 12 # forecast horizon
-ForecastList$t0Sample <- 1 # initial sample date
-ForecastList$t0Forecast <- 50 # last sample date for the first forecast
+#########################################################################################################
+############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
+#########################################################################################################
 
-##########################################################################################################
-############################### NO NEED TO MAKE CHANGES FROM HERE ########################################
-##########################################################################################################
-# 2) Minor preliminary work
-C <- length(Economies)
-FactorLabels <- LabFac(N, DomVar,GlobalVar, Economies, ModelType)
-ZZ <- RiskFactors
-
-Data <- list()
-Data$GVARFactors <- GVARFactors
-Data$Wgvar <-  Trade_Flows
-mat <- Maturities(Yields, Economies, UnitYields = UnitMatYields)
-
-# 2.1) Generate GVARinputs, JLLinputs and BRWinputs
-ModInputs <- ListModelInputs(ModelType, Data, Economies, VARXtype, t_First_Wgvar, t_Last_Wgvar, W_type)
-
-GVARinputs <- ModInputs$GVARinputs
-JLLinputs <- NULL
+# 2) Minor preliminary work: get the sets of factor labels and  a vector of common maturities
+FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+GVARlist$DataConnectedness <- Trade_Flows
 
 # 3) Prepare the inputs of the likelihood function
-# 3.1) Compute the inputs that go directly into the log-likelihood function
-ATSMInputs <- InputsForMLEdensity(ModelType, Yields, ZZ, FactorLabels, mat, Economies, DataFrequency, 
-                                  JLLinputs, GVARinputs)
+ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacro, DomMacro, FactorLabels,
+                           Economies, DataFreq, GVARlist)
 
-# 3.2) Initial guesses for Variables that will be concentrared out of from the log-likelihood function
-K1XQ <- ATSMInputs$K1XQ
-SSZ <- ATSMInputs$SSZ
+# 4) Optimization of the ATSM (Point Estimates)
+ModelParaList <- Optimization(ATSMInputs, StatQ, DataFreq, FactorLabels, Economies, ModelType)
 
-# 4) Build the objective function
-f <- Functionf(ATSMInputs, Economies, mat, DataFrequency, FactorLabels, ModelType)
+# 5) Numerical and graphical outputs
+# a) Prepare list of inputs for graphs and numerical outputs
+InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
+                                    WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishFPremia,
+                                    FPmatLim, WishBootstrap, BootList)
 
-# 5) Set the optimization settings
-VarLab <- ParaLabels(ModelType, StationarityUnderQ)
-
-varargin <- list()
-varargin$K1XQ <-list(K1XQ, VarLab[[ModelType]][["K1XQ"]] , NULL , NULL)
-varargin$SSZ <- list(SSZ, VarLab[[ModelType]][["SSZ"]], NULL, NULL)
-varargin$r0 <- list(NULL, VarLab[[ModelType]][["r0"]], NULL, NULL)
-varargin$se <- list(NULL, VarLab[[ModelType]][["se"]], 1e-6, NULL)
-varargin$K0Z <- list(NULL, VarLab[[ModelType]][["K0Z"]], NULL, NULL)
-varargin$K1Z <- list(NULL, VarLab[[ModelType]][["K1Z"]], NULL, NULL)
-varargin$OptRun <-  c("iter off")
-
-LabelVar<- c('Value', 'Label', 'LB', 'UB') # Elements of each parameter
-for (d in 1:(length(varargin)-1)){ names(varargin[[d]]) <-  LabelVar}
-
-tol <- 1e-4
-# 6) Optimization of the model
-ModelParaList <- list()
-ModelParaList[[ModelType]] <- Optimization(f, tol, varargin, FactorLabels, Economies, ModelType, 
-                                           JLLinputs, GVARinputs)$Summary
-
-# 7) Numerical Outputs
-InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StationarityUnderQ,
-                                     UnitMatYields, WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs,
-                                     WishForwardPremia, FPmatLim, WishBootstrap, Bootlist, WishForecast,
-                                     ForecastList)
-# A) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
+# b) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
 NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)
 
-# B) Bootstrap
-Bootstrap <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, mat, Economies, InputsForOutputs,
-                       FactorLabels, DataFrequency, varargin, JLLinputs, GVARinputs)
+# c) Confidence intervals (bootstrap analysis)
+BootstrapAnalysis <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, Economies, InputsForOutputs,
+                               FactorLabels, JLLlist = NULL, GVARlist)
+
 ```

### 2024-10-15 inst/doc/my-vignette.R

@@ -1,423 +1,0 @@
-## -----------------------------------------------------------------------------
-library(MultiATSM)
-
-## ----ModFea, message=FALSE, echo=FALSE----------------------------------------
-ModelLabels <- c("JPS", "JPS jointP", "VAR jointQ", "GVAR sepQ", "GVAR jointQ", 
-                 "JLL original", "JLL NoDomUnit", "JLL jointSigma")
-
-# Rows
-Tab <- data.frame(matrix(nrow = length(ModelLabels), ncol = 0)) 
-rownames(Tab) <- ModelLabels
-
-# Empty columns
-EmptyCol <- c("", "", "", "", "", "", "", "") 
-Tab$EmptyCol0 <- EmptyCol
-# P-dynamics + 2 empty spaces
-Tab$PdynIndUnco <- c("x", "", "", "", "", "", "", "")
-Tab$PdynIndCo <- c("", "", "", "", "", "", "", "")
-Tab$PdynJointUnco <- c("", "x", "x", "", "", "", "", "")
-Tab$PdynJointJLL <- c("", "", "", "", "", "x", "x", "x")
-Tab$PdynJointGVAR <- c("", "", "", "x", "x", "", "", "")
-Tab$EmptyCol1 <- EmptyCol
-Tab$EmptyCol2 <- EmptyCol
-# Q-dynamics + 2 empty spaces
-Tab$QdynInd <- c("x", "x", "", "x", "", "", "", "")   
-Tab$QdynJoint <- c("", "", "x", "", "x", "x", "x", "x") 
-Tab$EmptyCol3 <- EmptyCol
-Tab$EmptyCol4 <- EmptyCol
-# Sigma + 2 empty spaces
-Tab$Ponly <-  c("", "", "", "", "", "x", "x", "")
-Tab$PandQ <- c("x", "x", "x", "x", "x", "", "", "x")
-Tab$EmptyCol5 <- EmptyCol
-Tab$EmptyCol6 <- EmptyCol
-# Dominant Unit
-Tab$DomUnit <- c("", "", "", "", "", "x", "", "x")
-
-# Adjust column names
-ColNames <- c("","","","","JLL", "GVAR", "", "", "", "", "", "", "","", "", "","")
-colnames(Tab) <- ColNames
-
-
-# Generate the table
-suppressWarnings(library(magrittr))
-
-
-kableExtra::kbl(Tab, align = "c", caption = "Model Features") %>%
[…]

### 2025-03-24 inst/doc/Guidelines.Rmd
version 1.3.0

@@
 the frameworks of this package are based on the unspanned economic risk
-framework from **(JPS, 2014)**. In essence, these models assume the
-absence of arbitrage opportunities, consider a linear state space
-representation of the yield curve dynamics and offer a tractable
-approach to simultaneously combine the traditional yield curve factors
-(spanned factors) along with macroeconomic and financial variables
+framework from @JoslinPriebschSingleton2014  **(JPS, 2014)**. In essence, 
+these models assume the absence of arbitrage opportunities, consider a 
+linear state space representation of the yield curve dynamics and offer 
+a tractable approach to simultaneously combine the traditional yield curve 
+factors (spanned factors) along with macroeconomic variables
 (unspanned factors).
@@
 originally used in @CandelonMoura2023 (CM, 2023) and @CandelonMoura2024
-(CM, 2024) due to limited availability of certain data sources. In some
-cases, simulated data replicating the original methodology is employed.
-To simplify the exposition, the database of CM (2024) is used throughout
-the following sections to demonstrate the package's functions. These
-datasets can be accessed using the `LoadData()` function with the
+(CM, 2024) due to limited availability of certain data sources. As such, 
+in some cases, simulated data replicating the original methodology is 
+employed. To simplify the exposition, the database of CM (2024) is used 
+throughout the following sections to demonstrate the package's functions.
+These datasets can be accessed using the `LoadData()` function with the
 argument *CM_2024*, and similarly, the CM (2023) data with *CM_2023*.
@@
 factors) and *(ii)* a set of country-specific and global **unspanned
-factors**, such as economic growth and inflation. Similar to the bond
-yield data, these economic indicators must be constructed by the user.
+factors**, namely macroeconomic variables such as economic growth and 
+inflation. Similar to the bond yield data, these economic indicators 
+must be constructed by the user.
 
@@
 In order to estimate any ATSM of this package, the user needs to specify
-several general model inputs, namely:
+several general inputs, namely:
 
@@
 t_Last_Wgvar <-  "2015" # Last year of the sample
+DataConnectedness <- TradeFlows # Measure of connectedness across countries
 ```
@@
 ```{r}
-GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean",
-                  t_First_Wgvar = "2000", t_Last_Wgvar = "2015") 
+GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2000",
+                   t_Last_Wgvar = "2015", DataConnectedness = TradeFlows) 
 ```
@@
 
-```{r,, eval=FALSE}
-data(CM_GlobalMacroFactors)
-data(CM_DomMacroFactors)
-data(CM_Yields)
+```{r, eval=FALSE} 
+LoadData("CM_2024")
 
@@
 ATSMInputs <- InputsForOpt(t0, tF, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,  
-                         FactorLabels, Economies, DataFreq,CheckInputs = FALSE)
+                           FactorLabels, Economies, DataFreq, CheckInputs = FALSE)
 ```
@@
 particular, a $n$-maturity bond yield can be decomposed as the sum of
-its expected ($Exp_t^{(n)}$) and term premia ($TP_t^{(n)}$) components.
+its expected short-rate ($Exp_t^{(n)}$) and term premia 
+($TP_t^{(n)}$) components.
 Technically: $$
@@
     WishFPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0 
-    FPmatLim <- c(60,120)
+    FPmatLim <- c(60, 120)
 ```
@@
 
-If the user intends to generate confidence bands around the point
-estimates of the numerical outputs previously described using a
-bootstrap procedure, an additional list of inputs is required.
-Specifically:
+An additional list of inputs is required if the user intends to 
+generate confidence bands around the point estimates of the numerical 
+outputs. These results are achieved using a bootstrap procedure.
+Specifically, the list of inputs include:
 
@@
 
-Once bond yields and economic time series data are gathered and model
-features are selected, ATSM estimation proceeds in three steps. First,
-estimate the country-specific spanned factors for inclusion in the
-global ATSM. Next, estimate the parameters governing risk factor
-dynamics under the $P-$measure. Finally, optimize the entire selected
-ATSM, including the $Q-$measure parameters.
+Once bond yields and other (macro)economic time series data are gathered 
+and model features are selected, ATSM estimation proceeds in three steps. 
+First, the package estimates the country-specific spanned factors for 
+inclusion in the global ATSM. Next, it estimates the parameters governing 
+risk factor dynamics under the $P-$measure. Finally, it optimizes the 
+entire selected ATSM, including the $Q-$measure parameters.
 
@@
 
-For *N=3*, the spanned factors are traditionally interpreted as level,
-slope, and curvature. The nature of such interpretability results from
-the features of the PC weight matrix as illustrated below:
+For $N = 3$, the spanned factors are traditionally interpreted as level,
+slope, and curvature.  This interpretation arises from the structure of 
+the principal component (PC) weight matrix, as shown below :
 
@@
 To directly obtain the time series of the country-specific spanned
-factors, the user can simply use the `Spanned_Factors()` function as
+factors, the user can use the `Spanned_Factors()` function as
 follows:
@@
 ```{r}
-FactorsChina <- RiskFactors[1:7,]
+FactorsChina <- RiskFactors[1:7, ]
 PdynPara <- VAR(FactorsChina, VARtype= "unconstrained")
@@
 is built from country-specific VARX(1,1,1) models. For a comprehensive
-discussion of GVAR models, see @ChudikPesaran2016.
+discussion of GVAR models, see, for instance, @ChudikPesaran2016.
 
@@
 function. However, the remaining parameters (*K1XQ* and *SSZ*) require
-numerical estimation. @LeSingleton2018 provides standardized routines
+numerical estimation. @LeSingleton2018 provide standardized routines
 (integrated into this package) to set effective initial values,
@@
 
-## General template
+## General template {#S:SectionTemplate}
 
@@
 GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2005",
-                  t_Last_Wgvar = "2019")
+                  t_Last_Wgvar = "2019", DataConnectedness <- TradeFlows ) 
 # VARXtype: Available options "unconstrained" or "constrained" (VARX)
@@
 # t_Last_Wgvar:  Last year of the sample (transition matrix)
+# DataConnectedness: measure of connectedness across countries
 #################################### JLL-based models ###################################################
@@
 FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
-if(any(ModelType == c("GVAR single", "GVAR multi"))){ GVARlist$DataConnectedness <- TradeFlows}
 
@@
 
+## User-Friendly Inspection of Key Elements in the Selected ATSM Specification
+
+Starting from version $1.3.0$, the *MultiATSM* package provides additional ways to concisely inspect key elements of the selected ATSM.
+In particular, users can now gain an alternative perspective on the general model inputs, optimized model parameters, and root mean square error (RMSE) from the forecast analysis.
+For clarity, the variable names used hereafter align with those presented in Section \@ref(S:SectionTemplate).
+
+### ATSM inputs
+The *ATSMInputs* variable is an object of class *ATSMModelInputs* that includes the $S3$ methods `print()` and `summary()`.
+The `print()` method provides an overview of both the model inputs and the economic system's general features, while the `summary()` method offers statistical summaries of the risk factors and bond yields.
+
+```{r}
+data("Out_Example")
+print(Out$ATSMInputs)
+```
+
+```{r}
+summary(Out$ATSMInputs) 
+```
+
+### ATSM parameters
+The `Optimization()` function returns an object of class *'ATSMModelOutputs'*, which includes the `summary()` $S3$ method. This method offers a parsimonious summary of key model outputs following 
+optimization process, including the maximum log-likelihood value, the intercept and slope coefficients from the model’s cross-sectional dimension, and the count of physical-measure eigenvalues greater than one—an indicator of potential explosive behavior in the model’s time series dimension.
+
+```{r}
+summary(Out$ModelParaList) 
+```
+
+### Forecasts
+The `ForecastYields()` function returns an object of class *'ATSMModelForecast'*, which can be visualized using the `plot()` $S3$ method.
+This method generates a graphical representation of the RMSEs across different countries and forecast horizons.
+
+```{r}
+plot(Out$Forecasts) 
+```
+
 # References

### 2025-03-24 R/Bootstrap.R

@@ -1,180 +1,177 @@
 #' Generates the bootstrap-related outputs
 #'
-#'@param ModelType   A character vector indicating the model type to be estimated.
-#'@param ModelParaPE A list containing the point estimates of the model parameters. For details, refer to the outputs from the \code{\link{Optimization}} function.
-#'@param NumOutPE    The point estimate derived from numerical outputs. See the outputs from the \code{\link{NumOutputs}} function for further information.
-#'@param Economies   A character vector containing the names of the economies included in the system.
-#'@param InputsForOutputs A list containing the necessary inputs for generating IRFs, GIRFs, FEVDs, GFEVDs and Term Premia.
-#'@param FactorLabels A list of character vectors with labels for all variables in the model.
-#'@param JLLlist List. Inputs for JLL model estimation (see \code{\link{JLL}} function). Default is NULL.
-#'@param GVARlist List. Inputs for GVAR model estimation (see \code{\link{GVAR}} function). Default is NULL.
-#'@param WishBC  Whether to estimate the physical parameter model with bias correction, based on the method by Bauer, Rudebusch and Wu (2012) (see \code{\link{Bias_Correc_VAR}} function). Default is set to 0.
-#'@param BRWlist  List of necessary inputs for performing the bias-corrected estimation (see \code{\link{Bias_Correc_VAR}} function).
+#' @param ModelType A character vector indicating the model type to be estimated.
+#' @param ModelParaPE A list containing the point estimates of the model parameters. For details, refer to the outputs from the \code{\link{Optimization}} function.
+#' @param NumOutPE The point estimate derived from numerical outputs. See the outputs from the \code{\link{NumOutputs}} function for further information.
+#' @param Economies A character vector containing the names of the economies included in the system.
+#' @param InputsForOutputs A list containing the necessary inputs for generating IRFs, GIRFs, FEVDs, GFEVDs and Term Premia.
+#' @param FactorLabels A list of character vectors with labels for all variables in the model.
[…]

### 2025-03-24 R/Buildx_xvec.R

@@ -19,24 +19,21 @@ Build_xvec <- function(ListInputSet, Economies, FactorLabels, JLLinputs = NULL){
   # 0) Initializing the outputs
   x <- c()
   Dim_x <- c()
-  Name_x <- c()
-
+  Name_x <- vector("character", length(ListInputSet) - 1)
 
   # 1) Build list of interest
-  for (i in 1:(length(ListInputSet) -1)){
-    si <- strfind(ListInputSet[[i]]$Label, ':')
-    Name_x[i] <- Remove_at( substr(ListInputSet[[i]]$Label, start = 1, stop = si-1) ) # remove @ from variables name, if exists
+  for (i in seq_len(length(ListInputSet) - 1)){
+    si <- unlist(gregexpr(":", ListInputSet[[i]]$Label))
+    Name_x[i] <- sub(".*@", "", substr(ListInputSet[[i]]$Label, start = 1, stop = si - 1)) # remove @ from variables name, if exists
 
     Temp_x <- GetAuxPara(ListInputSet[[i]]$Value, ListInputSet[[i]]$Label, ListInputSet[[i]]$LB,
-                         ub= ListInputSet[[i]]$UB, Economies, FactorLabels, JLLinputs)
+                         ub = ListInputSet[[i]]$UB, Economies, FactorLabels, JLLinputs)
 
-
-    x <- rbind(x, t(t(as.vector(Temp_x))) )
-    Dim_x <- rbind(Dim_x, dim(Temp_x) )
+    x <- rbind(x, t(t(as.vector(Temp_x))))
+    Dim_x <- rbind(Dim_x, dim(Temp_x))
   }
 
   Outputs <- list(x0 = x, Dim_x = Dim_x, Name_x = Name_x)
-
   return(Outputs)
 }
 
@@ -64,8 +61,6 @@ Build_xvec <- function(ListInputSet, Economies, FactorLabels, JLLinputs = NULL){
 #' @param FactorLabels   string-list based which contains the labels of all the variables present in the model
 #' @param JLLinputs     list of necessary inputs for the estimation of JLL-based models (see "JLL" function)
 #'
-#' @importFrom pracma tril
-#'
 #'
 #'@keywords internal
 #'
@@ -77,61 +72,31 @@ Build_xvec <- function(ListInputSet, Economies, FactorLabels, JLLinputs = NULL){
 #'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
 #'  Available at: https://cepr.org/40029
 
-
-
[…]

### 2025-03-24 R/Classes_Methods_All.R

@@ -1,0 +1,271 @@
+##########################################################################################################################
+#' Print method for ATSMModelInputs objects
+#' @param x An object of class 'ATSMModelInputs'
+#' @param ... Additional arguments (not used)
+#'
+#' @method print ATSMModelInputs
+#' @usage \method{print}{ATSMModelInputs}(x, ...)
+#'
+#' @export
+
+print.ATSMModelInputs <- function(x, ...) {
+
+  info <- attr(x, "ModelInfo")
+
+  cat("General Model Inputs: \n")
+  cat("------------------------------------------\n")
+  cat("Model Type:", info$ModelType, "\n")
+  cat("Economic System:", paste(info$Economies, collapse = ", "), "\n")
+  cat("Sample Span:", info$InitialSampleDate, "-", info$FinalSampleDate, "\n")
+  cat("Data Frequency:", info$DataFrequency, "\n")
+  cat("Common Maturities across Countries:", info$Maturities, "years \n")
+  cat("------------------------------------------\n\n")
+
+  cat("Key Structure of the Economic System: \n")
+  cat("------------------------------------------\n")
+  cat("Total amount of spanned factors in the system:", info$NC, "\n")
+  cat("Total amount of global unspanned factors in the system:", info$G, "\n")
+  cat("Total amount of country-specific unspanned factors in the system:", info$MC, "\n")
+  cat("Total amount of risk factors in the system:", info$NC + info$G + info$MC, "\n")
+  cat("------------------------------------------\n")
+
+  invisible(x)
+}
+##########################################################################################################################
+#' Summary method for ATSMModelInputs objects
+#' @param object An object of class 'ATSMModelInputs'
+#' @param ... Additional arguments (not used)
+#'
+#' @method summary ATSMModelInputs
+#' @usage \method{summary}{ATSMModelInputs}(object, ...)
+#' @export
+
+summary.ATSMModelInputs <- function(object, ...) {
+
+  info <- attr(object, "ModelInfo")
+
+  # Function to print the summary statistics
[…]

### 2025-03-24 R/ForecastYields.R

@@ -1,52 +1,61 @@
 #' Generates forecasts of bond yields for all model types
 #'
-#'@param ModelType A character vector indicating the model type to be estimated.
-#'@param ModelPara  A list containing the point estimates of the model parameters. For details, refer to the outputs from the \code{\link{Optimization}} function.
-#'@param InputsForOutputs  A list containing the necessary inputs for generating IRFs, GIRFs, FEVDs, GFEVDs and Term Premia.
-#'@param FactorLabels  A list of character vectors with labels for all variables in the model.
-#'@param Economies A character vector containing the names of the economies included in the system.
-#'@param JLLlist A list of necessary inputs for the estimation of JLL-based models (see the \code{\link{JLL}} function).
-#'@param GVARlist A list containing the necessary inputs for the estimation of GVAR-based models (see the \code{\link{GVAR}} function).
-#'@param WishBRW Whether to estimate the physical parameter model with bias correction, based on the method by Bauer, Rudebusch and Wu (2012) (see \code{\link{Bias_Correc_VAR}} function). Default is set to 0.
-#'@param BRWlist  List of necessary inputs for performing the bias-corrected estimation (see \code{\link{Bias_Correc_VAR}} function).
+#' @param ModelType A character vector indicating the model type to be estimated.
+#' @param ModelPara A list containing the point estimates of the model parameters. For details, refer to the outputs from the \code{\link{Optimization}} function.
+#' @param InputsForOutputs A list containing the necessary inputs for generating IRFs, GIRFs, FEVDs, GFEVDs and Term Premia.
+#' @param FactorLabels A list of character vectors with labels for all variables in the model.
+#' @param Economies A character vector containing the names of the economies included in the system.
+#' @param JLLlist A list of necessary inputs for the estimation of JLL-based models (see the \code{\link{JLL}} function).
+#' @param GVARlist A list containing the necessary inputs for the estimation of GVAR-based models (see the \code{\link{GVAR}} function).
+#' @param WishBRW Whether to estimate the physical parameter model with bias correction, based on the method by Bauer, Rudebusch and Wu (2012) (see \code{\link{Bias_Correc_VAR}} function). Default is set to 0.
[…]

### 2025-03-24 R/Functionf_vectorized.R

@@ -23,52 +23,50 @@
 #'  Available at: https://cepr.org/40029
 #'
 
-
 Functionf_vectorized <- function(x, sizex, f, con, ListInputSet, ModelType, FactorLabels, Economies, JLLinputs,
                                  GVARinputs, WithEstimation){
 
-
   FF <- 1e9; out <- NULL; ParaList_Upd <- NULL
 
-  if ( any(is.na(x)==0) & any(is.infinite(x) ==0) & any(Im(x)==0) ) {
+  if (all(!is.na(x)) & all(!is.infinite(x)) & all(Im(x) == 0)) {
+    ParaList_Upd <- Update_ParaList(x, sizex, con, FactorLabels, Economies, JLLinputs, GVARinputs, ListInputSet)
+    # extract only the numerical parameters of each variable that will not be concentrated out of the likelihood function.
+    Para_Upds <- lapply(ParaList_Upd[-length(ParaList_Upd)], function(x) x$Value)
 
-  ParaList_Upd <- Update_ParaList(x, sizex, con, FactorLabels, Economies, JLLinputs, GVARinputs, ListInputSet)
-  Para_Upds <- getpara(ParaList_Upd) # extract only the numerical parameters of each variable that will not be concentrated out of the likelihood function.
-
-# Gather the estimated outputs after the optimization
-  if (!WithEstimation){
-
-      if (any(ModelType == c( "JLL original", "JLL No DomUnit"))){
-        FF <- unlist(f(K1XQ= Para_Upds[[1]],nargout = 2)$llk)
-        out <- f(K1XQ= Para_Upds[[1]], nargout = 2)
-      } else{
-        FF <- unlist(f(K1XQ= Para_Upds[[1]], SSZ=Para_Upds[[2]], nargout = 2)$llk)
-        out <- f(K1XQ= Para_Upds[[1]], SSZ = Para_Upds[[2]], nargout = 2)
+    # Gather the estimated outputs after the optimization
+    if (!WithEstimation){
+      if (any(ModelType == c("JLL original", "JLL No DomUnit"))){
+        FF <- unlist(f(K1XQ = Para_Upds[[1]], nargout = 2)$llk)
+        out <- f(K1XQ = Para_Upds[[1]], nargout = 2)
+      } else {
+        FF <- unlist(f(K1XQ = Para_Upds[[1]], SSZ = Para_Upds[[2]], nargout = 2)$llk)
[…]

### 2025-03-24 R/GVAR.R

@@ -1,7 +1,6 @@
 #' Estimates a GVAR(1) and a VARX(1,1,1) models
 #'
-#
-#'@param GVARinputs GVARinputs List. Inputs for GVAR model estimation:
+#' @param GVARinputs List of inputs for GVAR model estimation:
 #' \enumerate{
 #'        \item  \code{Economies}:  A character vector containing the names of the economies included in the system.
 #'        \item  \code{GVARFactors}: A list of all variables used in the estimation of the VARX model \cr
@@ -19,31 +18,28 @@
 #'          \item \code{Wgvar}: The GVAR transition matrix (C x C) used in the model solution. \cr
 #'                                (See the output from the \code{\link{Transition_Matrix}} function.).
 #' }
-#'@param N    Integer. Number of country-specific spanned factors.
-#'@param CheckInputs A logical flag to indicate whether to perform a prior consistency check on the inputs provided in \code{GVARinputs}. The default is set to FALSE
+#' @param N Integer. Number of country-specific spanned factors.
+#' @param CheckInputs A logical flag to indicate whether to perform a prior consistency check on the inputs provided in \code{GVARinputs}. The default is set to FALSE
 #'
-#'
-#'
-#'@return   A list containing
-#'\enumerate{
-#'\item parameters of the country-specific VARX(1,1,1)
-#'\itemize{
-#'\item intercept (M+Nx1);
-#'\item phi_1   (M+N x M+N);
-#'\item phi_1^star (M+N x M+N);
-#'\item phi_g (M+N x M+N);
-#'\item Sigma (M+N x G)
-#'}
-#'\item parameters of the GVAR.
-#'\itemize{
+#' @return A list containing
+#' \enumerate{
+#' \item parameters of the country-specific VARX(1,1,1)
+#' \itemize{
+#' \item intercept (M+Nx1);
+#' \item phi_1   (M+N x M+N);
+#' \item phi_1^star (M+N x M+N);
+#' \item phi_g (M+N x M+N);
+#' \item Sigma (M+N x G)
+#' }
[…]

### 2025-03-24 R/InputsForOpt.R

@@ -1,36 +1,35 @@
-#' Generates several  inputs that are necessary to build the likelihood function
+#' Generates inputs necessary to build the likelihood function for the ATSM model
 #'
-#'@param InitialSampleDate Start date of the sample period in the format "dd-mm-yyyy"
-#'@param FinalSampleDate End date of the sample period in the format "dd-mm-yyyy"
-#'@param ModelType A character vector indicating the model type to be estimated.
-#'@param Yields A numerical matrix with time series of yields (JxT or CJ x T)
-#'@param GlobalMacro A numerical matrix with time series of the global risk factors (G x T)
-#'@param DomMacro A numerical matrix with time series of the country-specific risk factors for all C countries (CM x T)
-#'@param FactorLabels A list of character vectors with labels for all variables in the model.
-#'@param Economies  A character vector containing the names of the economies included in the system.
-#'@param DataFrequency  A character vector specifying the frequency of the data. Available options are:
-#'        "Daily All Days", "Daily Business Days", "Weekly", "Monthly", "Quarterly", or "Annually".
-#'@param GVARlist A list containing the necessary inputs for the estimation of GVAR-based models
-#'@param JLLlist  A list of necessary inputs for the estimation of JLL-based models. If the chosen model is "JLL original" or "JLL joint Sigma", then a dominant unit economy must be chosen. Otherwise, this list must be set as 'None'.
-#'@param WishBRW Whether to estimate the physical parameter model with bias correction, based on the method by Bauer, Rudebusch and Wu (2012) (see \code{\link{Bias_Correc_VAR}} function). Default is set to 0.
-#'@param BRWlist List of necessary inputs for performing the bias-corrected estimation (see \code{\link{Bias_Correc_VAR}} function).
-#'@param UnitYields A character string indicating the maturity unit of yields. Options are: (i) "Month" for yields expressed in months, or (ii) "Year" for yields expressed in years. Default is "Month".
[…]

### 2025-03-24 R/NumOutputsBootstrap.R

@@ -1,19 +1,17 @@
 #' Numerical outputs (IRFs, GIRFs, FEVD, and GFEVD) for bootstrap
 #'
-#'@param ModelType A character vector indicating the model type to be estimated.
-#'@param ModelParaBoot A list of model parameter estimates (see the "Optimization" function) after a bootstrap draw
-#'@param InputsForOutputs A list containing the necessary inputs for generating IRFs, GIRFs, FEVDs, GFEVDs and Term Premia.
-#'@param FactorLabels  A list of character vectors with labels for all variables in the model.
-#'@param Economies  A character vector containing the names of the economies included in the system.
+#' @param ModelType A character vector indicating the model type to be estimated.
+#' @param ModelParaBoot A list of model parameter estimates (see the "Optimization" function) after a bootstrap draw
+#' @param InputsForOutputs A list containing the necessary inputs for generating IRFs, GIRFs, FEVDs, GFEVDs and Term Premia.
+#' @param FactorLabels A list of character vectors with labels for all variables in the model.
+#' @param Economies A character vector containing the names of the economies included in the system.
 #'
-#'@keywords internal
+#' @keywords internal
 
-
-NumOutputs_Bootstrap <- function(ModelType, ModelParaBoot, InputsForOutputs, FactorLabels, Economies){
+NumOutputs_Bootstrap <- function(ModelType, ModelParaBoot, InputsForOutputs, FactorLabels, Economies) {
 
   # IRF and GIRF
   IRFout <- IRFandGIRF_BS(ModelType, ModelParaBoot, InputsForOutputs[[ModelType]]$IRF$horiz, FactorLabels, Economies)
-
   # FEVD and GFEVD
   FEVDout <- FEVDandGFEVD_BS(ModelType, ModelParaBoot, InputsForOutputs[[ModelType]]$FEVD$horiz, FactorLabels, Economies)
 
@@ -23,121 +21,122 @@ NumOutputs_Bootstrap <- function(ModelType, ModelParaBoot, InputsForOutputs, Fac
   return(NumericalOutputs)
 }
 
-
[…]

### 2025-03-24 R/Reg_K1Q.R

@@ -22,6 +22,10 @@
 
 Reg_K1Q <- function(Y, mat, Z, dt,type){
 
+  # Validate inputs
+  if (!is.numeric(dt) || dt <= 0) { stop("dt must be a positive numeric value.")  }
+  if (!is.character(type) || !(type %in% c("Jordan", "NULL"))) { stop("type must be either 'Jordan' or 'NULL'.") }
+
   # Step 1: Define the boundaries of the maturities that will be interpolated
   M <- round(max(mat)/dt) # longest maturity adjusted for the time units of the model
   m <- round(min(mat)/dt) # shortest maturity adjusted for the time units of the model
@@ -29,7 +33,7 @@ Reg_K1Q <- function(Y, mat, Z, dt,type){
   # Step 2: Interpolate yields by spline. Interpolation is made for each time unit of the model.
   # Define target maturities
   target_maturities <- seq(dt, M * dt, by = dt)
-  b <- matrix(data=NA, ncol= ncol(Y) , nrow = M )
+  b <- matrix(data = NA, ncol = ncol(Y) , nrow = M )
 
   # Interpolate using splinefun for each column
   for (i in seq_len(ncol(Y))) {
@@ -51,7 +55,9 @@ Reg_K1Q <- function(Y, mat, Z, dt,type){
   R <- b
 
   # Step 3: regress by OLS interpolated yields onto pricing factors to obtain B: Y_t = A +B*Z_t
-  Bn <- stats::lm( Jmisc::demean(t(R))~ Jmisc::demean(t(Z))-1)$coefficients
+  lhs <-  scale(t(R), center = TRUE, scale = FALSE)
+  rhs <- scale(t(Z), center = TRUE, scale = FALSE)
+  Bn <-stats::lm( lhs ~ rhs - 1)$coefficients
   Bn <- t(Bn)
 
   # Step 4: set h such that b_h will be obtained without the need of extrapolation beyond the current maturity range
@@ -104,7 +110,6 @@ Reg_K1Q <- function(Y, mat, Z, dt,type){
 #
 #'@param K1XQ  squared matrix in non-Jordan form
 #'
-#'@importFrom pracma numel
 #'
 #'@return squared matrix in Jordam form
[…]

### 2025-03-24 R/Transition_Matrix.R

@@ -1,27 +1,25 @@
 #' Computes the transition matrix required in the estimation of the GVAR model
 #'
-#'@param t_First Sample starting date (in the format: yyyy).
-#'@param t_Last  Sample ending date (in the format: yyyy).
-#'@param Economies    A character vector containing the names of the economies included in the system.
-#'@param type   A character string indicating the method for computing interdependence. Possible options include:
-#'\itemize{
+#' @param t_First Sample starting date (in the format: yyyy).
+#' @param t_Last  Sample ending date (in the format: yyyy).
+#' @param Economies A character vector containing the names of the economies included in the system.
+#' @param type A character string indicating the method for computing interdependence. Possible options include:
+#' \itemize{
 #'      \item \code{Time-varying}: Computes time-varying interdependence and returns the weight matrices for each year based on available data (may extrapolate the sample period).
 #'      \item \code{Sample Mean}: Returns a single weight matrix containing the average weights over the entire sample period, suitable for time-invariant interdependence.
 #'      \item A specific year (e.g., "1998", "2005"): Used to compute time-invariant interdependence for the specified year.
-#'}
+#' }
+#' @param DataConnectedness Data used to compute the transition matrix. Default is set to NULL.
+#' @param DataPath Path to the Excel file containing the data (if applicable). The default is linked to the Excel file available in the package.
 #'
-#'@param DataConnectedness    Data used to compute the transition matrix. Default is set to NULL.
-#'@param DataPath Path to the Excel file containing the data (if applicable). The default is linked to the Excel file available in the package.
+#' @return matrix or list of matrices
 #'
-#
-#'@return matrix or list of matrices
-#'
-#'@details
+#' @details
 #' If there is missing data for any country of the system for that particularly year,
 #' then the transition matrix will include only NAs.
 #'
-#'@examples
[…]

### 2025-05-11 inst/doc/Guidelines.Rmd
version 1.3.1

@@
 
-The previous function loads four types of datasets. The first dataset
+The previous function loads three types of datasets. The first dataset
 consists of time series for zero-coupon bond yields from four emerging
@@
 yield maturities must be consistent across all countries (although the
-`DataForEstimation()` function can take different maturities as inputs,
-it outputs bond yields with uniform maturities for all economies); and
-*(ii)* the bond yields should be expressed in percentage terms per
-annum, not in basis points. The *MultiATSM* package does not include
-routines for bootstrapping zero-coupon yields from coupon bonds, so
-users must handle this data manipulation themselves if needed.
+`DataForEstimation()` and `InputsForOpt()` functions can take different 
+maturities as inputs, it outputs bond yields with uniform maturities for 
+all economies); and *(ii)* the bond yields should be expressed in percentage 
+terms per annum, not in basis points. The *MultiATSM* package does not include 
+routines for bootstrapping zero-coupon yields from coupon bonds, so users must
+handle this data manipulation themselves if needed. The data can also be retrieved 
+using:
 
@@
 
-The second dataset focuses on the time series of risk factors. Using the
-terminology of JPS (2014), this dataset includes *(i)* country-specific
-**spanned factors** (commonly represented by level, slope, and curvature
--- see Section \@ref(S:SpaFac) for an intuitive explanation of these
-factors) and *(ii)* a set of country-specific and global **unspanned
-factors**, namely macroeconomic variables such as economic growth and 
-inflation. Similar to the bond yield data, these economic indicators 
-must be constructed by the user.
+Bond yield data are also used to construct the country-specific **spanned** factors,
+according to the terminology of JPS (2014). These factors typically correspond to 
+the standard yield curve components, namely level, slope, and curvature. An intuitive 
+explanation of their statistical properties is provided in Section \@ref(S:SpaFac).
+
+The second dataset contains time series for macroeconomic risk factors — also 
+referred to as **unspanned** factors, following the classification of JPS (2014). 
+These include measures of economic growth and inflation for *(i)* each of the 
+four countries in the sample and *(ii)* the global economy. Similar to the bond 
+yield data, these economic indicators must be constructed by the user.
 
 ```{r}
-data('CM_Factors')
+data("CM_GlobalMacroFactors")
+data('CM_DomMacroFactors')
 ```
 
-The final two datasets are essential for estimating GVAR-based models.
-The trade flows data, sourced from the International Monetary Fund
-(IMF), presents the sum of the value of all goods imports and exports
-between any two countries of the sample on a yearly basis since 1948.
-All the values are free on board and are expressed in U.S. dollars.
-These data are used to construct the transition matrix of the GVARs
-models.
+The final dataset captures cross-country interconnectedness measures, proxied by 
+bilateral trade flows, and is essential for estimating GVAR-based models. 
+Specifically, the trade data—sourced from the International Monetary Fund (IMF) —
+reports the annual value of total goods traded (imports plus exports)
+between each pair of countries in the sample, available from 1948 onward. All the values are free on board 
+and are expressed in U.S. dollars. These data are specifically used to 
+construct the transition matrix of the GVAR models.
 
@@
 data('CM_Trade')
-```
-
-The GVAR factors database casts country-specific lists of all the
-factors that are used in the estimation of each country's VARX(1,1,1).
-The function titled `DatabasePrep()` provides assistance to structure
-the data in a similar fashion to this form. Moreover, a specific
-function for computing the star risk factors is detailed in the Section
-\@ref(S:SectionGVAR).
-
-```{r}
-data('CM_Factors_GVAR')
 ```
@@
 ```{r}
+data('CM_Factors_GVAR')
 GVARinputs <- list(Economies = Economies, GVARFactors = FactorsGVAR, VARXtype ="constrained: Inflation")
@@
 data("CM_Factors_GVAR")
-
 GVARinputs <- list(Economies = Economies, GVARFactors = FactorsGVAR, VARXtype = "unconstrained", 

### 2025-10-23 inst/doc/MultiATSM.Rmd
version 1.5.0

--- 2025-05-11 inst/doc/Guidelines.Rmd
@@
 
-Due to peculiar features of JPS-based specifications, an efficient
+Due to peculiar features of JPS-based specifications (see @JoslinSingletonZhu2011), an efficient
 estimation of the parameters governing the risk-neutral ($Q$-measure)
@@
 
+<a id="tab-ModFea"></a>
+
 ```{r ModFea, message=FALSE, echo=FALSE}
-ModelLabels <- c("JPS original", "JPS global", "JPS multi", "GVAR single", "GVAR multi", 
-                 "JLL original", "JLL No DomUnit", "JLL joint Sigma")
+ModelLabels <- c(
+  "JPS original", "JPS global", "JPS multi", "GVAR single", "GVAR multi",
+  "JLL original", "JLL No DomUnit", "JLL joint Sigma"
+)
 
@@
 # Adjust column names
-ColNames <- c("","","","","JLL", "GVAR", "", "", "", "", "", "", "","", "", "","")
+ColNames <- c("", "", "", "", "JLL", "GVAR", "", "", "", "", "", "", "", "", "", "", "")
 colnames(Tab) <- ColNames
@@
 
-kableExtra::kbl(Tab, align = "c", caption = "Model Features") %>%
+kableExtra::kbl(
+  Tab,
+  align = "c",
+  caption = "Model Features"
+) %>%
   kableExtra::kable_classic("striped", full_width = F) %>%
   kableExtra::row_spec(0, font_size = 10) %>%
-  kableExtra::add_header_above(c(" "=2, "Unrestricted" = 1, "Restricted" = 1, "Unrestricted" = 1, "Restricted" = 2, " " = 11)) %>%
- kableExtra::add_header_above(c(" "=2, "Individual" = 2, "Joint" = 3, " "=2, "Individual" = 1, "Joint" = 1, " "=2, "P only" = 1, "P and Q" = 1, " " = 3))  %>%
-  kableExtra::add_header_above(c( " "=2, "P-dynamics"= 5, " "=2, "Q-dynamics"= 2, " "=2, "Sigma matrix estimation" = 2, " "=2, "Dominant Country"=1), bold = T) %>%
-kableExtra::pack_rows("Unrestricted VAR", 1, 3 , label_row_css = "background-color: #666; color: #fff;")  %>%
+  kableExtra::add_header_above(c(" " = 2, "Unrestricted" = 1, "Restricted" = 1, "Unrestricted" = 1, "Restricted" = 2, " " = 11)) %>%
+  kableExtra::add_header_above(c(" " = 2, "Individual" = 2, "Joint" = 3, " " = 2, "Individual" = 1, "Joint" = 1, " " = 2, "P only" = 1, "P and Q" = 1, " " = 3)) %>%
+  kableExtra::add_header_above(c(" " = 2, "P-dynamics" = 5, " " = 2, "Q-dynamics" = 2, " " = 2, "Sigma matrix estimation" = 2, " " = 2, "Dominant Country" = 1), bold = T) %>%
+  kableExtra::pack_rows("Unrestricted VAR", 1, 3, label_row_css = "background-color: #666; color: #fff;") %>%
   kableExtra::pack_rows("Restricted VAR (GVAR)", 4, 5, label_row_css = "background-color: #666; color: #fff;") %>%
@@
 In what follows, the focus will be devoted to describe the various
-pieces that are necessary to implement ATSMs. Specifically, Section
-\@ref(S:SectionData) describes the data-set available at this package
+pieces that are necessary to implement ATSMs. Specifically, [Section
+Data](#SectionData) describes the data-set available at this package
 and the set of functions that are useful to retrieve data from Excel
-files. Section \@ref(S:SectionInputs) exposes the necessary inputs that
-have to be specified by the user. In Section \@ref(S:SectionEstimation)
+files. [Section Required user inputs](#SectionInputs) exposes the necessary inputs that
+have to be specified by the user. In Section [Model Estimation](#SectionEstimation)
 the estimation procedure is detailed. Section
-\@ref(S:SectionReplication) shows how to use the *MultiATSM* package to
+[Model Estimation](#SectionEstimation) shows how to use the *MultiATSM* package to
 estimate ATSMs from scratch.
 
-# Data {#S:SectionData}
+# Data {#SectionData}
 
@@
 ```{r}
-data('CM_Yields')
+data("Yields")
 ```
@@
 the standard yield curve components, namely level, slope, and curvature. An intuitive 
-explanation of their statistical properties is provided in Section \@ref(S:SpaFac).
+explanation of their statistical properties is provided in Section [Spanned Factors](#SpaFac).
 
@@
 ```{r}
-data("CM_GlobalMacroFactors")
-data('CM_DomMacroFactors')
+data("GlobalMacro")
+data("DomMacro")
 ```
@@
 ```{r}
-data('CM_Trade')
+data("TradeFlows")
 ```
@@
 *(iii)* have identical variable labels across all the tabs within each
-file. For illustration, see the Excel file available at the package. One
-example of list of inputs to be provided is, for instance,
-
+file. For illustration, see the Excel file available at the package. 
+The workflow is as follows:
+
+  1. Retrieve the full dataset from excel files using `Load_Excel_Data()`:
+```{r}
+MacroData <- Load_Excel_Data(system.file("extdata", "MacroData.xlsx", package = "MultiATSM"))
+YieldData <- Load_Excel_Data(system.file("extdata", "YieldsData.xlsx", package = "MultiATSM"))
+```
+
+  2. Specify a subset of the selected inputs. For instance,
 ```{r}
@@
 
-Using these inputs, the variable *RiskFac_TS* can be constructed,
-containing the full time series of risk factors required for the model.
+  3. Conditional on the previous step, the variable *RiskFac_TS* can be constructed. This variable
+contains the full time series of risk factors required for the model.
 
@@
 FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
-RiskFac_TS <- DataForEstimation(Initial_Date, Final_Date, Economies, N, FactorLabels, ModelType,
-                                DataFrequency)
-```
-
-# Required user inputs {#S:SectionInputs}
+RiskFac_TS <- DataForEstimation(
+  Initial_Date, Final_Date, Economies, N, FactorLabels, ModelType,
+  DataFrequency, MacroData, YieldData
+)
+```
+
+# Required user inputs {#SectionInputs}
 
@@
 -   *Model type*: a character vector indicating the model type to be
-    estimated, as presented in Table \@ref(tab:ModFea).
+    estimated, as presented in [Table&nbsp;0.1](#tab-ModFea).
 
@@
 
--   A binary variable that equals $1$ if the user wishes to impose the
+-   A logical variable. Choose *TRUE* to impose the
     constraint that the largest eigenvalue under the $Q$-measure is
-    strictly less than $1$ otherwise, it should be set to $0$. While
+    strictly less than $1$ otherwise, it should be set to *FALSE*. While
     enforcing this stationary constraint may prolong model estimation,
@@
 
-**6. Output label**
+**6. Selected folder to save outputs**
+
+-  Path where the selected graphical outputs will be saved. If set to `NULL`, the outputs are stored in the user's *temporary directory* (accessible via `tempdir()`).
+
+**7. Output label**
 
@@
 # 5) Risk-neutral stationary constraint
-StationarityUnderQ <- 0 # 1 = set stationary condition under the Q; 0 = no stationary condition
-
-# 6) Output label
+StationarityUnderQ <- FALSE
+
+# 6) Path to save outputs
+Folder2Save <- NULL
+
+# 7) Output label
 OutputLabel <- "Model_demo"
@@
 *GVARlist* and *JLLlist* for clarity. This section outlines the general
-structure of both lists, while Sections \@ref(S:SectionGVAR) and
-\@ref(S:SectionJLL) provide detailed explanations and available options.
+structure of both lists, while Sections [GVAR](#SectionGVAR) and
+[JLL](#SectionJLL) provide detailed explanations and available options.
 
@@
 ```{r}
-W_type <- 'Sample Mean' # Method to compute the transition matrix
+W_type <- "Sample Mean" # Method to compute the transition matrix
 t_First_Wgvar <- "2000" # First year of the sample
@@
 ```{r}
-GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2000",
-                   t_Last_Wgvar = "2015", DataConnectedness = TradeFlows) 
+GVARlist <- list(
+  VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2000",
+  t_Last_Wgvar = "2015", DataConnectedness = TradeFlows
+)
 ```
@@
 
--   *Mean or median of OLS estimates* *(flag_mean)*: the user must
-    decide whether to compute the mean or median of the OLS estimates
+-   *Mean or median of P-dynamics estimates* *(Cent_Measure)*: the user must
+    decide whether to compute the mean or median of the physical dynamics estimates
     after each bootstrap iteration. To compute the mean (median), the
-    user must set *TRUE* (*FALSE*);
+    user must set *Mean* (*Median*);
 
@@
 
-```{r}
-BRWlist <- within(list(flag_mean = TRUE, gamma = 0.2, N_iter = 500, B = 50, 
-                       checkBRW = TRUE, B_check = 1000), N_burn <- round(N_iter * 0.15))
+-   *Eigenvalue restriction (Eigen_rest)*: set a restriction on the largest eigenvalue 
+     under the $P$-measure. Default is $1$.
+
+```{r}
+BRWlist <- within(list(
+  Cent_Measure = "Mean", gamma = 0.2, N_iter = 500, B = 50,
+  checkBRW = TRUE, B_check = 1000, Eigen_rest = 1
+), N_burn <- round(N_iter * 0.15))
 ```
@@
 
-ATSMInputs <- InputsForOpt(t0, tF, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,  
-                           FactorLabels, Economies, DataFreq, CheckInputs = FALSE)
+ATSMInputs <- InputsForOpt(t0, tF, ModelType, Yields, GlobalMacro, DomMacro,
+  FactorLabels, Economies, DataFreq,
+  CheckInputs = FALSE
+)
 ```
@@
 ```{r}
-WishGraphRiskFac <- 0 #   YES: 1; No = 0.
-WishGraphYields <- 1 #    YES: 1; No = 0.
-WishOrthoJLLgraphs <- 0 # YES: 1; No = 0.
+WishGraphRiskFac <- FALSE
+WishGraphYields <- TRUE
+WishOrthoJLLgraphs <- FALSE
 ```
@@
 customizing the features of the wished outputs. Conditional to these
-settings, individual folders are created at the user's *temporary
-directory* (which can be accessed using `tempdir()` ) to store the
-different types of the desired graphical outputs.
+settings, individual folders are created at the selected `Folder2Save` to store the
+different types of the desired graphical outputs. The selected outputs are also printed to the user' console.
 
@@
 $FP_t^{(n)}$ corresponds to the forward premia. In this case, the user
-must specify a binary variable set to $1$ if the computation of forward
-premia is desired, or $0$ otherwise. If set to $1$, the user must also
+must specify a logical variable set to *TRUE* if the computation of forward
+premia is desired, or *FALSE* otherwise. If set to *TRUE*, the user must also
 provide a two-element numerical vector containing the maturities
@@
 ```{r}
-    WishFPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0 
+WishFPremia <- TRUE
 FPmatLim <- c(60, 120)
@@
 ```{r}
-Bootlist <- list(methodBS = 'block', BlockLength = 4, ndraws =  50, pctg   =  95)
+Bootlist <- list(methodBS = "block", BlockLength = 4, ndraws = 50, pctg = 95)
 ```
@@
 
-# Model Estimation {#S:SectionEstimation}
+# Model Estimation {#SectionEstimation}
 
@@
 
-## Spanned Factors {#S:SpaFac}
+## Spanned Factors {#SpaFac}
 
@@
 
-w_pca <- data.frame(t(w[1:N,]))
+w_pca <- data.frame(t(w[1:N, ]))
 colnames(w_pca) <- LabSpaFac
@@
 
-g <-  ggplot2::ggplot(data = w_pca, ggplot2::aes(x=  mat)) +  
+g <- ggplot2::ggplot(data = w_pca, ggplot2::aes(x = mat)) +
   ggplot2::geom_line(ggplot2::aes(y = Level, color = "Level"), linewidth = 0.7) +
@@
   ggplot2::geom_line(ggplot2::aes(y = Curvature, color = "Curvature"), linewidth = 0.7) +
-      ggplot2::labs(color = "Legend") + ggplot2::scale_color_manual(values = colors) + ggplot2::theme_classic() +
-    ggplot2::theme(axis.title.y= ggplot2::element_blank(), legend.position="top", legend.title=ggplot2::element_blank(), legend.text= ggplot2::element_text(size=8) ) + 
-   ggplot2::xlab("Maturity (Years)") + ggplot2::geom_hline(yintercept=0)
+  ggplot2::labs(color = "Legend") +
+  ggplot2::scale_color_manual(values = colors) +
+  ggplot2::theme_classic() +
+  ggplot2::theme(axis.title.y = ggplot2::element_blank(), legend.position = "top", legend.title = ggplot2::element_blank(), legend.text = ggplot2::element_text(size = 8)) +
+  ggplot2::xlab("Maturity (Years)") +
+  ggplot2::geom_hline(yintercept = 0)
 
@@
 ```{r}
-data('CM_Yields')
+data("Yields")
 Economies <- c("China", "Brazil", "Mexico", "Uruguay")
@@
 
-As presented in Table \@ref(tab:ModFea), the dynamics of the risk
+As presented in [Table&nbsp;0.1](#tab-ModFea), the dynamics of the risk
 factors under the $P$-measure evolve according to a VAR(1) model, which
@@
 ```{r}
-data("CM_Factors")
-PdynPara <- VAR(RiskFactors, VARtype= "unconstrained")
+data(RiskFacFull)
+PdynPara <- VAR(RiskFacFull, VARtype = "unconstrained")
 ```
@@
 ```{r}
-FactorsChina <- RiskFactors[1:7, ]
-PdynPara <- VAR(FactorsChina, VARtype= "unconstrained")
+FactorsChina <- RiskFacFull[1:7, ]
+PdynPara <- VAR(FactorsChina, VARtype = "unconstrained")
 ```
@@
 
-### GVAR {#S:SectionGVAR}
+### GVAR {#SectionGVAR}
 
@@
 ```{r}
-data('CM_Factors_GVAR')
-GVARinputs <- list(Economies = Economies, GVARFactors = FactorsGVAR, VARXtype ="constrained: Inflation")
+data(GVARFactors)
+GVARinputs <- list(Economies = Economies, GVARFactors = GVARFactors, VARXtype = "constrained: Inflation")
 ```
@@
 ```{r}
-data("CM_Trade")
+data("TradeFlows")
 t_First <- "2006"
@@
 type <- "Sample Mean"
-W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, DataPath = NULL, TradeFlows)
+W_gvar <- Transition_Matrix(t_First, t_Last, Economies, type, TradeFlows)
 print(W_gvar)
@@
 ```{r}
-data("CM_Factors_GVAR")
-GVARinputs <- list(Economies = Economies, GVARFactors = FactorsGVAR, VARXtype = "unconstrained", 
-                   Wgvar = W_gvar)
+data("GVARFactors")
+GVARinputs <- list(
+  Economies = Economies, GVARFactors = GVARFactors, VARXtype = "unconstrained",
+  Wgvar = W_gvar
+)
 N <- 3
@@
 
-A separate routine is provided for computing the foreign-specific
-factors (also commonly referred to as the star variables) used in the
-estimation of the VARX models.
-
-```{r}
-data('CM_Factors')
-StaFac <- StarFactors(RiskFactors, Economies, W_gvar)
-```
-
-### JLL {#S:SectionJLL}
+### JLL {#SectionJLL}
 
@@
 
-3.  *Wish the estimation of the sigmas matrices*: this is a binary
-    variable which assumes value equal to $1$ if the user wishes the
+3.  *Wish the estimation of the sigmas matrices*: this is a logical
+    variable which assumes value equal to *TRUE* if the user wishes the
     estimation of all JLL sigma matrices (i.e. variance-covariance and
-    the Cholesky factorization matrices) and, $0$ otherwise. Since this
+    the Cholesky factorization matrices) and, *FALSE* otherwise. Since this
     estimation is done numerically, its estimation can take several
@@
 5.  *JLL type*: a string-vector containing the label of the JLL model to
-    be estimated according to the features described in Table
-    \@ref(tab:ModFea).
+    be estimated according to the features described in [Table&nbsp;0.1](#tab-ModFea).
 
@@
 ModelType <- "JLL original"
-JLLinputs <- list(Economies = Economies, DomUnit = "China", WishSigmas = 1,  SigmaNonOrtho = NULL,
-                  JLLModelType = ModelType)
+JLLinputs <- list(
+  Economies = Economies, DomUnit = "China", WishSigmas = TRUE, SigmaNonOrtho = NULL,
+  JLLModelType = ModelType
+)
 ```
@@
 ```{r, eval=FALSE}
-data("CM_Factors")
+data("RiskFacFull")
 N <- 3
-JLLpara <- JLL(RiskFactors, N, JLLinputs, CheckInputs = TRUE)
+JLLpara <- JLL(RiskFacFull, N, JLLinputs, CheckInputs = TRUE)
 ```
@@
 ATSM estimation involves optimizing the log-likelihood function by
-optimally selecting the model parameters. The optimization structure in
-this package is based on routines from the term structure package by
-@LeSingleton2018. The ATSM otimization is carried out using the
-`Optimization()` function.
-
+optimally selecting the model parameters via the `Optimization()` function.
 Broadly, the unspanned risk framework from JPS (2014) (and, therefore,
-all its multi-country extensions) requires the estimation parameter set
-consisting of 6 elements, namely: the risk-neutral long-run mean of the
-short rate (*r0*), the risk-neutral feedback matrix (*K1XQ*), the
-variance-covariance matrix (*SSZ*) from the VAR processes, the standard
-deviation of the errors from the portfolios of yields observed with
-error (*se*), in addition to the intercept (*K0Z*) and the feedback
-(*K1Z*) matrices of the physical dynamics.
+all its multi-country extensions) follows a model parameterization similar to that proposed by @JoslinSingletonZhu2011. Specifically, it requires the estimation parameter set
+consisting of 6 elements, namely the:
+
+1.    Risk-neutral long-run mean of the
+short rate (*r0*);
+
+2.    Risk-neutral feedback matrix (*K1XQ*);
+
+3.    Variance-covariance matrix (*SSZ*) from the VAR processes;
+
+4.    The standard deviation of the errors from the portfolios of yields observed with
+error (*se*);
+
+5.    Intercept of the physical dynamics (*K0Z*);
+
+6.    Feedback matrices of the physical dynamics (*K1Z*).
 
@@
 function. However, the remaining parameters (*K1XQ* and *SSZ*) require
-numerical estimation. @LeSingleton2018 provide standardized routines
-(integrated into this package) to set effective initial values,
-facilitating the optimization process.
-
-The *MultiATSM* package offers two optimization algorithms: *fminunc*
-and *fminsearch*. By default, both algorithms are run sequentially twice
+numerical estimation. This package provides standardized routines
+to set effective initial values, facilitating the optimization process.
+
+The *MultiATSM* package offers two optimization algorithms: *L-BFGS-B* and *Nelder-Mead*. By default, both algorithms are run sequentially twice
 to ensure the model reaches the global optimum. Specifically in the
-default bootstrap setup, only *fminunc* is used to save time.
+default bootstrap setup, only *Nelder-Mead* is used to save time.
 
@@
 results presented in the original papers of JPS (2014), CM (2023) and CM
-(2024). See the [Paper Replication Vignette](Paper-Replications.html) of
-this package for a detailed reproduction.
+(2024). See the **paper replication vignette** (included with this package) for a detailed reproduction.
 
@@
 ########################################################################################################
+library(MultiATSM)
 # A) Load database data
@@
 # B) GENERAL model inputs
-ModelType <- "JPS multi" # available options: "JPS original", "JPS global", "GVAR single", "JPS multi",
-                          #"GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".
-
-Economies <- c("China", "Brazil", "Mexico") # Names of the economies from the economic system
-GlobalVar <- c("Gl_Eco_Act")# Global Variables
-DomVar <- c("Eco_Act", "Inflation") # Country-specific variables
+ModelType <- "JPS original" # available options: "JPS original", "JPS global", "GVAR single", "JPS multi",
+# "GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".
+
+Economies <- c("China", "Brazil") # Names of the economies from the economic system
+GlobalVar <- c("Gl_Eco_Act") # Global Variables
+DomVar <- c("Eco_Act") # Country-specific variables
 N <- 2 # Number of spanned factors per country
 
-t0_sample <- "01-05-2005" # Format: "dd-mm-yyyy"
+t0_sample <- "01-12-2007" # Format: "dd-mm-yyyy"
 tF_sample <- "01-12-2019" # Format: "dd-mm-yyyy"
@@
 OutputLabel <- "Test" # label of the model for saving the file
-DataFreq <-"Monthly" # Frequency of the data
-
-StatQ <- 0 # Wish to impose stationary condition for the eigenvalues of each country: YES: 1,NO:0
+DataFreq <- "Monthly" # Frequency of the data
+
+Folder2Save <- NULL # Folder to save the required graphical outputs. NULL -> temporary directory
+StatQ <- FALSE # Wish to impose stationary condition for the eigenvalues of each country
 
@@
 #################################### GVAR-based models ##################################################
-GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2005",
-                  t_Last_Wgvar = "2019", DataConnectedness <- TradeFlows ) 
+GVARlist <- list(
+  VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2005",
+  t_Last_Wgvar = "2019", DataConnectedness = TradeFlows
+)
 # VARXtype: Available options "unconstrained" or "constrained" (VARX)
@@
 ###################################### BRW inputs  ######################################################
-WishBC <- 0 # Wish to estimate the model with the bias correction method of BRW (2012): #YES: 1, NO:0
-BRWlist <- within(list(flag_mean = TRUE, gamma = 0.05, N_iter = 250, B = 50, checkBRW = TRUE,
-                       B_check = 1000),  N_burn <- round(N_iter * 0.15))
-# flag_mean: TRUE = compute the mean; FALSE = compute the median
+WishBC <- FALSE # Wish to estimate the model with the bias correction method of BRW (2012)
+BRWlist <- within(list(
+  Cent_Measure = "Mean", gamma = 0.05, N_iter = 250, B = 50, checkBRW = TRUE,
+  B_check = 1000, Eigen_rest = 1
+), N_burn <- round(N_iter * 0.15))
+# Cent_Measure: "Mean" = compute the mean; "Median" = compute the median
 # gamma: Adjustment parameter
@@
 # C) Decide on Settings for numerical outputs
-WishFPremia <- 1 # Wish to estimate the forward premia: YES: 1, NO:0
-FPmatLim <- c(60,120) #  If the forward premia is desired, then choose the Min and max maturities of the
+WishFPremia <- TRUE # Wish to estimate the forward premia
+FPmatLim <- c(60, 120) #  If the forward premia is desired, then choose the Min and max maturities of the
 # forward premia. Otherwise set NA
@@
 DesiredGraphs <- c("Fit", "IRF", "TermPremia") # "Fit", "IRF", "FEVD", "GIRF", "GFEVD", "TermPremia"
-WishGraphRiskFac <- 0
-WishGraphYields <- 1
-WishOrthoJLLgraphs <- 0
+WishGraphRiskFac <- FALSE
+WishGraphYields <- TRUE
+WishOrthoJLLgraphs <- FALSE
 
 # D) Bootstrap settings
-WishBootstrap <- 1 #  YES: 1; No = 0.
-BootList <- list(methodBS = 'bs', BlockLength = 4, ndraws = 5, pctg =  95)
+WishBootstrap <- TRUE
+BootList <- list(methodBS = "bs", BlockLength = 4, ndraws = 4, pctg = 95)
 # methodBS: bootstrap method. Available options: (i) 'bs' ; (ii) 'wild'; (iii) 'block'
@@
 # E) Out-of-sample forecast
-WishForecast <- 1 #  YES: 1; No = 0.
-ForecastList <- list(ForHoriz = 12,  t0Sample = 1, t0Forecast = 162, ForType = "Rolling")
+WishForecast <- TRUE
+ForecastList <- list(ForHoriz = 12, t0Sample = 1, t0Forecast = 131, ForType = "Rolling")
 # ForHoriz: forecast horizon
@@
 #########################################################################################################
-############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
+## NO NEED TO MAKE CHANGES FROM HERE:
+## The sections below automatically process the inputs provided above, run the model
+## estimation, generate the numerical and graphical outputs, and save results.
 #########################################################################################################
@@
 # 3) Prepare the inputs of the likelihood function
-ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,
-                           FactorLabels, Economies, DataFreq, GVARlist, JLLlist, WishBC, BRWlist)
+ATSMInputs <- InputsForOpt(
+  t0_sample, tF_sample, ModelType, Yields, GlobalMacro, DomMacro,
+  FactorLabels, Economies, DataFreq, GVARlist, JLLlist, WishBC, BRWlist
+)
 
@@
 # a) Prepare list of inputs for graphs and numerical outputs
-InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
+InputsForOutputs <- InputsForOutputs(
+  ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
   WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishFPremia,
-                                    FPmatLim, WishBootstrap, BootList, WishForecast, ForecastList)
+  FPmatLim, WishBootstrap, BootList, WishForecast, ForecastList
+)
 
 # b) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
-NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)
+NumericalOutputs <- NumOutputs(
+  ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
+  Folder2Save
+)
 
 # c) Confidence intervals (bootstrap analysis)
-BootstrapAnalysis <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, Economies, InputsForOutputs,
-                               FactorLabels, JLLlist, GVARlist, WishBC, BRWlist)
+BootstrapAnalysis <- Bootstrap(
+  ModelType, ModelParaList, NumericalOutputs, Economies, InputsForOutputs,
+  FactorLabels, JLLlist, GVARlist, WishBC, BRWlist, Folder2Save
+)
 
 # 6) Out-of-sample forecasting
-Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
-                            JLLlist, GVARlist, WishBC, BRWlist)
+Forecasts <- ForecastYields(
+  ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
+  JLLlist, GVARlist, WishBC, BRWlist, Folder2Save
+)
 ```
@@
 Starting from version $1.3.0$, the *MultiATSM* package provides additional ways to concisely inspect key elements of the selected ATSM.
-In particular, users can now gain an alternative perspective on the general model inputs, optimized model parameters, and root mean square error (RMSE) from the forecast analysis.
-For clarity, the variable names used hereafter align with those presented in Section \@ref(S:SectionTemplate).
+In particular, users can now gain an alternative perspective on the general model inputs, optimized model parameters, graphocal outputs and root mean square error (RMSE) from the forecast analysis.
+For clarity, the variable names used hereafter align with those presented in Section [General template](#SectionTemplate).
 
@@
 data("Out_Example")
-print(Out$ATSMInputs)
-```
-
-```{r}
-summary(Out$ATSMInputs) 
+print(Out_Example$ATSMInputs)
+```
+
+```{r}
+summary(Out_Example$ATSMInputs)
 ```
@@
 ```{r}
-summary(Out$ModelParaList) 
-```
+summary(Out_Example$ModelPara)
+```
+
+### Graphical outputs
+The `NumOutputs()` function returns an object of class *ATSMNumOutputs*, which includes an `autoplot()` $S3$ method.
+This method allows users to inspect the empirical properties of the chosen specification by displaying graphical outputs in the console, regardless of whether they were saved to *Folder2Save*.
+The user must specify a plot type: `RiskFactors`, `Fit`, `TermPremia`, or one of `IRF`, `FEVD`, `GIRF`, `GFEVD` (each suffixed with `_Factors` or `_Yields`).
+For JLL-based models, an additional `_Ortho` suffix generates orthogonalized outputs.
+The examples below illustrate the responses of the Brazilian and Chinese yield curves to a global economic activity shock:
+
+```{r, fig.alt="GIRFs - Responses of yields curves in China and Brazil to a global economic shocks"}
+GIRF_all <- autoplot(Out_Example$NumOut, type = "GIRF_Yields")
+GIRF_all$Brazil$Gl_Eco_Act
+GIRF_all$China$Gl_Eco_Act
+```
+
+Additionally, the `autoplot()` $S3$ method for *ATSMModelBoot* objects—generated by the `Bootstrap()` function—allows users to assess the statistical significance of the results.
+The example below shows the responses of 12-month bond yields to several shocks.
+
+```{r, fig.alt="GIRFs - Responses of 12-month maturity bonds in China and Brazil to several shocks"}
+GIRF_all_Boot <- autoplot(Out_Example$Bootstrap, Out_Example$NumOut, type = "GIRF_Yields_Boot")
+GIRF_all_Boot$China$Y12M_China
+GIRF_all_Boot$Brazil$Y12M_Brazil
+```
+
+In this context, the `autoplot()` method requires two additional inputs beyond the *ATSMModelBoot* object.
+First, it takes an *ATSMNumOutputs* object to provide the point estimates of the selected output.
+Second, the user must specify the plot type, which follows the same structure as for *ATSMNumOutputs* objects, but with the added `_Boot` suffix.
+It is worth noting that, in this example, the point estimates (black lines) occasionally lie outside the confidence bands.
+This is not surprising, as the bootstrap was carried out with only a few draws (4).
 
@@
 
-```{r}
-plot(Out$Forecasts) 
+```{r  fig.alt="Root mean square error (RMSE)"}
+plot(Out_Example$Forecasts)
 ```

### 2025-10-23 inst/doc/Paper-Replications.Rmd
version 1.5.0

@@
 
-A thorough explanation on the usage of the additional functionalities of the *MultiATSM* package is available at the [package vignette](Guidelines.html).
+A thorough explanation on the usage of the additional functionalities of the *MultiATSM* package are documented in the vignette **General Guidelines**.
 
@@
 
-```{r, eval=FALSE}
+```{r}
 # A) Load database data
@@
 
-Economies <- c("U.S.") # Names of the economies from the economic system
+Economies <- c("US") # Names of the economies from the economic system
 GlobalVar <- c() # Global Variables
@@
 
-StatQ <- 0 # Stationary condition
+StatQ <- FALSE # Stationary condition
 
@@
 ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,
-                           FactorLabels, Economies, DataFreq)
+  FactorLabels, Economies, DataFreq,
+  verbose = FALSE
+)
 
 # 4) Optimization of the model
-ModelPara <- Optimization(ATSMInputs, StatQ, DataFreq, FactorLabels, Economies, ModelType)
+ModelPara <- Optimization(ATSMInputs, StatQ, DataFreq, FactorLabels, Economies, ModelType,
+  verbose = FALSE
+)
 ```
@@
 options(scipen = 100) # eliminate the scientific notation
-data("BR_jps_gro_R3")
-data("JPSrep")
 
-RowsQ <- c("$r0$", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_3$" )
-TableQ <- data.frame(matrix(NA, ncol = 0, nrow =length(RowsQ)))
+RowsQ <- c("$r0$", "$\\lambda_1$", "$\\lambda_2$", "$\\lambda_3$")
+TableQ <- data.frame(matrix(NA, ncol = 0, nrow = length(RowsQ)))
 row.names(TableQ) <- RowsQ
 
-PackageQ<- c(ModelPara$ests$r0, diag(ModelPara$ests$K1XQ))
+PackageQ <- c(ModelPara$`JPS original`$US$ModEst$Q$r0, diag(ModelPara$`JPS original`$US$ModEst$Q$K1XQ))
 BRq <- c(BR_jps_out$est.llk$rho0.cP, diag(BR_jps_out$est.llk$KQ.XX))
 TableQ$MultiATSM <- PackageQ
-TableQ$'BR (2017)' <- BRq
+TableQ$"BR (2017)" <- BRq
 
@@
 
+<a id="tab-PdynTab"></a>
+
 ```{r PdynTab, echo= FALSE}
-data("BR_jps_gro_R3")
-data("JPSrep")
-
 RowsP <- c("PC1", "PC2", "PC3", "GRO", "INF")
@@
 # 1) K0Z and K1Z
-# Bauer and Rudebusch coefficients
-TablePbr <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
+# a) Bauer and Rudebusch coefficients
+TablePbr <- data.frame(matrix(NA, ncol = length(ColP), nrow = length(RowsP)))
 row.names(TablePbr) <- RowsP
@@
 TablePbr[[ColP[1]]] <- BR_jps_out$est.llk$KP.0Z
-for(j in 1:length(RowsP) ){TablePbr[[RowsP[j]]] <- BR_jps_out$est.llk$KP.ZZ[,j]}
+for (j in 1:length(RowsP)) {
+  TablePbr[[RowsP[j]]] <- BR_jps_out$est.llk$KP.ZZ[, j]
+}
 
@@
 
-# MultiATSM coefficients
-TablePMultiATSM <- data.frame(matrix(NA, ncol = length(ColP), nrow =length(RowsP)))
+# b) MultiATSM coefficients
+TablePMultiATSM <- data.frame(matrix(NA, ncol = length(ColP), nrow = length(RowsP)))
 row.names(TablePMultiATSM) <- RowsP
@@
 
-IdxVar <- c(3:5, 1:2) # indexes to flip the order of the spanned and unspanned factors
-TablePMultiATSM[[ColP[1]]] <- ModelPara$ests$K0Z[IdxVar]
-ModelPara$ests$K1Z <- ModelPara$ests$K1Z[IdxVar,IdxVar]
-for(j in 1:length(RowsP) ){ TablePMultiATSM[[RowsP[j]]] <- ModelPara$ests$K1Z[,j]}
+# Estimate the P-dynamics using the weights provided by BR necessary for replication)
+PP <- BR_jps_out$W[1:N, ] %*% Yields
+ZZ <- rbind(PP, DomesticMacroVar)
+Pdyncoef <- VAR(ZZ, "unconstrained")
+
+TablePMultiATSM[[ColP[1]]] <- Pdyncoef$K0Z
+for (j in 1:length(RowsP)) {
+  TablePMultiATSM[[RowsP[j]]] <- Pdyncoef$K1Z[, j]
+}
 
 TablePMultiATSM <- round(TablePMultiATSM, digits = 5)
-TableP <- rbind(TablePbr,TablePMultiATSM)
-row.names(TableP) <- c(RowsP,paste(RowsP," ",sep="")) # trick to avoid rows to have the same name
+TableP <- rbind(TablePbr, TablePMultiATSM)
+row.names(TableP) <- c(RowsP, paste(RowsP, " ", sep = "")) # trick to avoid rows to have the same name
 
@@
   kableExtra::row_spec(0, font_size = 14) %>%
-  kableExtra::add_header_above(c(" "= 1, "K0Z"=1, "K1Z" = 5), bold = T) %>%
+  kableExtra::add_header_above(c(" " = 1, "K0Z" = 1, "K1Z" = 5), bold = T) %>%
   kableExtra::pack_rows("BR (2017)", 1, 5) %>%
@@
 ```{r, echo= FALSE}
-data("BR_jps_gro_R3")
-data("JPSrep")
-
-se <- data.frame(BR_jps_out$est.llk$sigma.e, ModelPara$ests$se )
+se <- data.frame(BR_jps_out$est.llk$sigma.e, ModelPara$`JPS original`$US$ModEst$Q$se)
 rownames(se) <- "se"
-colnames(se) <- c("MultiATSM","BR (2017)")
+colnames(se) <- c("MultiATSM", "BR (2017)")
 se <- round(se, digits = 7)
 
-kableExtra::kbl(se, align = "c", caption ="Portfolio of yields with errors") %>%
+kableExtra::kbl(se, align = "c", caption = "Portfolio of yields with errors") %>%
   kableExtra::kable_classic("striped", full_width = F) %>%
@@
 
-For replicability, it is important to note that the $P-$dynamics results reported in Table \@ref(tab:PdynTab) for the *MultiATSM* package make use of the principal component weights provided by BR (2017).
+For replicability, it is important to note that the $P-$dynamics results reported in [Table&nbsp;1.2](#tab-PdynTab) for the *MultiATSM* package make use of the principal component weights provided by BR (2017).
 Such a matrix is simply a scaled-up version of the one provided by the function `pca_weights_one_country()` of this package.
@@
 
-# Candelon and Moura (forthcoming)
+# Candelon and Moura (2024)
 
@@
 OutputLabel <- "CM_jfec"
-DataFreq <-"Monthly"
+DataFreq <- "Monthly"
+Folder2Save <- NULL
 
-StatQ <- 0
+StatQ <- FALSE
 
@@
 #################################### GVAR-based models ##################################################
-GVARlist <- list( VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2004",
-                  t_Last_Wgvar = "2019")
+GVARlist <- list(
+  VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2004",
+  t_Last_Wgvar = "2019", DataConnectedness = TradeFlows
+)
 #################################### JLL-based models ###################################################
@@
 ###################################### BRW inputs  ######################################################
-WishBC <- 1
-BRWlist <- within(list(flag_mean = TRUE, gamma = 0.001, N_iter = 200, B = 50, checkBRW = TRUE,
-                       B_check = 1000),  N_burn <- round(N_iter * 0.15))
+WishBC <- TRUE
+BRWlist <- within(list(
+  Cent_Measure = "Mean", gamma = 0.001, N_iter = 200, B = 50, checkBRW = TRUE,
+  B_check = 1000, Eigen_rest = 1
+), N_burn <- round(N_iter * 0.15))
 
 # C) Decide on Settings for numerical outputs
-WishFPremia <- 1
-FPmatLim <- c(24,36)
+WishFPremia <- TRUE
+FPmatLim <- c(24, 36)
 
@@
 DesiredGraphs <- c("GIRF", "GFEVD", "TermPremia")
-WishGraphRiskFac <- 0
-WishGraphYields <- 1
-WishOrthoJLLgraphs <- 1
+WishGraphRiskFac <- FALSE
+WishGraphYields <- TRUE
+WishOrthoJLLgraphs <- TRUE
 
 # D) Bootstrap settings
-WishBootstrap <- 0 #  Set it to 1, if bootstrap is desired
-BootList <- list(methodBS = 'bs', BlockLength = 4, ndraws = 3, pctg =  95)
+WishBootstrap <- FALSE #  Set it to TRUE, if bootstrap is desired
+BootList <- list(methodBS = "bs", BlockLength = 4, ndraws = 1000, pctg = 95)
 
 # E) Out-of-sample forecast
-WishForecast <- 1
+WishForecast <- TRUE
 ForecastList <- list(ForHoriz = 12, t0Sample = 1, t0Forecast = 100, ForType = "Rolling")
@@
 FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
-if(any(ModelType == c("GVAR single", "GVAR multi"))){ GVARlist$DataConnectedness <- TradeFlows}
 
 # 3) Prepare the inputs of the likelihood function
-ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,
-                           FactorLabels, Economies, DataFreq, GVARlist, JLLlist, WishBC, BRWlist)
+ATSMInputs <- InputsForOpt(
+  t0_sample, tF_sample, ModelType, Yields, GlobalMacro, DomMacro,
+  FactorLabels, Economies, DataFreq, GVARlist, JLLlist, WishBC, BRWlist
+)
 
@@
 # a) Prepare list of inputs for graphs and numerical outputs
-InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
+InputsForOutputs <- InputsForOutputs(
+  ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
   WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishFPremia,
-                                    FPmatLim, WishBootstrap, BootList, WishForecast, ForecastList)
+  FPmatLim, WishBootstrap, BootList, WishForecast, ForecastList
+)
 
 # b) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
-NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)
+NumericalOutputs <- NumOutputs(
+  ModelType, ModelParaList, InputsForOutputs, FactorLabels,
+  Economies, Folder2Save
+)
 
 # c) Confidence intervals (bootstrap analysis)
-BootstrapAnalysis <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, Economies, InputsForOutputs,
-                               FactorLabels, JLLlist, GVARlist, WishBC, BRWlist)
+BootstrapAnalysis <- Bootstrap(
+  ModelType, ModelParaList, NumericalOutputs, Economies, InputsForOutputs,
+  FactorLabels, JLLlist, GVARlist, WishBC, BRWlist, Folder2Save
+)
 
 # 6) Out-of-sample forecasting
-Forecasts <- ForecastYields(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
-                            JLLlist, GVARlist, WishBC, BRWlist)
+Forecasts <- ForecastYields(
+  ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies,
+  JLLlist, GVARlist, WishBC, BRWlist, Folder2Save
+)
 ```
@@
 GlobalVar <- c("US_Output_growth", "China_Output_growth", "SP500")
-DomVar <- c("Inflation","Output_growth", "CDS", "COVID")
+DomVar <- c("Inflation", "Output_growth", "CDS", "COVID")
 N <- 2
@@
 OutputLabel <- "CM_EM"
-DataFreq <-"Weekly"
+DataFreq <- "Weekly"
+F2S <- NULL
 
-StatQ <- 0
+StatQ <- FALSE
 
@@
 #################################### GVAR-based models ##################################################
-GVARlist <- list( VARXtype = "constrained: COVID", W_type = "Sample Mean", t_First_Wgvar = "2015",
-                  t_Last_Wgvar = "2020")
+GVARlist <- list(
+  VARXtype = "constrained: COVID", W_type = "Sample Mean", t_First_Wgvar = "2015",
+  t_Last_Wgvar = "2020", DataConnectedness = TradeFlows_covid
+)
 ###################################### BRW inputs  ######################################################
-WishBC <- 0
+WishBC <- FALSE
 
 # C) Decide on Settings for numerical outputs
-WishFPremia <- 1
-FPmatLim <- c(47,48)
+WishFPremia <- TRUE
+FPmatLim <- c(47, 48)
 
@@
 DesiredGraphs <- c("GIRF", "GFEVD", "TermPremia")
-WishGraphRiskFac <- 0
-WishGraphYields <- 1
-WishOrthoJLLgraphs <- 0
+WishGraphRiskFac <- FALSE
+WishGraphYields <- TRUE
+WishOrthoJLLgraphs <- FALSE
 
 # D) Bootstrap settings
-WishBootstrap <- 1 #  YES: 1; No = 0.
-BootList <- list(methodBS = 'bs', BlockLength = 4, ndraws = 100, pctg =  95)
+WishBootstrap <- TRUE #  YES: 1; No = 0.
+BootList <- list(methodBS = "bs", BlockLength = 4, ndraws = 100, pctg = 95)
 
@@
 FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
-GVARlist$DataConnectedness <- Trade_Flows
 
 # 3) Prepare the inputs of the likelihood function
-ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacro, DomMacro, FactorLabels,
-                           Economies, DataFreq, GVARlist)
+ATSMInputs <- InputsForOpt(
+  t0_sample, tF_sample, ModelType, Yields_covid, GlobalMacro_covid, DomMacro_covid, FactorLabels,
+  Economies, DataFreq, GVARlist
+)
 
@@
 # a) Prepare list of inputs for graphs and numerical outputs
-InputsForOutputs <- InputsForOutputs(ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
+InputsForOutputs <- InputsForOutputs(
+  ModelType, Horiz, DesiredGraphs, OutputLabel, StatQ, DataFreq,
   WishGraphYields, WishGraphRiskFac, WishOrthoJLLgraphs, WishFPremia,
-                                    FPmatLim, WishBootstrap, BootList)
+  FPmatLim, WishBootstrap, BootList
+)
 
 # b) Fit, IRF, FEVD, GIRF, GFEVD, and Term Premia
-NumericalOutputs <- NumOutputs(ModelType, ModelParaList, InputsForOutputs, FactorLabels, Economies)
+NumericalOutputs <- NumOutputs(
+  ModelType, ModelParaList, InputsForOutputs, FactorLabels,
+  Economies, F2S
+)
 
@@
 BootstrapAnalysis <- Bootstrap(ModelType, ModelParaList, NumericalOutputs, Economies, InputsForOutputs,
-                               FactorLabels, JLLlist = NULL, GVARlist)
-
+  FactorLabels,
+  JLLlist = NULL, GVARlist, WishBC, BRWlist, F2S
+)
 ```

### 2025-10-23 R/A0N__BnAn.R

@@ -1,65 +1,66 @@
 #' Compute the cross-section loadings of yields of a canonical A0_N model
 #'
-#'@param mat       vector of maturities (J x 1). Maturities are in multiples of the discrete interval used in the model
-#'@param K1XQ      risk-neutral feedback matrix (N x N)
-#'@param ModelType string-vector containing the label of the model to be estimated
-#'@param dX        state loadings for the one-period rate (1xN). Default is a vector of ones
-#'@param r0        the long run risk neutral mean of the short rate (scalar)
-#'@param SSX       the covariance matrix of the errors (N x N)
-#'@param Economies string-vector containing the names of the economies which are part of the economic system
+#' @param mat       vector of maturities (J x 1).
+#' @param K1XQ      risk-neutral feedback matrix (N x N)
+#' @param ModelType string-vector containing the label of the model to be estimated
+#' @param r0        the long run risk neutral (scalar)
+#' @param SSX       covariance matrix of the latent states (N x N)
+#' @param Economies string-vector containing the names of the economies which are part of the economic system
 #'
-#'@keywords internal
 #'
-#'@return
-#' List containing:
-#' \itemize{
-#' \item Intercept (Jx1)
-#' \item slope (JxN)
-#' \item the betan (JX1, part of the intercepts unrelated to the long run risk neutral mean r0)
-#' coefficients of a canonical A_0(N) model.
-#' }
-#'@references
+#' @references
 #'  \itemize{
-#'  \item This function is a modified version of the "A0N__computeBnAn" function by Le and Singleton (2018). \cr
-#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
-#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
-#'  Available at: https://cepr.org/40029
-#'
-#'  \item Dai and Singleton (2000). "Specification Analysis of Affine Term Structure Models" (The Journal of Finance)
[…]

### 2025-10-23 R/BR_jps_out.R

@@ -1,0 +1,20 @@
+#' @title Replications of the JPS (2014) outputs by Bauer and Rudebusch (2017)
+#'
+#' @description Unspanned macro risk model outputs by Bauer and Rudebusch (2017)
+#' @name BR_jps_out
+#' @aliases BR_jps_out
+#' @docType data
+#' @usage data("BR_jps_out")
+#' @format Unspanned macro risk model outputs by Bauer and Rudebusch (2017)
+#' \describe{
+#' \item{est.llk}{summary list of log-likelihood estimations}
+#'    \item{M.o}{time series of unspanned factors}
+#'    \item{pars}{additional summary list of log-likelihood estimations}
+#'    \item{W}{Weight matrix that results from principal components analysis}
+#'    \item{Y}{time series of bond yields}
+#'    \item{N}{total number of  risk factor of the model (spanned and unspanned)}
+#'    \item{R}{total number of  spanned factor of the model}
+#'   }
+#' @references Bauer, M. and Rudebusch, G. "Resolving the Spanning Puzzle in Macro-Finance Term Structure Models"
+#' @keywords replication outputs
+NULL

### 2025-10-23 R/Bootstrap.R

@@ -1,17 +1,25 @@
 #' Generates the bootstrap-related outputs
 #'
-#' @param ModelType A character vector indicating the model type to be estimated.
-#' @param ModelParaPE A list containing the point estimates of the model parameters. For details, refer to the outputs from the \code{\link{Optimization}} function.
-#' @param NumOutPE The point estimate derived from numerical outputs. See the outputs from the \code{\link{NumOutputs}} function for further information.
-#' @param Economies A character vector containing the names of the economies included in the system.
-#' @param InputsForOutputs A list containing the necessary inputs for generating IRFs, GIRFs, FEVDs, GFEVDs and Term Premia.
-#' @param FactorLabels A list of character vectors with labels for all variables in the model.
-#' @param JLLlist List. Inputs for JLL model estimation (see \code{\link{JLL}} function). Default is NULL.
-#' @param GVARlist List. Inputs for GVAR model estimation (see \code{\link{GVAR}} function). Default is NULL.
-#' @param WishBC Whether to estimate the physical parameter model with bias correction, based on the method by Bauer, Rudebusch and Wu (2012) (see \code{\link{Bias_Correc_VAR}} function). Default is set to 0.
-#' @param BRWlist List of necessary inputs for performing the bias-corrected estimation (see \code{\link{Bias_Correc_VAR}} function).
-#' @param Folder2save Folder path where the outputs will be stored. Default option saves the outputs in a temporary directory.
-#' @param verbose Logical flag controlling function messaging. Default is TRUE.
+#' @param ModelType character. Model type to be estimated. Permissible choices: "JPS original", "JPS global", "GVAR single", "JPS multi", "GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".
+#' @param ModelParaPE list. Point estimates of the model parameters. See outputs from \code{\link{Optimization}}.
+#' @param NumOutPE list. Point estimates from numerical outputs. See outputs from \code{\link{NumOutputs}}.
[…]

### 2025-10-23 R/Buildx_xvec.R

@@ -1,262 +1,211 @@
-#' Obtain the auxiliary values corresponding to each parameter, its size and its name
-
-#'@param ListInputSet variable inputs used in the optimization (see "optimization" function)
-#'@param Economies string-vector containing the names of the economies which are part of the economic system
-#'@param FactorLabels list of necessary inputs for the estimation of JLL-based models (see \code{JLL} function)
-#'@param JLLinputs     Necessary inputs for the estimation of the JLL-based models
-#'
-#'
-#'@references
-#'  This function is a based on the "getx" function by Le and Singleton (2018). \cr
-#'  "A Small Package of Matlab Routines for the Estimation of Some Term Structure Models." \cr
-#'  (Euro Area Business Cycle Network Training School - Term Structure Modelling).
-#'
-#'@keywords internal
-
-
-Build_xvec <- function(ListInputSet, Economies, FactorLabels, JLLinputs = NULL){
-
-  # 0) Initializing the outputs
-  x <- c()
-  Dim_x <- c()
-  Name_x <- vector("character", length(ListInputSet) - 1)
-
-  # 1) Build list of interest
-  for (i in seq_len(length(ListInputSet) - 1)){
-    si <- unlist(gregexpr(":", ListInputSet[[i]]$Label))
-    Name_x[i] <- sub(".*@", "", substr(ListInputSet[[i]]$Label, start = 1, stop = si - 1)) # remove @ from variables name, if exists
-
-    Temp_x <- GetAuxPara(ListInputSet[[i]]$Value, ListInputSet[[i]]$Label, ListInputSet[[i]]$LB,
-                         ub = ListInputSet[[i]]$UB, Economies, FactorLabels, JLLinputs)
-
-    x <- rbind(x, t(t(as.vector(Temp_x))))
-    Dim_x <- rbind(Dim_x, dim(Temp_x))
-  }
-
-  Outputs <- list(x0 = x, Dim_x = Dim_x, Name_x = Name_x)
-  return(Outputs)
-}
-
 #################################################################################################################
-#################################################################################################################
-#' Map constrained parameters b to unconstrained auxiliary parameters a.
+#' Compute the auxiliary parameters a.
 #'
-#' @param ParaValue     Constrained parameter
+#' @param ParaValue     Parameter value
[…]

### 2025-10-23 R/CM_DomMacroFactors.R

@@ -1,10 +1,0 @@
-#'@title Data: Risk Factors - Candelon and Moura (2024, JFEC)
-#'
-#' @description Risk factors data used in Candelon and Moura (2024, JFEC)
-#' @name DomesticMacroVar
-#' @docType data
-#' @usage data("CM_DomMacroFactors")
-#' @format matrix containing the  risk factors of the models
-#' @references Candelon, B. and Moura, R. (2024) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
-#' @keywords Domestic unspanned risk factors
-NULL

### 2025-10-23 R/CM_DomMacroFactors_2023.R

@@ -1,10 +1,0 @@
-#'@title Data: Risk Factors for the GVAR - Candelon and Moura (2023)
-#'
-#' @description Risk factors data used in the GVAR models - Candelon and Moura (2023)
-#' @name DomMacro
-#' @docType data
-#' @usage data("CM_DomMacro_2023")
-#' @format list containing the variables used in the GVAR models
-#' @references Candelon, B. and Moura, R. (2023) "Sovereign yield curves and the COVID-19 in emerging markets". (Economic Modelling)
-#' @keywords Risk Factors GVAR
-NULL

### 2025-10-23 R/CM_Factors.R

@@ -1,10 +1,0 @@
-#'@title Data: Risk Factors - Candelon and Moura (2024, JFEC)
-#'
-#' @description Risk factors data used in Candelon and Moura (2024, JFEC)
-#' @name RiskFactors
-#' @docType data
-#' @usage data("CM_Factors")
-#' @format matrix containing the  risk factors of the models
-#' @references Candelon, B. and Moura, R. (2024) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
-#' @keywords Risk Factors
-NULL

### 2025-10-23 R/CM_Factors_GVAR.R

@@ -1,10 +1,0 @@
-#'@title Data: Risk Factors for the GVAR - Candelon and Moura (2024, JFEC)
-#'
-#' @description Risk factors data used in the GVAR models - Candelon and Moura (2024, JFEC)
-#' @name FactorsGVAR
-#' @docType data
-#' @usage data("CM_Factors_GVAR")
-#' @format list containing the variables used in the GVAR models
-#' @references Candelon, B. and Moura, R. (2024) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
-#' @keywords Risk Factors GVAR
-NULL

### 2025-10-23 R/CM_GlobalMacroFactors.R

@@ -1,10 +1,0 @@
-#'@title Data: Risk Factors - Candelon and Moura (2024, JFEC)
-#'
-#' @description Risk factors data used in Candelon and Moura (2024, JFEC)
-#' @name GlobalMacroVar
-#' @docType data
-#' @usage data("CM_GlobalMacroFactors")
-#' @format matrix containing the  risk factors of the models
-#' @references Candelon, B. and Moura, R. (2024) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
-#' @keywords Risk Factors
-NULL

### 2025-10-23 R/CM_GlobalMacro_2023.R

@@ -1,10 +1,0 @@
-#'@title Data: Risk Factors - Candelon and Moura (2023)
-#'
-#' @description Risk factors data used in Candelon and Moura (2023)
-#' @name GlobalMacro
-#' @docType data
-#' @usage data("CM_GlobalMacro_2023")
-#' @format matrix containing the  risk factors of the models
-#' @references Candelon, B. and Moura, R. (2023) "Sovereign yield curves and the COVID-19 in emerging markets". (Economic Modelling)
-#' @keywords Risk Factors
-NULL

### 2025-10-23 R/CM_Trade.R

@@ -1,10 +1,0 @@
-#'@title Data: Trade Flows - Candelon and Moura (2024, JFEC)
-#'
-#' @description Trade Flows data used in Candelon and Moura (2024, JFEC)
-#' @name TradeFlows
-#' @docType data
-#' @usage data("CM_Trade")
-#' @format list containing the bilateral trade flows
-#' @references Candelon, B. and Moura, R. (2024) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
-#' @keywords Trade Flows
-NULL

### 2025-10-23 R/CM_Trade_2023.R

@@ -1,10 +1,0 @@
-#'@title Data: Trade Flows - Candelon and Moura (2023)
-#'
-#' @description Trade Flows data used in Candelon and Moura (2023)
-#' @name Trade_Flows
-#' @docType data
-#' @usage data("CM_Trade")
-#' @format list containing the bilateral trade flows
-#' @references Candelon, B. and Moura, R. (2023) "Sovereign yield curves and the COVID-19 in emerging markets". (Economic Modelling)
-#' @keywords Trade Flows
-NULL

### 2025-10-23 R/CM_Yields.R

@@ -1,10 +1,0 @@
-#'@title Data: Yields - Candelon and Moura (2024, JFEC)
-#'
-#' @description Yields data used in Candelon and Moura (2024, JFEC)
-#' @name Yields
-#' @docType data
-#' @usage data("CM_Yields")
-#' @format matrix containing the Yields of the models
-#' @references Candelon, B. and Moura, R. (Forthcoming) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
-#' @keywords Yields
-NULL

### 2025-10-23 R/CM_Yields_2023.R

@@ -1,10 +1,0 @@
-#'@title Data: Yields - Candelon and Moura (2023)
-#'
-#' @description Bond yield data used in Candelon and Moura (2023)
-#' @name Yields
-#' @docType data
-#' @usage data("CM_Yields_2023")
-#' @format matrix containing the Yields of the models
-#' @references Candelon, B. and Moura, R. (2023) "Sovereign yield curves and the COVID-19 in emerging markets". (Economic Modelling)
-#' @keywords Yields
-NULL

### 2025-10-23 R/CM_repJPS.R

@@ -1,15 +1,0 @@
-#'@title Replications of the JPS (2014) outputs by the MultiATSM package
-#'
-#' @description Unspanned macro risk model outputs by the MultiATSM package
-#' @name ModelPara
-#' @docType data
-#' @usage data("JPSrep")
-#' @format list of inputs and outputs
-#' \describe{
-#'    \item{inputs}{general model inputs}
-#'    \item{ests}{model parameters estimates (JPS form)}
-#'    \item{llk}{log-likelihood of the observations}
-#'    \item{rot}{model parameters estimates (rotation form)}
-#'   }
-#' @keywords JPS (2014) outputs
-NULL

### 2025-10-23 R/ConfidenceBoundsBootstrap.R

@@ -11,7 +11,6 @@
 #' @keywords internal
 
 BootstrapBoundsSet <- function(ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies, Folder2save, verbose) {
-
   # Generate the graph paths and the graph folders
   FolderPath <- if (is.null(Folder2save)) tempdir() else Folder2save
   dir.create(paste(FolderPath, "/Outputs/", ModelType, "/Bootstrap", sep = ""))
@@ -25,10 +24,14 @@ BootstrapBoundsSet <- function(ModelType, ModelBootstrap, NumOutPE, InputsForOut
 
   AllNumOutputs <- list()
 
-  IRFandGIRF <- IRFandGIRFbs(ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies, PathsGraphs,
-                             Folder2save, verbose)
-  FEVDandGFEVD <- FEVDandGFEVDbs(ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies, PathsGraphs,
-                                 Folder2save, verbose)
+  IRFandGIRF <- IRFandGIRFbs(
+    ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies, PathsGraphs,
+    Folder2save, verbose
+  )
+  FEVDandGFEVD <- FEVDandGFEVDbs(
+    ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies, PathsGraphs,
+    Folder2save, verbose
+  )
 
   AllNumOutputs$Factors <- append(IRFandGIRF$Factors, FEVDandGFEVD$Factors)
   AllNumOutputs$Yields <- append(IRFandGIRF$Yields, FEVDandGFEVD$Yields)
@@ -54,7 +57,6 @@ BootstrapBoundsSet <- function(ModelType, ModelBootstrap, NumOutPE, InputsForOut
 
 IRFandGIRFbs <- function(ModelType, ModelBootstrap, NumOutPE, InputsForOutputs, Economies, PathsGraphs,
                          Folder2save, verbose) {
-
   ndraws <- InputsForOutputs[[ModelType]]$Bootstrap$ndraws
   pctg <- InputsForOutputs[[ModelType]]$Bootstrap$pctg
 
[…]

### 2025-10-23 R/DataForEstimation.R

@@ -1,62 +1,68 @@
-#' Retrieves data from Excel and build the database used in the model estimation
+#' Retrieves data from Excel and builds the database used in the model estimation
 #'
-#' @param t0 Start date of the sample period in the format yyyy-mm-dd.
-#' @param tF End date of the sample period in the format yyyy-mm-dd.
-#' @param Economies A character vector containing the names of the economies included in the system.
-#' @param N Integer. Number of country-specific spanned factors.
-#' @param FactorLabels String-list based which contains the labels of all the variables present in the model
-#' @param ModelType String-vector containing the label of the model to be estimated
-#' @param DataFrequency Character-based-vector. Available options are: "Daily All Days", "Daily Business Days", "Weekly", "Monthly", "Quarterly", "Annually"
-#' @param W_type Three possibilities:
-#' \itemize{
-#'      \item \code{Full Sample}: if one wishes ALL weight matrices of each year from which data is available (it may extrapolate the sample period);
-#'      \item \code{Sample Mean}: if one wishes a SINGLE weight matrix containing the average of weights over of the entire sample period;
-#'      \item Some year in particular (e.g. "1998", "2005" ...).
-#' }
-#' @param t_First_Wgvar Sample starting date (year)
-#' @param t_Last_Wgvar Sample last date (year)
-#' @param DataPathMacro Path of the Excel file containing the macroeconomic data (if any). The default is linked to the excel file present in the package.
-#' @param DataPathYields Path of the Excel file containing the yields data (if any). The default is linked to the excel file present in the package.
-#' @param DataPathTrade Path of the Excel file containing the trade data (if any). The default is linked to the excel file present in the package.
+#' @param t0 character. Start date of the sample period in the format yyyy-mm-dd.
+#' @param tF character. End date of the sample period in the format yyyy-mm-dd.
+#' @param Economies character vector. Names of the \code{C} economies included in the system.
+#' @param N positive integer. Number of country-specific spanned factors.
[…]

### 2025-10-23 R/Data_GVARFactors.R

@@ -1,0 +1,19 @@
+#' @title Data: Risk Factors for the GVAR - Candelon and Moura (2024, JFEC)
+#'
+#' @description Risk factors data used in the GVAR-ATSM from  Candelon and Moura (2024, JFEC)
+#' @name GVARFactors
+#' @aliases GVARFactors
+#' @docType data
+#' @usage data("GVARFactors")
+#' @format List of risk factors organized for GVAR estimation. It includes global unspanned factors (economic activity, inflation) and domestic factors—both unspanned (economic activity, inflation) and spanned (level, slope, curvature) with their starred counterparts.
+#' The dataset covers Brazil, China, Mexico, and Uruguay at a monthly frequency from June 2004 to January 2020.
+#' @source
+#' \describe{
+#' \item{Global unspanned factors}{ See \code{data("GlobalMacro")} for a detailed data description.}
+#' \item{Domestic unspanned factors}{See \code{data("DomMacro")} for a detailed data description.}
+#' \item{Domestic spanned factors}{First three principal components of each country set of bond yields. See \code{data("Yields")} for a detailed data description.}
+#' \item{Domestic star factors}{Weighted average of foreign factors. See \code{\link{Transition_Matrix}} for the computation of weights.}
+#' }
+#' @references Candelon, B. and Moura, R. (2024) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
+#' @keywords Risk Factors GVAR
+NULL

### 2025-10-23 R/Data_Trade.R

@@ -1,0 +1,13 @@
+#' @title Data: Trade Flows - Candelon and Moura (2024, JFEC)
+#'
+#' @description Trade Flows data used in Candelon and Moura (2024, JFEC)
+#' @name TradeFlows
+#' @aliases TradeFlows
+#' @docType data
+#' @usage data("TradeFlows")
+#' @format A list. Contains bilateral trade flow data for Brazil, China, Mexico, and Uruguay,
+#' covering the sample period from 1948 to 2019.
+#' @source International Monetary Fund - International Trade in Goods (IMTS) <https://data.imf.org/en/datasets/IMF.STA:IMTS>
+#' @references Candelon, B. and Moura, R. (2024) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
+#' @keywords Trade Flows
+NULL

### 2025-10-23 R/Data_Yields.R

@@ -1,0 +1,19 @@
+#' @title Data: bond yield data - Candelon and Moura (2024, JFEC)
+#'
+#' @description Yields data used in Candelon and Moura (2024, JFEC)
+#' @name Yields
+#' @aliases Yields
+#' @docType data
+#' @usage data("Yields")
+#' @format A matrix containing bond yields with maturities of 3, 6, 12, 36, 60, and 120 months for Brazil, China, Mexico, and Uruguay.
+#' The data are at monthly frequency and cover the period from June 2004, to January 2020.
+#' @source
+#' \describe{
+#' \item{Brazil}{Swap fixed-DI contracts (derivative securities indexed to the interbank loan rates). <https://www.b3.com.br/pt_br/market-data-e-indices/servicos-de-dados/market-data/consultas/mercado-de-derivativos/precos-referenciais/taxas-referenciais-bm-fbovespa/>}
+#' \item{China}{Government bond yield data in domestic currency. Simulated data constructed using Bloomberg bond yield series.}
+#' \item{Mexico}{Government bond yield data in domestic currency. Simulated data constructed using Bloomberg bond yield series.}
+#' \item{Uruguay}{Government bond yield data in US dollars <https://web.bevsa.com.uy/CurvasVectorPrecios/CurvasIndices/CUD.aspx>.}
+#' }
+#' @references Candelon, B. and Moura, R. (2024) "A Multicountry Model of the Term Structures of Interest Rates with a GVAR". (Journal of Financial Econometrics)
+#' @keywords Bond yields
+NULL

### 2025-10-23 R/DatabasePrep.R

@@ -1,68 +1,73 @@
 #' Gather data of several countries in a list. Particularly useful for GVAR-based setups (Compute "GVARFactors")
 #'
-#'@param t_First        Start date of the sample period in the format yyyy-mm-dd.
-#'@param t_Last         End date of the sample period in the format yyyy-mm-dd.
-#'@param Economies      A character vector containing the names of the economies included in the system.
-#'@param N              Integer. Number of country-specific spanned factors.
-#'@param FactorLabels   A list of character vectors with labels for all variables in the model.
-#'@param ModelType      A character vector indicating the model type to be estimated.
-#'@param Wgvar          GVAR transition matrix of size C x C, applicable if a GVAR-type model is selected. Default is NULL.
-#'@param DataPathMacro  File path to the Excel file containing macroeconomic data, if provided. The default path points to the Excel file available within the package.
-#'@param DataPathYields File path to the Excel file containing yields data, if provided. The default path points to the Excel file available within the package
+#' @param t_First character. Start date of the sample period in the format yyyy-mm-dd.
+#' @param t_Last character. End date of the sample period in the format yyyy-mm-dd.
+#' @param Economies character vector. Names of the \code{C} economies included in the system.
+#' @param N positive integer. Number of country-specific spanned factors per country.
+#' @param FactorLabels list. Labels for all variables present in the model, as returned by \code{\link{LabFac}}.
+#' @param ModelType character. Model type to be estimated. Permissible choices: "JPS original", "JPS global", "GVAR single", "JPS multi", "GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".
+#' @param Macro_FullData list. Full set of macroeconomic data, as returned by \code{\link{Load_Excel_Data}}.
+#' @param Yields_FullData list. Full set of bond yield data, as returned by \code{\link{Load_Excel_Data}}.
[…]

### 2025-10-23 R/Example_InpOut.R

@@ -1,10 +1,10 @@
-#'@title Example of list inputs used in the construction of several model outputs
+#' @title Example of list inputs used in the construction of several model outputs
 #'
 #' @description List of inputs of a JPS-based model for Brazilian data
 #' @name InpForOutEx
+#' @aliases InpForOutEx
 #' @docType data
 #' @usage data("InpForOutEx")
 #' @format list of inputs
-#'
-#' @keywords  JPS model - Brazil
+#' @keywords  list of inputs for an illustrative JPS model for Brazil
 NULL

### 2025-10-23 R/Examples_Out.R

@@ -1,7 +1,8 @@
-#'@title Complete list of several outputs from an ATSM
+#' @title Complete list of several outputs from an ATSM
 #'
 #' @description Example for illustration used in the package vignette
-#' @name Out
+#' @name Out_Example
+#' @aliases Out_Example
 #' @docType data
 #' @usage data("Out_Example")
 #' @format several model classes
@@ -12,5 +13,5 @@
 #'    \item{NumOut}{List of numerical outputs}
 #'    \item{Bootstrap}{List of set of outputs after bootstrap}
 #'   }
-#' @keywords examples
+#' @keywords Numerical outputs for an illustrative JPS model for Brazil
 NULL

### 2025-10-23 R/ForecastYields.R

@@ -1,16 +1,24 @@
 #' Generates forecasts of bond yields for all model types
 #'
-#' @param ModelType A character vector indicating the model type to be estimated.
-#' @param ModelPara A list containing the point estimates of the model parameters. For details, refer to the outputs from the \code{\link{Optimization}} function.
-#' @param InputsForOutputs A list containing the necessary inputs for generating IRFs, GIRFs, FEVDs, GFEVDs and Term Premia.
-#' @param FactorLabels A list of character vectors with labels for all variables in the model.
-#' @param Economies A character vector containing the names of the economies included in the system.
-#' @param JLLlist A list of necessary inputs for the estimation of JLL-based models (see the \code{\link{JLL}} function).
-#' @param GVARlist A list containing the necessary inputs for the estimation of GVAR-based models (see the \code{\link{GVAR}} function).
-#' @param WishBRW Whether to estimate the physical parameter model with bias correction, based on the method by Bauer, Rudebusch and Wu (2012) (see \code{\link{Bias_Correc_VAR}} function). Default is set to 0.
-#' @param BRWlist List of necessary inputs for performing the bias-corrected estimation (see \code{\link{Bias_Correc_VAR}} function).
-#' @param Folder2save Folder path where the outputs will be stored. Default option saves the outputs in a temporary directory.
-#' @param verbose Logical flag controlling function messaging. Default is TRUE.
+#' @param ModelType character. Model type to be estimated. Permissible choices: "JPS original", "JPS global", "GVAR single", "JPS multi", "GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".
+#' @param ModelPara list. Point estimates of the model parameters. See outputs from \code{\link{Optimization}}.
+#' @param InputsForOutputs list. Inputs for generating IRFs, GIRFs, FEVDs, GFEVDs, and Term Premia.
+#' @param FactorLabels list. Labels for all variables present in the model, as returned by \code{\link{LabFac}}.
+#' @param Economies character vector. Names of the \code{C} economies included in the system.
[…]

### 2025-10-23 R/FunctionML_vec.R

@@ -1,0 +1,450 @@
+#' Use function ML to generate the outputs from a ATSM
+#'
+#' @param x vector containing all the vectorized auxiliary parameters
+#' @param ML_fun vector-valued objective function (function)
+#' @param ListInputSet variable inputs used in the optimization (see inputs from "optimization" function)
+#' @param ModelType string-vector containing the label of the model to be estimated
+#' @param FactorLabels string-list based which contains the labels of all the variables present in the model
+#' @param Economies string-vector containing the names of the economies which are part of the economic system
+#' @param JLLinputs Set of necessary inputs used in the estimation of the JLL-based models (see "JLL" function)
+#' @param GVARinputs Set of necessary inputs used in the estimation of the GVAR-based models (see "GVAR" function)
+#' @param WithEstimation if TRUE, returns only the values of the likelihood function, else generates the entire set of outputs
+#'
+#'
+#' @keywords internal
+
+FunctionML_vec <- function(x, ML_fun, ListInputSet, ModelType, FactorLabels, Economies, JLLinputs,
+                           GVARinputs, WithEstimation) {
+  Para_Upds <- Update_ParaList(x, ModelType, FactorLabels, Economies, JLLinputs, GVARinputs, ListInputSet)
+
+  # Perform the numerical optimization
+  if (WithEstimation) {
+    if (any(ModelType == c("JLL original", "JLL No DomUnit"))) {
+      MLE <- ML_fun(K1XQ = Para_Upds$K1XQ$Value, ExportListOut = FALSE)
+    } else {
+      MLE <- ML_fun(K1XQ = Para_Upds$K1XQ$Value, SSZ = Para_Upds$SSZ$Value, ExportListOut = FALSE)
+    }
+  } else {
+    # Gather the estimated outputs after the optimization
+    if (any(ModelType == c("JLL original", "JLL No DomUnit"))) {
+      ListOut <- ML_fun(K1XQ = Para_Upds$K1XQ$Value)
+    } else {
+      ListOut <- ML_fun(K1XQ = Para_Upds$K1XQ$Value, SSZ = Para_Upds$SSZ$Value)
+    }
+  }
+
+  if (WithEstimation) {
+    return(MLE)
+  } else {
+    return(ListOut)
+  }
+}
+
+###########################################################################################################
[…]

### 2025-10-23 R/GVAR.R

@@ -1,60 +1,73 @@
-#' Estimates a GVAR(1) and a VARX(1,1,1) models
+#' Estimates a GVAR(1) and VARX(1,1,1) models
 #'
-#' @param GVARinputs List of inputs for GVAR model estimation:
+#' @param GVARinputs list. Inputs for GVAR model estimation:
 #' \enumerate{
-#'        \item  \code{Economies}:  A character vector containing the names of the economies included in the system.
-#'        \item  \code{GVARFactors}: A list of all variables used in the estimation of the VARX model \cr
-#'                (see e.g. \code{CM_Factors_GVAR} file for details);
-#'        \item \code{VARXtype}: A character vector with three possible options:
+#'        \item  \code{Economies}:  character vector. Contains the \code{C} names of the economies included in the system.
+#'        \item  \code{GVARFactors}: list. All variables used in the estimation of the VARX model \cr
+#'                (see e.g. \code{GVARFactors} file for details);
+#'        \item \code{VARXtype}: Permissible:
 #'  \itemize{
 #'        \item \code{'unconstrained'}: model is estimated without constraints (each equation is estimated individually by ordinary least square);
 #'        \item \code{'constrained: Spanned Factors'}: The model is estimated with the restriction that foreign pricing factors do NOT
-#'                                affect (i) domestic economic variables and (ii) domestic pricing factors. \cr
-#'                               (Equations are estimated using restricted least squares)
+#'                                affect (i) domestic economic variables and (ii) domestic pricing factors (estimation via restricted least squares).
 #'        \item \code{'constrained : [factor_name]'}: The model is estimated with the restriction that the specified risk factor
 #'              is influenced only by its own lagged values and the lagged values of its corresponding star variables.
-#'        (Equations are estimated using restricted least squares.)
+#'        (estimation via restricted least squares.)
 #'          }
[…]

### 2025-10-23 R/GraphicalOutputs.R

@@ -16,7 +16,6 @@
 
 GraphicalOutputs <- function(ModelType, ModelPara, NumOut, InputsForOutputs, Economies, FactorLabels,
                              Folder2save, verbose) {
-
   # Generate the graph paths and the graph folders
   dirs <- c("Outputs", paste0("Outputs/", ModelType), paste0("Outputs/", ModelType, "/Point Estimate"))
   lapply(file.path(Folder2save, dirs), dir.create, recursive = TRUE, showWarnings = FALSE)
@@ -24,9 +23,11 @@ GraphicalOutputs <- function(ModelType, ModelPara, NumOut, InputsForOutputs, Eco
   PathsGraphs <- FolderCreationPoint(ModelType, Economies, Folder2save)
 
   # 1) Plot the set of risk factors
-  RiskFactorsGraphs(ModelType, ModelPara, Economies, FactorLabels, Folder2save)
-
   if (verbose) message("2.3) Generating the graphs of interest")
+  RiskFactorsGraphs(
+    ModelType, InputsForOutputs[[ModelType]]$RiskFactors$WishGraphs, ModelPara, Economies,
+    FactorLabels, Folder2save, verbose
+  )
 
   if (ModelType %in% c("JPS original", "JPS global", "GVAR single")) {
     for (i in 1:length(Economies)) {
@@ -35,8 +36,10 @@ GraphicalOutputs <- function(ModelType, ModelPara, NumOut, InputsForOutputs, Eco
   }
 
   # 2) Model fit
-  Fitgraphs(ModelType, InputsForOutputs[[ModelType]]$Fit$WishGraphs, ModelPara, NumOut, Economies,
-            PathsGraphs, Folder2save, verbose)
+  Fitgraphs(
+    ModelType, InputsForOutputs[[ModelType]]$Fit$WishGraphs, ModelPara, NumOut, Economies,
+    PathsGraphs, Folder2save, verbose
+  )
 
   # 3) IRF and GIRF
   if (any(ModelType == c("JLL original", "JLL No DomUnit", "JLL joint Sigma"))) {
@@ -48,7 +51,9 @@ GraphicalOutputs <- function(ModelType, ModelPara, NumOut, InputsForOutputs, Eco
   for (j in 1:length(OutType)) {
     WG <- Wished_Graphs_IRFandGIRF(InputsForOutputs, OutType[j], ModelType)
[…]

### 2025-10-23 R/InputsForOpt.R

@@ -2,24 +2,60 @@
 #'
 #' @param InitialSampleDate Start date of the sample period in the format "dd-mm-yyyy"
 #' @param FinalSampleDate End date of the sample period in the format "dd-mm-yyyy"
-#' @param ModelType A character vector indicating the model type to be estimated. Available options: "JPS original", "JPS global", "GVAR single", "JPS multi", "GVAR multi", "JLL original", "JLL No DomUnit", "JLL joint Sigma".
-#' @param Yields A numerical matrix with time series of yields (JxT or CJ x T)
-#' @param GlobalMacro A numerical matrix with time series of the global risk factors (G x T)
-#' @param DomMacro A numerical matrix with time series of the country-specific risk factors for all C countries (CM x T)
-#' @param FactorLabels A list of character vectors with labels for all variables in the model.
-#' @param Economies A character vector containing the names of the economies included in the system.
-#' @param DataFrequency A character vector specifying the frequency of the data. Available options are: "Daily All Days", "Daily Business Days", "Weekly", "Monthly", "Quarterly", or "Annually".
-#' @param GVARlist A list containing the necessary inputs for the estimation of GVAR-based models
-#' @param JLLlist A list of necessary inputs for the estimation of JLL-based models. If the chosen model is "JLL original" or "JLL joint Sigma", then a dominant unit economy must be chosen. Otherwise, this list must be set as 'None'.
-#' @param WishBRW Logical. Whether to estimate the physical parameter model with bias correction, based on the method by Bauer, Rudebusch and Wu (2012). Default is FALSE.
-#' @param BRWlist List of necessary inputs for performing the bias-corrected estimation.
-#' @param UnitYields A character string indicating the maturity unit of yields. Options are: "Month" for yields expressed in months, or "Year" for yields expressed in years. Default is "Month".
-#' @param CheckInputs Logical. Whether to perform a prior check on the consistency of the provided input list. Default is TRUE.
[…]

### 2025-10-23 R/JLL.R

@@ -1,50 +1,57 @@
 #' Estimates the P-dynamics from JLL-based models
 #'
-#' @param NonOrthoFactors A numeric matrix (F x T) representing the time series of risk factors before the orthogonalization process.
-#' @param N Integer. Number of country-specific spanned factors.
-#' @param JLLinputs List of necessary inputs to estimate JLL models:
-#'  \enumerate{
-#'      \item Economies:  set of economies that are part of the economic system (string-vector)
-#'      \item \code{DomUnit}: A string specifying the name of the economy assigned as the dominant unit. \cr
-#'                  If no dominant unit is assigned, set this variable to "None".
-#'      \item \code{WishSigmas}: Set to "1" if the user wishes to estimate the variance-covariance matrices and Cholesky factorizations \cr
-#'                  (this can take a long time). Set to "0" if not.
-#'      \item \code{SigmaNonOrtho}: A NULL value or an F x F matrix from the non-orthogonalized dynamics.
-#'      \item \code{JLLModelType}: A string specifying the type of JLL model. Available options are: "JLL original", "JLL joint Sigma", or "JLL No DomUnit".
+#' @param NonOrthoFactors numeric matrix (\code{K x Td}). Time series of risk factors before orthogonalization.
+#' @param N positive integer. Number of country-specific spanned factors.
+#' @param JLLinputs list. Necessary inputs to estimate JLL models:
+#' \enumerate{
+#'   \item \code{Economies}: character vector. Set of \code{C} economies in the system.
+#'   \item \code{DomUnit}: character. Name of the dominant economy, or "None" if not assigned (for "JLL No DomUnit" model).
+#'   \item \code{WishSigmas}: logical. TRUE to estimate variance-covariance matrices and Cholesky factorizations; FALSE otherwise.
+#'   \item \code{SigmaNonOrtho}: NULL or F x F matrix from non-orthogonalized dynamics.
[…]

### 2025-10-23 R/NumOutputsBootstrap.R

@@ -9,7 +9,6 @@
 #' @keywords internal
 
 NumOutputs_Bootstrap <- function(ModelType, ModelParaBoot, InputsForOutputs, FactorLabels, Economies) {
-
   # IRF and GIRF
   IRFout <- IRFandGIRF_BS(ModelType, ModelParaBoot, InputsForOutputs[[ModelType]]$IRF$horiz, FactorLabels, Economies)
   # FEVD and GFEVD
@@ -35,7 +34,6 @@ NumOutputs_Bootstrap <- function(ModelType, ModelParaBoot, InputsForOutputs, Fac
 #' @keywords internal
 
 IRFandGIRF_BS <- function(ModelType, ModelParaBoot, IRFhoriz, FactorLabels, Economies) {
-
   C <- length(Economies)
   N <- length(FactorLabels$Spanned)
   G <- length(FactorLabels$Global)
@@ -47,19 +45,18 @@ IRFandGIRF_BS <- function(ModelType, ModelParaBoot, IRFhoriz, FactorLabels, Econ
 
   # 1) SINGLE COUNTRY MODELS
   if (ModelType %in% c("JPS original", "JPS global", "GVAR single")) {
-
     for (economy in Economies) {
       # a) Extract general parameters
       para_economy <- ModelParaBoot$ParaDraws[[ModelType]][[economy]]
       ndraws <- length(para_economy)
-      K <- nrow(para_economy[[1]]$ests$K1Z)
+      K <- nrow(para_economy[[1]]$ModEst$P$K1Z)
       J <- length(ModelParaBoot$GeneralInputs$mat)
-      YieldsLabel <- rownames(para_economy[[1]]$inputs$Y)
+      YieldsLabel <- rownames(para_economy[[1]]$Inputs$Y)
 
       # b) Compute IRFs and GIRFs
       results <- lapply(seq_len(ndraws), function(tt) {
         # Extract parameters of each draw
-        para_tt <- para_economy[[tt]]$ests
+        para_tt <- para_economy[[tt]]$ModEst$P
         SIGMA <- para_tt$SSZ # K x K (variance-covariance matrix)
         K1Z <- para_tt$K1Z # K x K (feedback matrix)
         G0.y <- para_tt$Gy.0
[…]

### 2025-10-23 R/Optimization.R

@@ -1,112 +1,89 @@
-#' Peform the minimization of mean(f)
+#' Perform the minimization of ML function
 #'
-#' @param f vector-valued objective function (function)
-#' @param ListInputSet list contain starting values and constraints:
-#'                        for each input argument K (of f), we need four inputs that look like:
+#' @param ML_fun vector-valued objective ML function
+#' @param ListInputSet list containing :
 #'    \enumerate{
-#'        \item a starting value: K0
-#'        \item a variable label ('K0') followed by a ':' followed by a type of constraint. The constraint can be:
+#'        \item a starting value for K1XQ and/or SSZ
+#'        \item a variable label among the following:
 #'                  \itemize{
-#'                      \item 'bounded': bounded matrix;
-#'                      \item 'Jordan' or 'Jordan MultiCountry': a matrix of Jordan type;
-#'                      \item 'psd': psd matrix;
-#'                      \item 'stationary': largest eigenvalue of the risk-neutral feedback matrix is strictly smaller than 1;
-#'                      \item 'diag' or 'BlockDiag': a diagonal or block diagonal matrix.
-#'                      \item 'JLLstructure': to impose the zero-restrictions on the variance-variance matrix along
+#'                      \item 'Jordan' or 'Jordan; stationary' for single countries setups or
+#'                      'Jordan MultiCountry' or 'Jordan MultiCountry; stationary'.  All cases related to the computation of a K1XQ parameter;
+#'                      \item 'psd': PSD matrix, used for JPS-based models. It relates to the SSZ parameter;
+#'                      \item  'BlockDiag': block diagonal matrix, used for JPS-based models. It relates to the SSZ parameter.
+#'                      \item 'JLLstructure': to impose the zero-restrictions on the SSZ term along
 #'                              the lines of the JLL models
 #'                          }
-#'        \item a lower bound lb (lb <- NULL -> no lower bound)
-#'        \item an upper bound ub (ub <- NULL -> no upper bound)
-#'        \item Specification of the optimization settings:
-#'        \itemize{
-#'                  \item 'iter off': hide the printouts of the numerical optimization routines;
[…]

### 2025-10-23 R/RiskFactorsPrep.R

@@ -1,6 +1,6 @@
 #' Builds the complete set of time series of the risk factors (spanned and unspanned)
 #'
-#' @param FactorSet Factor set list (see e.g. "CM_Factors_GVAR" data file)
+#' @param FactorSet Factor set list (see e.g. "GVARFactors" data file)
 #' @param Economies A character vector containing the names of the economies included in the system.
 #' @param FactorLabels A list of character vectors with labels for all variables in the model.
 #' @param Initial_Date Start date of the sample period in the format yyyy-mm-dd
@@ -9,15 +9,14 @@
 #'                      "Weekly", "Monthly", "Quarterly", "Annually".
 #'
 #'
-#'@keywords internal
+#' @keywords internal
 #'
 #' @return
 #' Risk factors used in the estimation of the desired ATSM
 #'
 
 
-RiskFactorsPrep <- function(FactorSet, Economies, FactorLabels, Initial_Date, Final_Date, DataFrequency){
-
+RiskFactorsPrep <- function(FactorSet, Economies, FactorLabels, Initial_Date, Final_Date, DataFrequency) {
   Initial_Date <- as.Date(Initial_Date)
   Final_Date <- as.Date(Final_Date)
 
@@ -28,56 +27,56 @@ RiskFactorsPrep <- function(FactorSet, Economies, FactorLabels, Initial_Date, Fi
   T_dim <- length(FactorSet[[1]][[1]][[1]])
 
 
-  ZZfull <- matrix(NA, nrow= C*(N+M)+G, ncol = T_dim)
-  ZZfull[seq_len(G),] <- do.call(rbind,lapply(FactorSet$Global, matrix, ncol=T_dim))
+  ZZfull <- matrix(NA, nrow = C * (N + M) + G, ncol = T_dim)
+  ZZfull[seq_len(G), ] <- do.call(rbind, lapply(FactorSet$Global, matrix, ncol = T_dim))
 
   # 1) Assign variables:
   idx0 <- G
-  for (i in 1:C){
-    idx1 <- idx0 + N+M
-    ZZfull[(idx0+1):idx1,] <- do.call(rbind,lapply(FactorSet[[Economies[i]]]$Factors[1:(N+M)], matrix, ncol=T_dim))
+  for (i in 1:C) {
+    idx1 <- idx0 + N + M
[…]

### 2025-10-23 R/Transition_Matrix.R

@@ -1,78 +1,83 @@
 #' Computes the transition matrix required in the estimation of the GVAR model
 #'
-#' @param t_First Sample starting date (in the format: yyyy).
-#' @param t_Last  Sample ending date (in the format: yyyy).
-#' @param Economies A character vector containing the names of the economies included in the system.
-#' @param type A character string indicating the method for computing interdependence. Possible options include:
+#' @param t_First character. Sample starting date (format: yyyy).
+#' @param t_Last character. Sample ending date (format: yyyy).
+#' @param Economies character vector. Names of the \code{C} economies included in the system.
+#' @param type character. Method for computing interdependence. Possible options:
 #' \itemize{
-#'      \item \code{Time-varying}: Computes time-varying interdependence and returns the weight matrices for each year based on available data (may extrapolate the sample period).
-#'      \item \code{Sample Mean}: Returns a single weight matrix containing the average weights over the entire sample period, suitable for time-invariant interdependence.
-#'      \item A specific year (e.g., "1998", "2005"): Used to compute time-invariant interdependence for the specified year.
+#'   \item \code{"Time-varying"}: Computes time-varying interdependence and returns weight matrices for each year.
+#'   \item \code{"Sample Mean"}: Returns a single weight matrix with average weights over the sample period.
+#'   \item Specific year (e.g., "1998", "2005"): Computes time-invariant interdependence for the specified year.
 #' }
-#' @param DataConnectedness Data used to compute the transition matrix. Default is set to NULL.
-#' @param DataPath Path to the Excel file containing the data (if applicable). The default is linked to the Excel file available in the package.
+#' @param DataConnectedness list or data frame. Data used to compute the transition matrix (e.g., trade flows).
 #'
-#' @return matrix or list of matrices
+#' @return matrix or list of matrices. Time-varying or time-invariant transition matrix depending on 'type'.
 #'
 #' @details
[…]

### 2025-10-23 R/VAR.R

@@ -1,46 +1,57 @@
 #' Estimates a standard VAR(1)
 #'
-#' @param RiskFactors A numeric matrix (F x T) representing the time series of risk factors.
-#' @param VARtype String vector with two possible values: 'unconstrained' or 'constrained'.
-#' @param Bcon_Mat Constraints matrix (F+1 x N), which includes an intercept. Entries containing NAs are treated as free parameters. \cr
-#'             Default is set to NULL.
+#' @param RiskFactors numeric matrix (\code{K x Td}). Time series of risk factors.
+#' @param VARtype character. Permissible choices: "unconstrained" or "constrained".
+#' @param Bcon_Mat matrix (\code{K x K + 1}). Constraints matrix (includes intercept). Entries containing NAs are treated as free parameters. Default is NULL.
 #'
-#' @return intercept, feedback matrix and the variance-covariance matrix of a VAR(1)
+#' @return list. Contains:
+#'   \itemize{
+#'   \item intercept (K x 1)
+#'   \item feedback matrix (K x K)
+#'   \item variance-covariance matrix (K x K) of a VAR(1)
+#' }
+#'
+#' @section General Notation:
+#' \itemize{
+#'   \item \code{Td}: model time series dimension
+#'   \item \code{N}: number of country-specific spanned factors
+#'   \item \code{K}: total number of risk factors
+#' }
+#'
 #' @examples
-#' data("CM_Factors")
+#' data(RiskFacFull)
 #' # Example 1: unconstrained case
-#' VAR(RiskFactors, VARtype= 'unconstrained')
+#' VAR_para1 <- VAR(RiskFacFull, VARtype = "unconstrained")
 #'
 #' # Example 2: constrained case
-#' K <- nrow(RiskFactors)
-#' Bcon_Mat <- matrix(0, nrow = K, ncol = K+1)
-#' Bcon_Mat[ , 1:3] <- NaN
-#' VAR(RiskFactors, VARtype= 'constrained', Bcon_Mat)
+#' K <- nrow(RiskFacFull)
+#' Bcon_Mat <- matrix(0, nrow = K, ncol = K + 1)
+#' Bcon_Mat[, 1:3] <- NaN
[…]

### 2025-10-23 R/meta-documentation.R

@@ -1,0 +1,34 @@
+#' Overview of Datasets Included in the MultiATSM Package
+#'
+#' The package includes several pre-processed datasets used for estimation and replication examples:
+#'
+#' \describe{
+#'   \item{GlobalMacro}{Global macro-financial risk factors, namely global economic activity and global inflation.}
+#'   \item{GlobalMacro_covid}{Global macro-financial risk factors, namely the output growth rate from the U.S. and China and the S&P 500 index.}
+#'   \item{DomMacro}{Domestic macroeconomic risk factors, namely economic activity and inflation.}
+#'   \item{DomMacro_covid}{Domestic macroeconomic risk factors, namely otput growth, inflation, CDS and the COVID-19 reproduction rate}
+#'   \item{TradeFlows}{Bilateral trade flow series used in GVAR examples as a proxy measure of cross-country conectdness.}
+#'   \item{TradeFlows_covid}{Bilateral trade flow series used in GVAR examples as a proxy measure of cross-country conectdness.}
+#'   \item{Yields}{Monthly series of bond yields by maturity for multiple economies.}
+#'   \item{Yields_covid}{Weekly series of sovereign bond yields by maturity for multiple economies.}
+#'   \item{RiskFacFull}{Full set of risk factors (global and domestic) data used throughout the package}
+#'   \item{GVARFactors}{List of risk factors used in the estimation of GVAR models.}
+#'   \item{BR_jps_out}{Replications of the JPS outputs by Bauer and Rudebusch (2017)}
+#'   \item{InpForOutEx}{List of inputs for an illustrative JPS model with Brazilian data}
+#'   \item{ParaSetEx}{List of set of parameterafter optimization for an illustrative JPS model with Brazilian data}
+#'   \item{NumOutEx}{List of numerical outputs for an illustrative JPS model with Brazilian data}
+#'   \item{Out_Example}{re-loaded examaple of a complete list of several model outputs. Used in the package vignette.}
+#' }
+#'
+#' Each dataset is documented separately using `?GlobalMacro`, `?DomMacro`, `?TradeFlows`, `?Yields`, etc.
+#' Datasets ending with the suffix \code{_covid} are based on those used in Candelon and Moura (2023) and cover Brazil, India, Mexico, and Russia.
+#' The remaining datasets correspond to Candelon and Moura (2024) and include Brazil, China, Mexico, and Uruguay.
+#' @references
[…]

### 2025-10-23 inst/doc/Guidelines.R

@@ -1,373 +1,0 @@
-## -----------------------------------------------------------------------------
-library(MultiATSM)
-
-## ----ModFea, message=FALSE, echo=FALSE----------------------------------------
-ModelLabels <- c("JPS original", "JPS global", "JPS multi", "GVAR single", "GVAR multi", 
-                 "JLL original", "JLL No DomUnit", "JLL joint Sigma")
-
-# Rows
-Tab <- data.frame(matrix(nrow = length(ModelLabels), ncol = 0)) 
-rownames(Tab) <- ModelLabels
-
-# Empty columns
-EmptyCol <- c("", "", "", "", "", "", "", "") 
-Tab$EmptyCol0 <- EmptyCol
-# P-dynamics + 2 empty spaces
-Tab$PdynIndUnco <- c("x", "", "", "", "", "", "", "")
-Tab$PdynIndCo <- c("", "", "", "", "", "", "", "")
-Tab$PdynJointUnco <- c("", "x", "x", "", "", "", "", "")
-Tab$PdynJointJLL <- c("", "", "", "", "", "x", "x", "x")
-Tab$PdynJointGVAR <- c("", "", "", "x", "x", "", "", "")
-Tab$EmptyCol1 <- EmptyCol
-Tab$EmptyCol2 <- EmptyCol
-# Q-dynamics + 2 empty spaces
-Tab$QdynInd <- c("x", "x", "", "x", "", "", "", "")   
-Tab$QdynJoint <- c("", "", "x", "", "x", "x", "x", "x") 
-Tab$EmptyCol3 <- EmptyCol
-Tab$EmptyCol4 <- EmptyCol
-# Sigma + 2 empty spaces
-Tab$Ponly <-  c("", "", "", "", "", "x", "x", "")
-Tab$PandQ <- c("x", "x", "x", "x", "x", "", "", "x")
-Tab$EmptyCol5 <- EmptyCol
-Tab$EmptyCol6 <- EmptyCol
-# Dominant Unit
-Tab$DomUnit <- c("", "", "", "", "", "x", "", "x")
-
-# Adjust column names
-ColNames <- c("","","","","JLL", "GVAR", "", "", "", "", "", "", "","", "", "","")
-colnames(Tab) <- ColNames
-
-
-# Generate the table
-suppressWarnings(library(magrittr))
-
-
-kableExtra::kbl(Tab, align = "c", caption = "Model Features") %>%
[…]

### 2025-10-23 inst/doc/MultiATSM.R

@@ -1,0 +1,408 @@
+## -----------------------------------------------------------------------------
+library(MultiATSM)
+
+## ----ModFea, message=FALSE, echo=FALSE----------------------------------------
+ModelLabels <- c(
+  "JPS original", "JPS global", "JPS multi", "GVAR single", "GVAR multi",
+  "JLL original", "JLL No DomUnit", "JLL joint Sigma"
+)
+
+# Rows
+Tab <- data.frame(matrix(nrow = length(ModelLabels), ncol = 0))
+rownames(Tab) <- ModelLabels
+
+# Empty columns
+EmptyCol <- c("", "", "", "", "", "", "", "")
+Tab$EmptyCol0 <- EmptyCol
+# P-dynamics + 2 empty spaces
+Tab$PdynIndUnco <- c("x", "", "", "", "", "", "", "")
+Tab$PdynIndCo <- c("", "", "", "", "", "", "", "")
+Tab$PdynJointUnco <- c("", "x", "x", "", "", "", "", "")
+Tab$PdynJointJLL <- c("", "", "", "", "", "x", "x", "x")
+Tab$PdynJointGVAR <- c("", "", "", "x", "x", "", "", "")
+Tab$EmptyCol1 <- EmptyCol
+Tab$EmptyCol2 <- EmptyCol
+# Q-dynamics + 2 empty spaces
+Tab$QdynInd <- c("x", "x", "", "x", "", "", "", "")
+Tab$QdynJoint <- c("", "", "x", "", "x", "x", "x", "x")
+Tab$EmptyCol3 <- EmptyCol
+Tab$EmptyCol4 <- EmptyCol
+# Sigma + 2 empty spaces
+Tab$Ponly <- c("", "", "", "", "", "x", "x", "")
+Tab$PandQ <- c("x", "x", "x", "x", "x", "", "", "x")
+Tab$EmptyCol5 <- EmptyCol
+Tab$EmptyCol6 <- EmptyCol
+# Dominant Unit
+Tab$DomUnit <- c("", "", "", "", "", "x", "", "x")
+
+# Adjust column names
+ColNames <- c("", "", "", "", "JLL", "GVAR", "", "", "", "", "", "", "", "", "", "", "")
+colnames(Tab) <- ColNames
+
+
+# Generate the table
+suppressWarnings(library(magrittr))
+
+
+kableExtra::kbl(
+  Tab,
+  align = "c",
+  caption = "Model Features"
+) %>%
[…]

### 2025-10-23 inst/doc/Paper-Replications.R

@@ -1,117 +1,120 @@
 ## -----------------------------------------------------------------------------
 library(MultiATSM)
 
-## ----eval=FALSE---------------------------------------------------------------
-# # A) Load database data
-# LoadData("BR_2017")
-# 
-# # B) GENERAL model inputs
-# ModelType <- "JPS original"
-# 
-# Economies <- c("US") # Names of the economies from the economic system
-# GlobalVar <- c() # Global Variables
-# DomVar <- c("GRO", "INF") # Country-specific variables
-# N <- 3 # Number of spanned factors per country
-# 
-# t0_sample <- "January-1985"
-# tF_sample <- "December-2007"
-# 
-# DataFreq <- "Monthly" # Frequency of the data
-# 
-# StatQ <- 0 # Stationary condition
-# 
-# #########################################################################################################
-# ############################### NO NEED TO MAKE CHANGES FROM HERE #######################################
-# #########################################################################################################
-# # 2) Minor preliminary work
-# FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
-# 
-# Yields <- t(BR_jps_out$Y)
-# DomesticMacroVar <- t(BR_jps_out$M.o)
-# GlobalMacroVar <- c()
-# 
-# # 3) Prepare the inputs of the likelihood function
-# ATSMInputs <- InputsForOpt(t0_sample, tF_sample, ModelType, Yields, GlobalMacroVar, DomesticMacroVar,
-#                            FactorLabels, Economies, DataFreq)
-# 
-# # 4) Optimization of the model
-# ModelPara <- Optimization(ATSMInputs, StatQ, DataFreq, FactorLabels, Economies, ModelType)
+## -----------------------------------------------------------------------------
+# A) Load database data
+LoadData("BR_2017")
+
+# B) GENERAL model inputs
+ModelType <- "JPS original"
+
+Economies <- c("US") # Names of the economies from the economic system
+GlobalVar <- c() # Global Variables
[…]

### 2025-10-23 tests/testthat/test_DataForEstimation.R

@@ -1,0 +1,72 @@
+library(testthat)
+library(MultiATSM)
+
+
+#  Retrieve data from excel files
+MacroAllData <- Load_Excel_Data(system.file("extdata", "MacroData.xlsx", package = "MultiATSM"))
+YieldsAllData <- Load_Excel_Data(system.file("extdata", "YieldsData.xlsx", package = "MultiATSM"))
+
+# Define common inputs of interest
+DomVar <- c("Eco_Act", "Inflation")
+GlobalVar <- c("GBC", "CPI_OECD")
+t0 <- "2006-09-01"
+tF <- "2011-09-01"
+Economies <- c("China", "Brazil")
+N <- 1
+
+# 1) Test output structure for non-GVAR model
+test_that("DataForEstimation returns correct structure for non-GVAR model", {
+  ModelType <- "JPS original"
+  K <- (N + length(DomVar)) * length(Economies) + length(GlobalVar)
+  FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  DataFrequency <- "Monthly"
+  res <- DataForEstimation(
+    t0, tF, Economies, N, FactorLabels, ModelType, DataFrequency,
+    MacroAllData, YieldsAllData
+  )
+  expect_type(res, "list")
+  expect_true(all(c("Yields", "RiskFactors", "GVARFactors") %in% names(res)))
+  expect_null(res$GVARFactors)
+  expect_equal(nrow(res$RiskFactors), K)
+})
+
+
+# 2) Test output structure for GVAR-based models
+test_that("DataForEstimation returns correct structure for GVAR model", {
+  ModelType <- "GVAR multi"
+  FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  DataFrequency <- "Monthly"
+  data("CM_Trade")
+  res <- DataForEstimation(t0, tF, Economies, N, FactorLabels, ModelType, DataFrequency, MacroAllData,
+    YieldsAllData, TradeFlows,
+    W_type = "Sample Mean",
+    t_First_Wgvar = "2006", t_Last_Wgvar = "2006"
+  )
+  expect_type(res, "list")
+  expect_true(all(c("Yields", "RiskFactors", "GVARFactors") %in% names(res)))
+  expect_type(res$GVARFactors, "list")
+  expect_equal(names(res$GVARFactors), c(Economies, "Global"))
+})
+
+
+# 3) Test error for invalid input
+# a) Type 1
[…]

### 2025-10-23 tests/testthat/test_DatabasePrep.R

@@ -1,0 +1,67 @@
+library(testthat)
+library(MultiATSM)
+
+#  1) Load all available data from excel files
+# a) Macro and bond yield data
+
+MacroDataAll <- Load_Excel_Data(system.file("extdata", "MacroData.xlsx", package = "MultiATSM"))
+YieldsDataAll <- Load_Excel_Data(system.file("extdata", "YieldsData.xlsx", package = "MultiATSM"))
+Trade_Alldata <- Load_Excel_Data(system.file("extdata", "TradeData.xlsx", package = "MultiATSM"))
+
+# Adjust trade data
+DataConnect <- lapply(Trade_Alldata, function(df) {
+  countries <- df[[1]]
+  df <- as.data.frame(df[-1])
+  rownames(df) <- countries
+  df
+})
+
+
+# 2) Select desired inputs
+Economies <- c("China", "Brazil", "Mexico", "Russia")
+DomVar <- c("Eco_Act", "Inflation")
+GlobalVar <- c("GBC", "CPI_OECD")
+N <- 1
+t0 <- "2007-09-01"
+tF <- "2019-01-01"
+
+
+############ 1) Test output structure ############
+test_that("DatabasePrep returns correct structure", {
+  ModelType <- "GVAR multi"
+  FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  Wgvar <- Transition_Matrix(t_First = "2006", t_Last = "2019", Economies, type = "Sample Mean", DataConnect)
+  res <- DatabasePrep(t0, tF, Economies, N, FactorLabels, ModelType, MacroDataAll, YieldsDataAll, Wgvar)
+  expect_type(res, "list")
+  expect_true(all(Economies %in% names(res)))
+  expect_true("Global" %in% names(res))
+})
+
+
+############ 2) Test error for missing transition matrix ############
+test_that("DatabasePrep issue an error for missing Wgvar a GVAR-based model", {
+  ModelType <- "GVAR multi"
+  FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  expect_error(DatabasePrep(t0, tF, Economies, N, FactorLabels, ModelType, MacroDataAll, YieldsDataAll, Wgvar = NULL))
+})
+
+
+############# 3) Test for all model types, but GVAR-based one ############
+test_that("DatabasePrep checks overall structure of non GVAR-based models", {
[…]

### 2025-10-23 tests/testthat/test_GVAR.R

@@ -1,0 +1,58 @@
+library(testthat)
+library(MultiATSM)
+
+# Load inputs
+data(GVARFactors)
+
+GVARinputs <- list(
+  Economies = c("China", "Brazil", "Mexico", "Uruguay"),
+  GVARFactors = GVARFactors, VARXtype = "unconstrained"
+)
+
+GVARinputs$Wgvar <- matrix(c(
+  0, 0.83, 0.86, 0.38,
+  0.65, 0, 0.13, 0.55,
+  0.32, 0.12, 0, 0.07,
+  0.03, 0.05, 0.01, 0
+), nrow = 4, ncol = 4)
+N <- 3
+
+
+# 1) Test output structure
+test_that("GVAR returns correct output structure", {
+  res <- GVAR(GVARinputs, N)
+  expect_type(res, "list")
+  expect_true(all(c("VARX", "Gy.0", "F0", "F1", "Sigma_y") %in% names(res)))
+  # Test for the absence of  NA, NaN, or Inf
+  expect_false(any(!is.finite(unlist(res))))
+})
+
+
+# 2) Test error for inconsistent N
+test_that("GVAR issues an error for inconsistent N", {
+  N <- 10
+  expect_error(GVAR(GVARinputs, N, CheckInputs = TRUE))
+})
+
+# 3) Test for restricted VARXtype elements
+test_that("GVAR checks for a constrained model version (Spanned Factors)", {
+  G <- length(GVARFactors$Global)
+  GVARinputs$VARXtype <- "constrained: Spanned Factors"
+  res <- GVAR(GVARinputs, N, CheckInputs = TRUE)
+  expect_true(any(res$F1[-(1:G), ] == 0))
+})
+
+
+# 4) Test for restricted VARXtype elements
+test_that("GVAR checks for a constrained model version (Inflation)", {
+  GVARinputs$VARXtype <- "constrained: Inflation"
+  res <- GVAR(GVARinputs, N, CheckInputs = TRUE)
+  infl_rows <- grep("Inflation", rownames(res$F1))
+  expect_true(length(infl_rows) > 0)
+})
+
+# 5) Test error for missing required GVARinputs elements
+test_that("GVAR throws error for invalid (incomplete) VARXtype element", {
+  GVARinputs$VARXtype <- "constrained"
+  expect_error(GVAR(GVARinputs, N, CheckInputs = TRUE))
+})

### 2025-10-23 tests/testthat/test_InputsForOpt.R

@@ -1,0 +1,200 @@
+library(testthat)
+library(MultiATSM)
+
+# Define inputs
+LoadData("CM_2024")
+InitialSampleDate <- "01-05-2007"
+FinalSampleDate <- "01-12-2018"
+DataFrequency <- "Monthly"
+GlobalVar <- c("Gl_Eco_Act")
+DomVar <- c("Eco_Act", "Inflation")
+N <- 2
+
+
+# 1) Test output structure (Model type "JPS original")
+# a) No bias correction
+test_that("InputsForOpt returns correct output structure (JPS model)", {
+  ModelType <- "JPS original"
+  Economies <- "Mexico"
+  FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  res <- InputsForOpt(InitialSampleDate, FinalSampleDate, ModelType, Yields, GlobalMacro, DomMacro,
+    FactorLabels, Economies, DataFrequency,
+    CheckInputs = FALSE, verbose = FALSE
+  )
+  expect_type(res, "list")
+  expect_s3_class(res, "ATSMModelInputs")
+  expect_true(all(c(
+    "mat", "Wpca", "We", "WpcaFull", "Yields", "SpaFact", "RiskFactors", "Gy.0", "K1XQ",
+    "SSZ", "K0Z", "K1Z"
+  ) %in% names(res$Mexico)))
+
+  # Test methods
+  out_print <- capture.output(print(res))
+  expect_true(length(out_print) > 0)
+  expect_type(out_print, "character")
+
+  out_summary <- capture.output(summary(res))
+  expect_true(length(out_summary) > 0)
+  expect_type(out_summary, "character")
+})
+
+# b) With bias correction
+test_that("InputsForOpt returns correct output structure (JPS model with bias correction)", {
+  ModelType <- "JPS original"
+  Economies <- "Mexico"
+  WishBC <- TRUE
+  BRWlist <- within(list(
+    Cent_Measure = "Mean", gamma = 0.1, N_iter = 10, B = 20, checkBRW = TRUE,
+    B_check = 1000, Eigen_rest = 1
+  ), N_burn <- round(N_iter * 0.15))
+
+  FactorLabels <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  res <- InputsForOpt(InitialSampleDate, FinalSampleDate, ModelType, Yields, GlobalMacro, DomMacro,
[…]

### 2025-10-23 tests/testthat/test_InputsForOutputs.R

@@ -1,0 +1,87 @@
+library(testthat)
+library(MultiATSM)
+
+# 1) Test basic output structure
+test_that("InputsForOutputs returns correct structure and options", {
+  ModelType <- "JPS original"
+  Horiz <- 100
+  DesiredOutputGraphs <- c("Fit", "GIRF", "GFEVD")
+  OutputLabel <- "Test"
+  WishStationarityQ <- TRUE
+  DataFrequency <- "Monthly"
+  WishGraphYields <- TRUE
+  WishGraphRiskFactors <- FALSE
+  res <- InputsForOutputs(
+    ModelType, Horiz, DesiredOutputGraphs, OutputLabel,
+    WishStationarityQ, DataFrequency, WishGraphYields, WishGraphRiskFactors
+  )
+  expect_type(res, "list")
+  expect_equal(res$`Label Outputs`, OutputLabel)
+  expect_equal(res$StationaryQ, WishStationarityQ)
+  expect_equal(res$DataFreq, DataFrequency)
+  expect_true(ModelType %in% names(res))
+  expect_true("Fit" %in% names(res[[ModelType]]))
+})
+
+
+# 2) Test JLL model
+test_that("InputsForOutputs returns correct structure for a JLL model", {
+  ModelType <- "JLL original"
+  Horiz <- 10
+  DesiredOutputGraphs <- c("IRF", "FEVD")
+  OutputLabel <- "Test"
+  WishStationarityQ <- TRUE
+  DataFrequency <- "Monthly"
+  WishGraphYields <- TRUE
+  WishGraphRiskFactors <- FALSE
+  WishOrthoJLLgraphs <- TRUE
+  res <- InputsForOutputs(
+    ModelType, Horiz, DesiredOutputGraphs, OutputLabel,
+    WishStationarityQ, DataFrequency, WishGraphYields, WishGraphRiskFactors
+  )
+  expect_type(res, "list")
+  expect_equal(res$`Label Outputs`, OutputLabel)
+  expect_equal(res$StationaryQ, WishStationarityQ)
+  expect_equal(res$DataFreq, DataFrequency)
+  expect_true(ModelType %in% names(res))
+  expect_true("Fit" %in% names(res[[ModelType]]))
+})
+
+
+
+# 3) Test bootstrap options
+test_that("InputsForOutputs handles bootstrap options", {
+  ModelType <- "JPS original"
+  Horiz <- 100
[…]

### 2025-10-23 tests/testthat/test_LabFac.R

@@ -1,0 +1,47 @@
+library(testthat)
+library(MultiATSM)
+
+
+# 1) Test basic label generation
+test_that("LabFac generates correct labels for typical input", {
+  N <- 2
+  DomVar <- c("inflation", "Output gap")
+  GlobalVar <- "Commodity Prices"
+  Economies <- c("U.S.", "Canada")
+  ModelType <- "JPS original"
+  res <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  expect_type(res, "list")
+  expect_equal(res$Spanned, c("Level", "Slope"))
+  expect_equal(res$Domestic, c("inflation", "Output gap", "Level", "Slope"))
+  expect_equal(res$Global, "Commodity Prices")
+  expect_equal(res$Tables$AllCountries, c(
+    "inflation U.S.", "Output gap U.S.", "Level U.S.", "Slope U.S.",
+    "inflation Canada", "Output gap Canada", "Level Canada", "Slope Canada"
+  ))
+})
+
+
+# 2) Test error for invalid N
+test_that("LabFac issues an error for invalid (large) N", {
+  N <- 10
+  DomVar <- c("inflation")
+  GlobalVar <- "Commodity Prices"
+  Economies <- c("U.S.")
+  ModelType <- "JPS original"
+  expect_error(LabFac(N, DomVar, GlobalVar, Economies, ModelType))
+})
+
+
+# 3) Test JLL model type labels
+test_that("LabFac generates JLL labels when ModelType is JLL", {
+  N <- 1
+  DomVar <- c("inflation")
+  GlobalVar <- "Commodity Prices"
+  Economies <- c("U.S.")
+  ModelType <- "JLL original"
+  res <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  expect_type(res, "list")
+  expect_equal(res$Spanned, c("Level"))
+  expect_equal(res$Domestic, c("inflation", "Level"))
+  expect_equal(res$Tables$AllCountriesJLL, c("inflation U.S. JLL", "Level U.S. JLL"))
+})

### 2025-10-23 tests/testthat/test_LoadData.R

@@ -1,0 +1,12 @@
+library(testthat)
+library(MultiATSM)
+
+
+test_that("LoadData loads correct datasets for each paper", {
+  # Test all valid cases
+  valid_cases <- c("BR_2017", "CM_2024", "CM_2023")
+
+  for (case in valid_cases) {
+    expect_silent(LoadData(case))
+  }
+})

### 2025-10-23 tests/testthat/test_Load_Excel_Data.R

@@ -1,0 +1,87 @@
+library(testthat)
+library(MultiATSM)
+
+
+# Unit test for Load_Excel_Data
+test_that("Load_Excel_Data loads package example files correctly", {
+  # Test with actual package data files
+  macro_file <- system.file("extdata", "MacroData.xlsx", package = "MultiATSM")
+  yields_file <- system.file("extdata", "YieldsData.xlsx", package = "MultiATSM")
+
+  # Test that files exist (sanity check)
+  expect_true(file.exists(macro_file))
+  expect_true(file.exists(yields_file))
+
+  # Test function with package files
+  expect_silent(macro_data <- Load_Excel_Data(macro_file))
+  expect_silent(yields_data <- Load_Excel_Data(yields_file))
+
+  # Test basic structure
+  expect_type(macro_data, "list")
+  expect_type(yields_data, "list")
+  expect_true(length(macro_data) > 0)
+  expect_true(length(yields_data) > 0)
+  expect_named(macro_data)
+  expect_named(yields_data)
+})
+
+test_that("Load_Excel_Data handles invalid inputs", {
+  # Test with non-existent file
+  expect_error(Load_Excel_Data("nonexistent_file.xlsx"))
+
+  # Test with NULL input
+  expect_error(Load_Excel_Data(NULL))
+
+  # Test with non-Excel file
+  temp_file <- tempfile(fileext = ".txt")
+  writeLines("test", temp_file)
+  expect_error(Load_Excel_Data(temp_file))
+  unlink(temp_file)
+})
+
+test_that("Load_Excel_Data returns correct structure", {
+  macro_file <- system.file("extdata", "MacroData.xlsx", package = "MultiATSM")
+  skip_if_not(file.exists(macro_file))
+
+  result <- Load_Excel_Data(macro_file)
+
+  # Test generic structure (without specific content checks)
+  expect_type(result, "list")
+  expect_true(length(result) > 0)
+  expect_true(all(sapply(result, is.data.frame)))
+  expect_named(result)
+})
+
+test_that("Load_Excel_Data preserves sheet names and order", {
+  # Use an existing file from your package that has multiple sheets
+  test_file <- system.file("extdata", "MacroData.xlsx", package = "MultiATSM")
+  skip_if_not(file.exists(test_file))
+
+  # First, check what sheets exist in the actual file
+  actual_sheets <- readxl::excel_sheets(test_file)
+  skip_if(length(actual_sheets) < 2) # Skip if file doesn't have multiple sheets
+
[…]

### 2025-10-23 tests/testthat/test_Opt_NumOut_Boot_Graphs_Forecast_GVAR.R

@@ -1,0 +1,186 @@
+library(testthat)
+library(MultiATSM)
+
+
+# 0) Load inputs
+extract_ggplots <- function(x) {
+  if (inherits(x, "ggplot")) {
+    list(x)
+  } else if (is.list(x)) {
+    unlist(lapply(x, extract_ggplots), recursive = FALSE)
+  } else {
+    list()
+  }
+}
+
+# Inputs required for the optimization
+LoadData("CM_2024")
+Economies <- c("Brazil", "Mexico")
+ModelType <- "GVAR multi"
+t0 <- "01-05-2007"
+tF <- "01-12-2018"
+N <- 1
+GlobalVar <- "Gl_Eco_Act"
+DomVar <- "Eco_Act"
+DataFreq <- "Monthly"
+StatQ <- FALSE
+WishFP <- FALSE
+FPLim <- c()
+WishBC <- FALSE
+
+# Inputs required for the computation of the numerical outputs
+Horiz <- 10
+OutputLabel <- "Model_demo"
+DesiredGraphs <- c()
+WGYields <- FALSE
+WGFac <- FALSE
+WGJLL <- FALSE
+
+# Bootstrap settings
+WishBoot <- TRUE
+BootList <- list(methodBS = "bs", BlockLength = 4, ndraws = 3, pctg = 95)
+
+# Forecasting setting
+WishFor <- TRUE
+ForList <- list(ForHoriz = 6, t0Sample = 1, t0Forecast = 132, ForType = "Expanding")
+
+test_that("Optimization + Outputs + Graphs return correct structure (GVAR model)", {
+  GVARlist <- list(
+    VARXtype = "unconstrained", W_type = "Sample Mean", t_First_Wgvar = "2005",
+    t_Last_Wgvar = "2019", DataConnectedness = TradeFlows
+  )
+
+
+  FacLab <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  ATSMInputs <- InputsForOpt(
+    t0, tF, ModelType, Yields, GlobalMacro, DomMacro,
+    FacLab, Economies, DataFreq, GVARlist,
+    CheckInputs = FALSE, verbose = FALSE
+  )
+
+  res_Opt <- Optimization(ATSMInputs, StatQ, DataFreq, FacLab, Economies, ModelType, verbose = FALSE)
+
+  # --- A) Optimization structure ---
+  expect_type(res_Opt, "list")
[…]

### 2025-10-23 tests/testthat/test_Opt_NumOut_Boot_Graphs_Forecast_JLL.R

@@ -1,0 +1,183 @@
+library(testthat)
+library(MultiATSM)
+
+
+# 0) Load inputs
+extract_ggplots <- function(x) {
+  if (inherits(x, "ggplot")) {
+    list(x)
+  } else if (is.list(x)) {
+    unlist(lapply(x, extract_ggplots), recursive = FALSE)
+  } else {
+    list()
+  }
+}
+
+# Inputs required for the optimization
+LoadData("CM_2024")
+Economies <- c("China", "Mexico")
+ModelType <- "JLL original"
+t0 <- "01-05-2007"
+tF <- "01-12-2018"
+N <- 1
+GlobalVar <- "Gl_Eco_Act"
+DomVar <- "Eco_Act"
+DataFreq <- "Monthly"
+StatQ <- FALSE
+WishFP <- TRUE
+FPLim <- c(24, 36)
+
+WishBC <- FALSE
+JLLlist <- list(DomUnit = "China")
+GVARlist <- NULL
+
+# Inputs required for the computation of the numerical outputs
+Horiz <- 10
+OutputLabel <- "Model_demo"
+DesiredGraphs <- c()
+WGYields <- FALSE
+WGFac <- FALSE
+WGJLL <- FALSE
+
+# Bootstrap settings
+WishBoot <- TRUE
+BootList <- list(methodBS = "bs", BlockLength = 4, ndraws = 3, pctg = 95)
+
+# Forecasting setting
+WishFor <- TRUE
+ForList <- list(ForHoriz = 6, t0Sample = 1, t0Forecast = 132, ForType = "Expanding")
+
+
+test_that("Optimization + Outputs + Graphs return correct structure (JLL model)", {
+  # Run optimization
+  FacLab <- LabFac(N, DomVar, GlobalVar, Economies, ModelType)
+  ATSMInputs <- InputsForOpt(
+    t0, tF, ModelType, Yields, GlobalMacro, DomMacro, FacLab, Economies,
+    DataFreq, GVARlist, JLLlist,
+    CheckInputs = FALSE, verbose = FALSE
+  )
+
+  res_Opt <- Optimization(ATSMInputs, StatQ, DataFreq, FacLab, Economies, ModelType, verbose = FALSE)
+
+  # --- A) Optimization structure ---
+  expect_type(res_Opt, "list")
+  expect_s3_class(res_Opt, "ATSMModelOutputs")
[…]

### 2025-10-23 tests/testthat/test_Opt_NumOut_Boot_Graphs_Forecast_JPS.R

@@ -1,0 +1,145 @@
+library(testthat)
+library(MultiATSM)
+
+
+# 0) Load inputs
+extract_ggplots <- function(x) {
+  if (inherits(x, "ggplot")) {
+    list(x)
+  } else if (is.list(x)) {
+    unlist(lapply(x, extract_ggplots), recursive = FALSE)
+  } else {
+    list()
+  }
+}
+
+# Inputs required for the optimization
+LoadData("CM_2024")
+Economy <- "Brazil"
+ModelType <- "JPS original"
+t0 <- "01-05-2007"
+tF <- "01-12-2018"
+N <- 2
+GlobalVar <- "Gl_Eco_Act"
+DomVar <- "Eco_Act"
+DataFreq <- "Monthly"
+StatQ <- TRUE
+WishFP <- FALSE
+FPLim <- c(24, 36)
+WishBC <- FALSE
+
+# Inputs required for the computation of the numerical outputs
+Horiz <- 10
+OutputLabel <- "Model_demo"
+DesiredGraphs <- c()
+WGYields <- FALSE
+WGFac <- FALSE
+WGJLL <- FALSE
+
+# Bootstrap settings
+WishBoot <- TRUE
+BootList <- list(methodBS = "bs", BlockLength = 4, ndraws = 3, pctg = 95)
+
+# Forecasting setting
+WishFor <- TRUE
+ForList <- list(ForHoriz = 6, t0Sample = 1, t0Forecast = 132, ForType = "Expanding")
+
+
+# 1) Set of tests
+test_that("Optimization + Outputs + Graphs return correct structure (JPS model)", {
+  # Run optimization
+  FacLab <- LabFac(N, DomVar, GlobalVar, Economy, ModelType)
+  ATSMInputs <- InputsForOpt(
+    t0, tF, ModelType, Yields, GlobalMacro, DomMacro,
+    FacLab, Economy, DataFreq,
+    CheckInputs = FALSE, verbose = FALSE
+  )
+
+  res_Opt <- Optimization(ATSMInputs, StatQ, DataFreq, FacLab, Economy, ModelType, verbose = FALSE)
+
+  # --- A) Optimization structure ---
+  expect_type(res_Opt, "list")
+  expect_s3_class(res_Opt, "ATSMModelOutputs")
+  out_summary <- capture.output(summary(res_Opt))
+  expect_true(length(out_summary) > 0)
+  expect_type(out_summary, "character")
+
[…]

### 2025-10-23 tests/testthat/test_Transition_Matrix.R

@@ -1,0 +1,39 @@
+library(testthat)
+library(MultiATSM)
+
+# Set inputs
+LoadData("CM_2024")
+Economies <- c("China", "Brazil", "Mexico", "Uruguay")
+
+# 1) Test Sample Mean output
+test_that("Transition_Matrix returns correct output for Sample Mean", {
+  W_mat <- Transition_Matrix("2000", "2002", Economies, "Sample Mean", DataConnectedness = TradeFlows)
+  expect_type(W_mat, "double")
+  expect_equal(dim(W_mat), c(length(Economies), length(Economies)))
+  expect_equal(rownames(W_mat), Economies)
+  expect_equal(colnames(W_mat), Economies)
+})
+
+# 2) Test Time-varying output
+test_that("Transition_Matrix returns correct output for Time-varying", {
+  W_list <- Transition_Matrix("2000", "2002", Economies, "Time-varying", DataConnectedness = TradeFlows)
+  expect_type(W_list, "list")
+  expect_true(all(sapply(W_list, function(x) all(dim(x) == c(length(Economies), length(Economies))))))
+})
+
+
+# 3) Test specific year output
+test_that("Transition_Matrix returns correct output for specific year", {
+  W_year <- Transition_Matrix("2000", "2002", Economies, "2001", DataConnectedness = TradeFlows)
+  expect_type(W_year, "double")
+  expect_equal(dim(W_year), c(length(Economies), length(Economies)))
+  expect_equal(rownames(W_year), Economies)
+  expect_equal(colnames(W_year), Economies)
+})
+
+
+# 4) Test error for missing data
+test_that("Transition_Matrix returns NA matrix for missing data", {
+  W_mat <- Transition_Matrix("1992", "2002", Economies, "Sample Mean", DataConnectedness = TradeFlows)
+  expect_true(all(is.na(W_mat)))
+})

### 2025-10-23 tests/testthat/test_VAR.R

@@ -1,0 +1,38 @@
+library(testthat)
+library(MultiATSM)
+
+# Load data
+LoadData("CM_2024")
+RiskFactors <- rbind(GlobalMacro, DomMacro[1:4, ])
+
+# 1) Test unconstrained VAR
+test_that("VAR returns correct output for unconstrained case", {
+  K <- nrow(RiskFactors)
+  res <- VAR(RiskFactors, VARtype = "unconstrained")
+  expect_type(res, "list")
+  expect_true(all(c("K0Z", "K1Z", "SSZ") %in% names(res)))
+  expect_equal(dim(res$K0Z), c(K, 1))
+  expect_equal(dim(res$K1Z), c(K, K))
+  expect_equal(dim(res$SSZ), c(K, K))
+  expect_false(anyNA(res)) # tests whether all entries are numeric
+})
+
+
+# 2) Test constrained VAR
+test_that("VAR returns correct output for constrained case", {
+  K <- nrow(RiskFactors)
+  Bcon_Mat <- matrix(0, nrow = K, ncol = K + 1)
+  Bcon_Mat[, 1:3] <- NaN
+  res <- VAR(RiskFactors, VARtype = "constrained", Bcon_Mat)
+  expect_type(res, "list")
+  expect_true(all(c("K0Z", "K1Z", "SSZ") %in% names(res)))
+  expect_equal(dim(res$K0Z), c(K, 1))
+  expect_equal(dim(res$K1Z), c(K, K))
+  expect_equal(dim(res$SSZ), c(K, K))
+  expect_false(anyNA(res)) # tests whether all entries are numeric
+})
+
+# 3) Test error for invalid VARtype
+test_that("VAR throws error for invalid VARtype", {
+  expect_error(VAR(RiskFactors, VARtype = "invalid"))
+})

### 2025-10-23 tests/testthat/test_pca_weights_one_country.R

@@ -1,0 +1,33 @@
+library(testthat)
+library(MultiATSM)
+
+# Simulate yield data for testing
+LoadData("CM_2024")
+Yields <- Yields[1:18, ]
+Economies <- c("China", "Brazil", "Mexico")
+
+# 1) Test output dimensions, type and absences of NAs
+test_that("pca_weights_one_country returns correct dimensions and type", {
+  J <- nrow(Yields) / length(Economies)
+  for (i in 1:length(Economies)) {
+    W <- pca_weights_one_country(Yields, Economies[i])
+    expect_type(W, "double")
+    expect_equal(dim(W), c(J, J))
+    expect_false(anyNA(W))
+  }
+})
+
+
+# 2) Test sign adjustment for level
+test_that("pca_weights_one_country properly adjusts sign for level weights", {
+  for (i in 1:length(Economies)) {
+    W <- pca_weights_one_country(Yields, Economies[i])
+    expect_true(all(W[1, ] >= 0) || all(W[1, ] <= 0))
+  }
+})
+
+
+# 3) Test error for missing economy
+test_that("pca_weights_one_country issues an error for missing economy", {
+  expect_error(pca_weights_one_country(Yields, "India"))
+})
